### 2021-03-29 proposal/proposal.Rmd
outlined project proposal

---
output: 
  stevetemplates::article:
    fig_caption: true
#bibliography: master.bib
urlcolor: blue
biblio-style: apsr
title: "Using Statistical Methods to Predict the Outcome of Basketball Games"
author:
- name: Sabrina Ball
  affiliation:
- name: Joseph Despres
  affiliation:
abstract: "Quick Summary"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
---

```{r eval=FALSE, include=FALSE}
op <- par(mar = rep(0, 4))
plot(imager::load.image("https://i.imgur.com/f0KNGaD.png"))
par(op)
```

```{r, include=FALSE}
source(here::here("R", "explatoryAnalysis.R"))
```

# 1. Data

Our data comes from three different data sources Kaggle, NCAA, and a March Madness bracket. First, a Kaggle^[https://www.kaggle.com/c/ncaam-march-mania-2021/data] machine learning challenge that has a comprehensive records of every US division I basketball game played from 2003 to 2020. Second, the NCAA^[http://stats.ncaa.org/rankings/change_sport_year_div] team statistics. Third, the tournament bracket in a nice .xls format we obtained from a blog^[https://plexkits.com/march-madness-bracket/].

## Variables

Our response is the outcome win/loss as a function of team statistics. 
win 
`r #TODO: fill this out maybe make this a list if its too big`

|  Variable name     |  Description  |   Type  |
| ------------------ | ------------- | ------- |
| win                |               |         |
| fg_percent         |               |         |
| opposingfg_percent |               |         |
| ft_percent         |               |         |
| opposingft_percent |               |         |
| rpg                |               |         |
| opposingrpg        |               |         |
| st                 |               |         |
| opposingst         |               |         |
| to                 |               |         |
| opposingto         |               |         |
| opposingbkpg       |               |         |
| bkpg               |               |         |

# 2. Questions

Can we use statistical methods to predict outcomes of the March Madness basketball games better than chance? Better than seed? Better than betting markets?

# 3. Learning

We would like to learn how to make predictions with logistic regression. We would like to come up with a good way to evaluate our predictions. This will be dificult since the outcome is binary so we would need to weight our predictions somehow and index our socres.

# 4. Methods

Logistic regression with a response prediction function.

# 5. Issues

The seeds are likely derrived from the predictors we are using. Multicolinearity, will be something to watchout for because better teams probally have better statistics.

### 2021-03-29 R/bracket.R
getting bracket without pockets

@@ -5,6 +5,9 @@ library(tidyverse)
 team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
 bracket <- readxl::read_excel(here::here("data", "prebracket.xlsx"))
 
+begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
+
+
 # Even better
 # bracket <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1zaWZ2Xh7sll-PZKA1DhKvS6TdfE0Yggk0m5OWY4aICc/edit#gid=0")
 # remember API times out...
@@ -59,3 +62,27 @@ bracket <- bracket %>%
 
 bracket
 
+
+# begining bracket (no pockets) -------------------------------------------
+
+begining_bracket <- begining_bracket %>% 
+  select(2, 35) %>% 
+  stack() %>% 
+  drop_na() %>% 
+  tibble() %>% 
+  rename(team = values) %>% 
+  mutate(team = tolower(team)) %>% 
+  mutate(game = 1:n(),
+         game = as.numeric(game),
+         otherteam  = if_else(game %% 2 == 1, team, lag(team)),
+         game = if_else(game %% 2 == 1, game + 1, game), .before = team) %>% 
+  filter(team != otherteam) %>% 
+  filter(otherteam != "first round") %>% 
+  left_join(team_names, by = c("team" = "TeamNameSpelling")) %>% 
+  rename(teamid = TeamID) %>% 
+  left_join(team_names, by = c("otherteam" = "TeamNameSpelling")) %>%
+  rename(otherteamid = TeamID) %>% 
+  mutate(teamid = replace(teamid, team == "app st.", 1111)) %>% 
+  mutate_if(is.numeric, as.character) %>% 
+  select(-ind, -game)
+  

### 2021-03-29 R/bracket.R
fixed data issue

@@ -65,6 +65,9 @@ bracket
 
 # begining bracket (no pockets) -------------------------------------------
 
+
+
+
 begining_bracket <- begining_bracket %>% 
   select(2, 35) %>% 
   stack() %>% 
@@ -86,3 +89,4 @@ begining_bracket <- begining_bracket %>%
   mutate_if(is.numeric, as.character) %>% 
   select(-ind, -game)
   
+begining_bracket

### 2021-03-29 R/modeling.R
re-ran model on bracket without pockets

@@ -33,9 +33,11 @@ fit <- glm(win ~ x3fg +
            data = merged, family = "binomial")
 summary(fit)
 
+begining_bracket
+
 s2021 <- team_stats %>% 
   filter(season == 2021, 
-         TeamID %in% c(bracket$teamid, bracket$otherteamid)) %>% 
+         TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
   distinct(TeamID, .keep_all = TRUE) %>% 
   select(-season) 
 
@@ -50,26 +52,29 @@ predictor_fn <- function (team_1_id, team_2_id) {
   return(pred)
 }
 
-possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = 0)
+possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")
 
 # app state is missing from the data
 # that 0 puts a zero pros of them winning to effectively drop them.
-
-probs <- map2_dbl(.x = bracket$teamid, 
-                  .y = bracket$otherteamid, 
+begining_bracket
+probs <- map2_dbl(.x = begining_bracket$teamid, 
+                  .y = begining_bracket$otherteamid, 
                   .f = possibly_predictor_fn)
 
-bracket <- bracket %>% 
+probs
+
+begining_bracket <- begining_bracket%>% 
   mutate(team_prob = probs,
          otherteam_prob = 1 - team_prob, 
          game = paste0(team, " vs. ", otherteam),
          predicted_winner = if_else(team_prob > 0.5, team, otherteam),
          predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid))
   
-games <- bracket %>% 
+games <- begining_bracket %>% 
   mutate(team_prob = as.numeric(probs),
          otherteam_prob = 1 - team_prob, 
          game = paste0(team, " vs. ", otherteam)) %>% 
   select(game, team_prob, otherteam_prob)
 
 games
+

### 2021-03-29 R/modeling.R
added predicitons

@@ -52,6 +52,9 @@ predictor_fn <- function (team_1_id, team_2_id) {
   return(pred)
 }
 
+
+# First round -------------------------------------------------------------
+
 possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")
 
 # app state is missing from the data
@@ -78,3 +81,156 @@ games <- begining_bracket %>%
 
 games
 
+
+# second round ------------------------------------------------------------
+
+second_round <- begining_bracket %>% 
+  select(predicted_winner, predicted_winnerid) %>% 
+  rename(team = predicted_winner, teamid = predicted_winnerid) %>% 
+  mutate(game = rep(c("team", "otherteam"), nrow(begining_bracket)/2))
+
+team <- second_round %>% 
+  filter(game != "team")
+
+other <- second_round %>% 
+  filter(game == "team")
+
+probs <- map2_dbl(.x = team$teamid, 
+                  .y = other$teamid, 
+                  .f = possibly_predictor_fn)
+
+second_round_prediction <- other %>% 
+  rename_all(~ paste0("opposing",.)) %>% 
+  bind_cols(team) %>% 
+  mutate(team_prob = as.numeric(probs),
+         otherteam_prob = 1 - team_prob, 
+         game = paste0(team, " vs. ", opposingteam), 
+         predicted_winner = if_else(team_prob > 0.5, team, opposingteam),
+         predicted_winnerid = if_else(team_prob > 0.5, teamid, opposingteamid))
+
+second_round_prediction %>% 
+  select(predicted_winner)
+
+# sweet sixteen -------------------------------------------------------------
+
+winner_selector <- function(df){
+    output <- df %>% 
+    select(predicted_winner, predicted_winnerid) %>% 
+    rename(team = predicted_winner, teamid = predicted_winnerid) %>% 
+    mutate(game = rep(c("team", "otherteam"), nrow(df)/2))
+    return(output)
+}
+
+sweet_sixteen <- winner_selector(df = second_round_prediction) 
+
+team <- sweet_sixteen %>% 
+  filter(game != "team")
+
+other <- sweet_sixteen %>% 
+  filter(game == "team")
+
+probs <- map2_dbl(.x = team$teamid, 
+                  .y = other$teamid, 
+                  .f = possibly_predictor_fn)
+
+sweet_sixteen <- other %>% 
+  rename_all(~ paste0("opposing",.)) %>% 
+  bind_cols(team) %>% 
[…]

### 2021-03-29 proposal/proposal.Rmd
adding proposal

@@
 ```{r, include=FALSE}
-source(here::here("R", "explatoryAnalysis.R"))
+# source(here::here("R", "explatoryAnalysis.R"))
 ```
@@
 # 1. Data
-
-Our data comes from three different data sources Kaggle, NCAA, and a March Madness bracket.
-First, a Kaggle^[https://www.kaggle.com/c/ncaam-march-mania-2021/data] machine learning challenge that has a comprehensive records of every US division I basketball game played from 2003 to 2020.
-Second, the NCAA^[http://stats.ncaa.org/rankings/change_sport_year_div] team statistics.
-Third, the tournament bracket in a nice .xls format we obtained from a blog^[https://plexkits.com/march-madness-bracket/].
+ ^[https://plexkits.com/march-madness-bracket/].
 
@@
 
-Our response is the outcome win/loss as a function of team statistics. 
-win 
-`r #TODO: fill this out maybe make this a list if its too big`
-
-|  Variable name     |  Description  |   Type  |
-| ------------------ | ------------- | ------- |
-| win                |               |         |
-| fg_percent         |               |         |
-| opposingfg_percent |               |         |
-| ft_percent         |               |         |
-| opposingft_percent |               |         |
-| rpg                |               |         |
-| opposingrpg        |               |         |
-| st                 |               |         |
-| opposingst         |               |         |
-| to                 |               |         |
-| opposingto         |               |         |
-| opposingbkpg       |               |         |
-| bkpg               |               |         |
+Our response is the outcome win/loss as a function of team statistics. Win or loss coded 0 or 1. field goal precentage, 
+win,fg_percent, opposingfg_percent, ft_percent, opposingft_percent, rpg, opposingrpg, st, opposingst, to, opposingto, opposingbkpg, bkpg
 

### 2021-03-29 proposal/proposal.tex
adding proposal

\documentclass[11pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\title{Using Statistical Methods to Predict the Outcome of Basketball
Games  }

\author{\Large Sabrina
Ball\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Joseph
Despres\vspace{0.05in} \newline\normalsize\emph{}  }

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Sabrina Ball () and Joseph Despres ()},
             pdfkeywords = {},
            pdftitle={Using Statistical Methods to Predict the Outcome
of Basketball Games},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

% CSL environment change -----

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1
%
% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12}
\textbf{\authorfont Sabrina
Ball} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Joseph
Despres} \hskip 15pt \emph{\small }

}

}

\begin{abstract}

    \hbox{\vrule height .2pt width 39.14pc}

    \vskip 8.5pt % \small

\noindent Quick Summary

    \hbox{\vrule height .2pt width 39.14pc}

\end{abstract}

\vskip -8.5pt

 % removetitleabstract

\noindent

\hypertarget{data}{%
\section{1. Data}\label{data}}

Our data comes from three different data sources Kaggle, NCAA, and a
March Madness bracket. First, a Kaggle\footnote{\url{https://www.kaggle.com/c/ncaam-march-mania-2021/data}}
machine learning challenge that has a comprehensive records of every US
division I basketball game played from 2003 to 2020. Second, the
NCAA\footnote{\url{http://stats.ncaa.org/rankings/change_sport_year_div}}
team statistics. Third, the tournament bracket in a nice .xls format we
obtained from a blog\footnote{\url{https://plexkits.com/march-madness-bracket/}}.

\hypertarget{variables}{%
\subsection{Variables}\label{variables}}

Our response is the outcome win/loss as a function of team statistics.
win \textbar{} Variable name \textbar{} Description \textbar{} Type
\textbar{} \textbar{} ------------------ \textbar{} -------------
\textbar{} ------- \textbar{} \textbar{} win \textbar{} \textbar{}
\textbar{} \textbar{} fg\_percent \textbar{} \textbar{} \textbar{}
\textbar{} opposingfg\_percent \textbar{} \textbar{} \textbar{}
\textbar{} ft\_percent \textbar{} \textbar{} \textbar{} \textbar{}
opposingft\_percent \textbar{} \textbar{} \textbar{} \textbar{} rpg
\textbar{} \textbar{} \textbar{} \textbar{} opposingrpg \textbar{}
\textbar{} \textbar{} \textbar{} st \textbar{} \textbar{} \textbar{}
\textbar{} opposingst \textbar{} \textbar{} \textbar{} \textbar{} to
\textbar{} \textbar{} \textbar{} \textbar{} opposingto \textbar{}
\textbar{} \textbar{} \textbar{} opposingbkpg \textbar{} \textbar{}
\textbar{} \textbar{} bkpg \textbar{} \textbar{} \textbar{}

\hypertarget{questions}{%
\section{2. Questions}\label{questions}}

Can we use statistical methods to predict outcomes of the March Madness
basketball games better than chance? Better than seed? Better than
betting markets?

\hypertarget{learning}{%
\section{3. Learning}\label{learning}}

We would like to learn how to make predictions with logistic regression.
We would like to come up with a good way to evaluate our predictions.
This will be dificult since the outcome is binary so we would need to
weight our predictions somehow and index our socres.

\hypertarget{methods}{%
\section{4. Methods}\label{methods}}

Logistic regression with a response prediction function.

\hypertarget{issues}{%
\section{5. Issues}\label{issues}}

The seeds are likely derrived from the predictors we are using.
Multicolinearity, will be something to watchout for because better teams
probally have better statistics.

\newpage
\singlespacing
\end{document}

### 2021-03-30 R/modeling.R
cleaning up the code breaking it into functions

@@ -33,8 +33,6 @@ fit <- glm(win ~ x3fg +
            data = merged, family = "binomial")
 summary(fit)
 
-begining_bracket
-
 s2021 <- team_stats %>% 
   filter(season == 2021, 
          TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
@@ -54,62 +52,94 @@ predictor_fn <- function (team_1_id, team_2_id) {
 
 
 # First round -------------------------------------------------------------
+begining_bracket
 
 possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")
 
-# app state is missing from the data
-# that 0 puts a zero pros of them winning to effectively drop them.
-begining_bracket
 probs <- map2_dbl(.x = begining_bracket$teamid, 
                   .y = begining_bracket$otherteamid, 
                   .f = possibly_predictor_fn)
 
 probs
 
-begining_bracket <- begining_bracket%>% 
-  mutate(team_prob = probs,
-         otherteam_prob = 1 - team_prob, 
-         game = paste0(team, " vs. ", otherteam),
-         predicted_winner = if_else(team_prob > 0.5, team, otherteam),
-         predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid))
+add_probs <- function(df, probs){
   
-games <- begining_bracket %>% 
-  mutate(team_prob = as.numeric(probs),
-         otherteam_prob = 1 - team_prob, 
-         game = paste0(team, " vs. ", otherteam)) %>% 
-  select(game, team_prob, otherteam_prob)
+  df <- df %>% 
+    mutate(team_prob = probs,
+           otherteam_prob = 1 - team_prob, 
+           game = paste0(team, " vs. ", otherteam),
+           predicted_winner = if_else(team_prob > 0.5, team, otherteam),
+           predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid),
+           game = paste0(team, " vs. ", otherteam))
+  return(df)
+}
 
-games
+begining_bracket <- add_probs(begining_bracket, probs = probs)
 
+begining_bracket %>% 
+  select(game, team_prob, otherteam_prob, predicted_winner)
 
 # second round ------------------------------------------------------------
 
-second_round <- begining_bracket %>% 
-  select(predicted_winner, predicted_winnerid) %>% 
[…]

### 2021-03-30 R/modeling.R
cleaned up the code and made repeated data manipulations into functions

@@ -5,8 +5,6 @@ source(here::here("R", 'cleaningScript.R'))
 
 # merged <- readr::read_csv(here::here("data", "merged.csv"))
 
-# Take a look -------------------------------------------------------------
-
 
 # model selection ---------------------------------------------------------
 
@@ -31,15 +29,13 @@ fit <- glm(win ~ x3fg +
          opposingbkpg +
                  bkpg,
            data = merged, family = "binomial")
-summary(fit)
+# summary(fit)
 
 s2021 <- team_stats %>% 
   filter(season == 2021, 
          TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
   distinct(TeamID, .keep_all = TRUE) %>% 
   select(-season) 
-
-s2021
 
 predictor_fn <- function (team_1_id, team_2_id) {
   stats_team_1 <- filter(s2021, TeamID == team_1_id)
@@ -50,8 +46,8 @@ predictor_fn <- function (team_1_id, team_2_id) {
   return(pred)
 }
 
-
 # First round -------------------------------------------------------------
+
 begining_bracket
 
 possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")
@@ -76,8 +72,17 @@ add_probs <- function(df, probs){
 
 begining_bracket <- add_probs(begining_bracket, probs = probs)
 
-begining_bracket %>% 
-  select(game, team_prob, otherteam_prob, predicted_winner)
+nice_format <- function(df){
+  output <- df %>%
+    mutate(pred_winner_prob = if_else(team_prob > 0.5, team_prob, otherteam_prob),
+           pred_loser_prob = if_else(team_prob < 0.5, team_prob, otherteam_prob)) %>% 
+    select(game, predicted_winner, 
+           pred_winner_prob, pred_loser_prob)
+  return(output)
+}
+
+
+nice_format(begining_bracket)
 
 # second round ------------------------------------------------------------
 
@@ -91,8 +96,6 @@ advance_round <- function(df){
 }
 
 second_round  <- advance_round(begining_bracket)
-
-second_round
 
 align_teams <- function(df){
   
@@ -114,153 +117,66 @@ probs <- map2_dbl(.x = second_round_games$teamid,
                   .y = second_round_games$otherteamid, 
                   .f = possibly_predictor_fn)
 
-probs
-
 second_round <- add_probs(second_round_games, probs)
 
-second_round
+nice_format(second_round)
 
[…]

### 2021-03-31 R/independent_modeling.R
changed script name and adding script for independent predictions

@@ -1,0 +1,22 @@
+library(tidymodels)
+source(here::here("R", "bracket.R"))
+source(here::here("R", 'cleaningScript.R'))
+theme_set(theme_test())
+options(tibble.print_max = 20, tibble.print_min = 20)
+
+s2021 <- team_stats %>% 
+  filter(season == 2021, 
+         TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
+  distinct(TeamID, .keep_all = TRUE) %>% 
+  select(-season) 
+
+team1 <- pull(s2021, team)
+  
+team2 <- pull(s2021, team) 
+  
+crossing(team1, team2) %>% 
+  distinct() %>%
+  filter(team1 != team2) %>% 
+  # note choose(64,2) = 2016 Im removing double counts.
+  top_n(2016) 
+

### 2021-03-31 R/logistic_dependent_model.R

@@ -1,10 +1,11 @@
 library(tidymodels)
-theme_set(theme_test())
 source(here::here("R", "bracket.R"))
 source(here::here("R", 'cleaningScript.R'))
+theme_set(theme_test())
+options(tibble.print_max = 20, tibble.print_min = 20)
+
 
 # merged <- readr::read_csv(here::here("data", "merged.csv"))
-
 
 # model selection ---------------------------------------------------------
 
@@ -29,7 +30,7 @@ fit <- glm(win ~ x3fg +
          opposingbkpg +
                  bkpg,
            data = merged, family = "binomial")
-# summary(fit)
+summary(fit)
 
 s2021 <- team_stats %>% 
   filter(season == 2021, 
@@ -180,3 +181,12 @@ championship_prediction  <- add_probs(championship_game, championship_probs)
 
 nice_format(championship_prediction)}
 
+# TODO: How do we score this model? train and test data set
+# check previous year 19
+# corss valad
+# caret package
+# weight prediction score pred_winner_prob pred_loser_prob
+# TODO: Run make it independent
+
+# multinomeal category low close high
+# problem with right or wrong...

### 2021-03-31 R/helper_functions.R
put all functions on a file preparing for joining

@@ -1,0 +1,53 @@
+align_teams <- function(df){
+  
+  team <- filter(df, game != "team")
+  other <- filter(df, game == "team")
+  
+  output <- other %>% 
+    rename_all(~ paste0("other",.)) %>% 
+    bind_cols(team) 
+  
+  return(output)
+}
+
+advance_round <- function(df){
+  df <- df %>% 
+    select(predicted_winner, predicted_winnerid) %>% 
+    rename(team = predicted_winner, teamid = predicted_winnerid) %>% 
+    mutate(game = rep(c("team", "otherteam"), nrow(df)/2))
+  
+  return(df)
+}
+
+nice_format <- function(df){
+  output <- df %>%
+    mutate(pred_winner_prob = if_else(team_prob > 0.5, team_prob, otherteam_prob),
+           pred_loser_prob = if_else(team_prob < 0.5, team_prob, otherteam_prob)) %>% 
+    select(game, predicted_winner, 
+           pred_winner_prob, pred_loser_prob)
+  
+  return(output)
+}
+
+add_probs <- function(df, probs){
+  
+  df <- df %>% 
+    mutate(team_prob = probs,
+           otherteam_prob = 1 - team_prob, 
+           game = paste0(team, " vs. ", otherteam),
+           predicted_winner = if_else(team_prob > 0.5, team, otherteam),
+           predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid),
+           game = paste0(team, " vs. ", otherteam))
+  
+  return(df)
+}
+
+predictor_fn <- function (team_1_id, team_2_id) {
+  stats_team_1 <- filter(s2021, TeamID == team_1_id)
+  stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
+    rename_all(~ paste0("opposing",.)) 
+  bound_cols <- bind_cols(stats_team_1, stats_team_2)
+  pred <- predict(fit, newdata = bound_cols, type = "response")
+  
+  return(pred)
+}

### 2021-03-31 R/independent_modeling.R

@@ -1,6 +1,7 @@
 library(tidymodels)
 source(here::here("R", "bracket.R"))
 source(here::here("R", 'cleaningScript.R'))
+source(here::here("R", "helper_functions.R"))
 theme_set(theme_test())
 options(tibble.print_max = 20, tibble.print_min = 20)
 
@@ -11,12 +12,12 @@ s2021 <- team_stats %>%
   select(-season) 
 
 team1 <- pull(s2021, team)
-  
 team2 <- pull(s2021, team) 
   
 crossing(team1, team2) %>% 
   distinct() %>%
   filter(team1 != team2) %>% 
   # note choose(64,2) = 2016 Im removing double counts.
-  top_n(2016) 
+  top_n(2016) %>% 
+  left_join(team_stats, by = team)
 

### 2021-03-31 R/logistic_dependent_model.R

@@ -1,9 +1,9 @@
 library(tidymodels)
 source(here::here("R", "bracket.R"))
 source(here::here("R", 'cleaningScript.R'))
+source(here::here("R", "helper_functions.R"))
 theme_set(theme_test())
-options(tibble.print_max = 20, tibble.print_min = 20)
-
+options(tibble.print_min = 38)
 
 # merged <- readr::read_csv(here::here("data", "merged.csv"))
 
@@ -38,14 +38,6 @@ s2021 <- team_stats %>%
   distinct(TeamID, .keep_all = TRUE) %>% 
   select(-season) 
 
-predictor_fn <- function (team_1_id, team_2_id) {
-  stats_team_1 <- filter(s2021, TeamID == team_1_id)
-  stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
-  rename_all(~ paste0("opposing",.)) 
-  bound_cols <- bind_cols(stats_team_1, stats_team_2)
-  pred <- predict(fit, newdata = bound_cols, type = "response")
-  return(pred)
-}
 
 # First round -------------------------------------------------------------
 
@@ -57,62 +49,16 @@ probs <- map2_dbl(.x = begining_bracket$teamid,
                   .y = begining_bracket$otherteamid, 
                   .f = possibly_predictor_fn)
 
-probs
-
-add_probs <- function(df, probs){
-  
-  df <- df %>% 
-    mutate(team_prob = probs,
-           otherteam_prob = 1 - team_prob, 
-           game = paste0(team, " vs. ", otherteam),
-           predicted_winner = if_else(team_prob > 0.5, team, otherteam),
-           predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid),
-           game = paste0(team, " vs. ", otherteam))
-  return(df)
-}
 
 begining_bracket <- add_probs(begining_bracket, probs = probs)
-
-nice_format <- function(df){
-  output <- df %>%
-    mutate(pred_winner_prob = if_else(team_prob > 0.5, team_prob, otherteam_prob),
-           pred_loser_prob = if_else(team_prob < 0.5, team_prob, otherteam_prob)) %>% 
-    select(game, predicted_winner, 
-           pred_winner_prob, pred_loser_prob)
-  return(output)
-}
-
 
 nice_format(begining_bracket)
 
 # second round ------------------------------------------------------------
 
-advance_round <- function(df){
-  df <- df %>% 
-    select(predicted_winner, predicted_winnerid) %>% 
[…]

### 2021-03-31 R/helper_functions.R
ran independent models in one DF

@@ -51,3 +51,5 @@ predictor_fn <- function (team_1_id, team_2_id) {
   
   return(pred)
 }
+
+possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")

### 2021-03-31 R/independent_modeling.R

@@ -14,10 +14,28 @@ s2021 <- team_stats %>%
 team1 <- pull(s2021, team)
 team2 <- pull(s2021, team) 
   
-crossing(team1, team2) %>% 
+teams <- crossing(team = team1, other = team2) %>% 
   distinct() %>%
-  filter(team1 != team2) %>% 
+  filter(team != other) %>% 
   # note choose(64,2) = 2016 Im removing double counts.
-  top_n(2016) %>% 
-  left_join(team_stats, by = team)
+  top_n(2016)
 
+
+team1id <- pull(s2021, TeamID)
+team2id <- pull(s2021, TeamID) 
+
+start <- system.time()
+ids <-  crossing(team = team1id, other = team2id) %>% 
+        distinct() %>%
+        filter(team != other) %>% 
+        # note choose(64,2) = 2016 Im removing double counts.
+        top_n(2016) %>% 
+        mutate(predicted_probs = map2_dbl(.x = ids$team, 
+                                          .y = ids$other, 
+                                          .f = possibly_predictor_fn))
+
+system.time() - start
+ids
+
+# TODO: join ids into a df.
+# TODO: rename possibly_predictor_fn

### 2021-03-31 R/logistic_dependent_model.R

@@ -43,8 +43,6 @@ s2021 <- team_stats %>%
 
 begining_bracket
 
-possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")
-
 probs <- map2_dbl(.x = begining_bracket$teamid, 
                   .y = begining_bracket$otherteamid, 
                   .f = possibly_predictor_fn)

### 2021-03-31 R/bracket.R
preped data for independent models

@@ -66,8 +66,6 @@ bracket
 # begining bracket (no pockets) -------------------------------------------
 
 
-
-
 begining_bracket <- begining_bracket %>% 
   select(2, 35) %>% 
   stack() %>% 

### 2021-03-31 R/cleaningScript.R

@@ -104,3 +104,11 @@ merged <- readr::read_csv(here::here("data", "merged.csv"))
 
 merged %>% 
   skimr::skim()
+
+
+
+s2021 <- team_stats %>% 
+  filter(season == 2021, 
+         TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
+  distinct(TeamID, .keep_all = TRUE) %>% 
+  select(-season)

### 2021-03-31 R/helper_functions.R

@@ -42,7 +42,31 @@ add_probs <- function(df, probs){
   return(df)
 }
 
-predictor_fn <- function (team_1_id, team_2_id) {
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+
+fit <- glm(win ~ x3fg +
+             opposingx3fg +
+             # field goal pct
+             fg_percent +
+             opposingfg_percent +
+             # free throws
+             ft_percent +
+             opposingft_percent +
+             # rebound per game
+             rpg + 
+             opposingrpg +
+             # steels
+             st +
+             opposingst +
+             #turnover
+             to +
+             opposingto +
+             # blocks
+             opposingbkpg +
+             bkpg,
+             data = merged, family = "binomial")
+
+logistic_predictor_fn <- function (team_1_id, team_2_id) {
   stats_team_1 <- filter(s2021, TeamID == team_1_id)
   stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
     rename_all(~ paste0("opposing",.)) 
@@ -52,4 +76,4 @@ predictor_fn <- function (team_1_id, team_2_id) {
   return(pred)
 }
 
-possibly_predictor_fn <- possibly(.f = predictor_fn, otherwise = "ERROR")
+logistic_predictor <- possibly(.f = logistic_predictor_fn, otherwise = "ERROR")

### 2021-03-31 R/independent_modeling.R

@@ -1,41 +1,74 @@
 library(tidymodels)
 source(here::here("R", "bracket.R"))
-source(here::here("R", 'cleaningScript.R'))
 source(here::here("R", "helper_functions.R"))
+source(here::here("R", 'cleaningScript.R'))
 theme_set(theme_test())
-options(tibble.print_max = 20, tibble.print_min = 20)
-
-s2021 <- team_stats %>% 
-  filter(season == 2021, 
-         TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
-  distinct(TeamID, .keep_all = TRUE) %>% 
-  select(-season) 
-
-team1 <- pull(s2021, team)
-team2 <- pull(s2021, team) 
-  
-teams <- crossing(team = team1, other = team2) %>% 
-  distinct() %>%
-  filter(team != other) %>% 
-  # note choose(64,2) = 2016 Im removing double counts.
-  top_n(2016)
-
+options(tibble.print_min = 20)
 
 team1id <- pull(s2021, TeamID)
 team2id <- pull(s2021, TeamID) 
 
-start <- system.time()
-ids <-  crossing(team = team1id, other = team2id) %>% 
-        distinct() %>%
-        filter(team != other) %>% 
-        # note choose(64,2) = 2016 Im removing double counts.
-        top_n(2016) %>% 
-        mutate(predicted_probs = map2_dbl(.x = ids$team, 
-                                          .y = ids$other, 
-                                          .f = possibly_predictor_fn))
+ids <- crossing(teamid = team1id, other = team2id) %>% 
+  distinct() %>%
+  filter(teamid != other) %>% 
+  # note choose(64,2) = 2016 Im removing double counts.
+  top_n(2016) %>% 
+  left_join(s2021, by = c("other" = "TeamID")) %>% 
+  rename(other_team = team) %>% 
+  select(teamid:other_team) %>% 
+  left_join(s2021, by = c("teamid" = "TeamID")) %>% 
+  mutate(game = paste0(team, " vs. ", other_team)) %>% 
+  select(team, other_team,game, teamid, otherid = other)
+  
+start <- Sys.time()
+predicted_games <- ids %>% 
+  mutate(predicted_probs = map2_dbl(.x = ids$teamid, 
+                                    .y = ids$otherid, 
+                                    .f = logistic_predictor))
+# about 30 seconds to run
+paste0("runtime: ", round(Sys.time() - start, 2))
 
-system.time() - start
-ids
+predicted_games %>% 
+  mutate(  
[…]

### 2021-03-31 R/logistic_dependent_model.R

@@ -1,4 +1,3 @@
-library(tidymodels)
 source(here::here("R", "bracket.R"))
 source(here::here("R", 'cleaningScript.R'))
 source(here::here("R", "helper_functions.R"))
@@ -45,7 +44,7 @@ begining_bracket
 
 probs <- map2_dbl(.x = begining_bracket$teamid, 
                   .y = begining_bracket$otherteamid, 
-                  .f = possibly_predictor_fn)
+                  .f = logistic_predictor)
 
 
 begining_bracket <- add_probs(begining_bracket, probs = probs)
@@ -60,7 +59,7 @@ second_round_games <- align_teams(second_round)
 
 probs <- map2_dbl(.x = second_round_games$teamid, 
                   .y = second_round_games$otherteamid, 
-                  .f = possibly_predictor_fn)
+                  .f = logistic_predictor)
 
 second_round <- add_probs(second_round_games, probs)
 
@@ -74,7 +73,7 @@ sweet_sixteen_games <- align_teams(sweet_sixteen)
 
 sweet_sixteen_probs <- map2_dbl(.x = sweet_sixteen_games$teamid, 
                                 .y = sweet_sixteen_games$otherteamid, 
-                                .f = possibly_predictor_fn)
+                                .f = logistic_predictor)
 
 sweet_sixteen_prediction <- add_probs(sweet_sixteen_games, probs = sweet_sixteen_probs)
 
@@ -89,7 +88,7 @@ elite_eight_games <- align_teams(elite_eight)
 
 elite_eight_probs <- map2_dbl(.x = elite_eight_games$teamid, 
                                 .y = elite_eight_games$otherteamid, 
-                                .f = possibly_predictor_fn)
+                                .f = logistic_predictor)
 
 elite_eight_prediction <- add_probs(elite_eight_games, elite_eight_probs)
 
@@ -104,7 +103,7 @@ final_four_games <- align_teams(final_four)
 
 final_four_probs <- map2_dbl(.x = final_four_games$teamid, 
                              .y = final_four_games$otherteamid, 
-                             .f = possibly_predictor_fn)
+                             .f = logistic_predictor)
 
 final_four_prediction <- add_probs(final_four_games, final_four_probs)
 
@@ -119,7 +118,7 @@ championship_game <- align_teams(championship)
 
 championship_probs <- map2_dbl(.x = championship_game$teamid, 
                                .y = championship_game$otherteamid, 
-                               .f = possibly_predictor_fn)
[…]

### 2021-04-01 R/bracket.R
cleaned up code and madsure the source calls were independent

@@ -4,9 +4,7 @@ library(tidyverse)
 
 team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
 bracket <- readxl::read_excel(here::here("data", "prebracket.xlsx"))
-
 begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
-
 
 # Even better
 # bracket <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1zaWZ2Xh7sll-PZKA1DhKvS6TdfE0Yggk0m5OWY4aICc/edit#gid=0")
@@ -64,7 +62,8 @@ bracket
 
 
 # begining bracket (no pockets) -------------------------------------------
-
+begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
+team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
 
 begining_bracket <- begining_bracket %>% 
   select(2, 35) %>% 

### 2021-04-01 R/cleaningScript.R

@@ -1,5 +1,5 @@
 library(tidyverse) 
-options(tibble.print_max = 35, tibble.print_min = 35)
+# options(tibble.print_min = 35)
 
 list_of_dfs <- sapply(paste0("rawdata/", dir("rawdata")), read_csv, USE.NAMES = TRUE)
 
@@ -106,9 +106,66 @@ merged %>%
   skimr::skim()
 
 
+# bracket no pockets ------------------------------------------------------
+
+begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
+team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
+
+begining_bracket <- begining_bracket %>% 
+  select(2, 35) %>% 
+  stack() %>% 
+  drop_na() %>% 
+  tibble() %>% 
+  rename(team = values) %>% 
+  mutate(team = tolower(team)) %>% 
+  mutate(game = 1:n(),
+         game = as.numeric(game),
+         otherteam  = if_else(game %% 2 == 1, team, lag(team)),
+         game = if_else(game %% 2 == 1, game + 1, game), .before = team) %>% 
+  filter(team != otherteam) %>% 
+  filter(otherteam != "first round") %>% 
+  left_join(team_names, by = c("team" = "TeamNameSpelling")) %>% 
+  rename(teamid = TeamID) %>% 
+  left_join(team_names, by = c("otherteam" = "TeamNameSpelling")) %>%
+  rename(otherteamid = TeamID) %>% 
+  mutate(teamid = replace(teamid, team == "app st.", 1111)) %>% 
+  mutate_if(is.numeric, as.character) %>% 
+  select(-ind, -game)
+
+begining_bracket
+
+# 2021 stats --------------------------------------------------------------
+
+
 
 s2021 <- team_stats %>% 
   filter(season == 2021, 
          TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
   distinct(TeamID, .keep_all = TRUE) %>% 
   select(-season)
+
+
+# logsitic fit ------------------------------------------------------------
+
+logistic_fit <- glm(win ~ x3fg +
+                      opposingx3fg +
+                      # field goal pct
+                      fg_percent +
+                      opposingfg_percent +
+                      # free throws
+                      ft_percent +
+                      opposingft_percent +
+                      # rebound per game
+                      rpg + 
+                      opposingrpg +
+                      # steels
+                      st +
+                      opposingst +
+                      #turnover
[…]

### 2021-04-01 R/helper_functions.R

@@ -42,38 +42,14 @@ add_probs <- function(df, probs){
   return(df)
 }
 
-merged <- readr::read_csv(here::here("data", "merged.csv"))
-
-fit <- glm(win ~ x3fg +
-             opposingx3fg +
-             # field goal pct
-             fg_percent +
-             opposingfg_percent +
-             # free throws
-             ft_percent +
-             opposingft_percent +
-             # rebound per game
-             rpg + 
-             opposingrpg +
-             # steels
-             st +
-             opposingst +
-             #turnover
-             to +
-             opposingto +
-             # blocks
-             opposingbkpg +
-             bkpg,
-             data = merged, family = "binomial")
-
 logistic_predictor_fn <- function (team_1_id, team_2_id) {
   stats_team_1 <- filter(s2021, TeamID == team_1_id)
   stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
-    rename_all(~ paste0("opposing",.)) 
+                         rename_all(~ paste0("opposing",.)) 
   bound_cols <- bind_cols(stats_team_1, stats_team_2)
-  pred <- predict(fit, newdata = bound_cols, type = "response")
+  pred <- predict(logistic_fit, newdata = bound_cols, type = "response")
   
   return(pred)
 }
 
-logistic_predictor <- possibly(.f = logistic_predictor_fn, otherwise = "ERROR")
+logistic_predictor <- purrr::possibly(.f = logistic_predictor_fn, otherwise = "ERROR")

### 2021-04-01 R/independent_modeling.R

@@ -1,7 +1,6 @@
-library(tidymodels)
-source(here::here("R", "bracket.R"))
+source(here::here("R", "cleaningScript.R"))
 source(here::here("R", "helper_functions.R"))
-source(here::here("R", 'cleaningScript.R'))
+library(tidymodels)
 theme_set(theme_test())
 options(tibble.print_min = 20)
 
@@ -11,8 +10,7 @@ team2id <- pull(s2021, TeamID)
 ids <- crossing(teamid = team1id, other = team2id) %>% 
   distinct() %>%
   filter(teamid != other) %>% 
-  # note choose(64,2) = 2016 Im removing double counts.
-  top_n(2016) %>% 
+  top_n(2016) %>%   # choose(64,2) = 2016 (removing double counts)
   left_join(s2021, by = c("other" = "TeamID")) %>% 
   rename(other_team = team) %>% 
   select(teamid:other_team) %>% 
@@ -25,20 +23,21 @@ predicted_games <- ids %>%
   mutate(predicted_probs = map2_dbl(.x = ids$teamid, 
                                     .y = ids$otherid, 
                                     .f = logistic_predictor))
-# about 30 seconds to run
+# This takes about 30 seconds to run
 paste0("runtime: ", round(Sys.time() - start, 2))
+
 
 predicted_games %>% 
   mutate(  
     predicted_winner = if_else(predicted_probs > .5, team, other_team),
     predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
     predicted_loser_probs = 1 - predicted_winner_probs,
+    prediction_points = predicted_winner_probs - predicted_loser_probs,
     .after = game
   )
 
-# TODO: rename possibly_predictor_fn
 
-# 
+
 # fit <- glm(win ~ x3fg +
 #              opposingx3fg +
 #              # field goal pct
@@ -59,7 +58,7 @@ predicted_games %>%
 #              # blocks
 #              opposingbkpg +
 #              bkpg,
-#            data = merged, family = "binomial")
+#              data = merged, family = "binomial")
 # 
 # logistic_predictor_fn <- function (team_1_id, team_2_id) {
 #   stats_team_1 <- filter(s2021, TeamID == team_1_id)

### 2021-04-01 R/logistic_dependent_model.R

@@ -1,4 +1,3 @@
-source(here::here("R", "bracket.R"))
 source(here::here("R", 'cleaningScript.R'))
 source(here::here("R", "helper_functions.R"))
 theme_set(theme_test())
@@ -29,14 +28,6 @@ fit <- glm(win ~ x3fg +
          opposingbkpg +
                  bkpg,
            data = merged, family = "binomial")
-summary(fit)
-
-s2021 <- team_stats %>% 
-  filter(season == 2021, 
-         TeamID %in% c(begining_bracket$teamid, begining_bracket$otherteamid)) %>% 
-  distinct(TeamID, .keep_all = TRUE) %>% 
-  select(-season) 
-
 
 # First round -------------------------------------------------------------
 

### 2021-04-01 R/cleaningScript.R
ade adj for poisson model

@@ -98,6 +98,7 @@ statoutcome %>%
   rename(ConfAbbrev = ConfAbbrev.x) %>% 
   janitor::clean_names() %>% 
   distinct() %>% 
+  mutate(team_score = ifelse(win == 1, w_score, l_score)) %>% 
   readr::write_csv(here::here("data", "merged.csv"))
 
 merged <- readr::read_csv(here::here("data", "merged.csv"))

### 2021-04-01 R/cleaningScript.R
added poison and mulitnom, with categories and points scored

@@ -98,7 +98,20 @@ statoutcome %>%
   rename(ConfAbbrev = ConfAbbrev.x) %>% 
   janitor::clean_names() %>% 
   distinct() %>% 
-  mutate(team_score = ifelse(win == 1, w_score, l_score)) %>% 
+  mutate(team_score = ifelse(win == 1, w_score, l_score),
+         other_team_score = ifelse(w_score == team_score, l_score, w_score),
+         diff = team_score - other_team_score, 
+         fifth_xtile = case_when(diff <  -12             ~ '<-12',
+                                 diff >= -12 & diff < -4 ~ '-12:-4',
+                                 diff >=  -4 & diff <  4 ~ '-4:4',
+                                 diff >=   4 & diff < 13 ~ '4:13',
+                                 diff >= 13               ~ '>13'
+         )) %>% 
+  mutate(x_tile = fct_relevel(fifth_xtile, levels = c("<-12",
+                                                      "-12:-4",
+                                                      "-4:4",
+                                                      "4:13",
+                                                      ">13"))) %>% 
   readr::write_csv(here::here("data", "merged.csv"))
 
 merged <- readr::read_csv(here::here("data", "merged.csv"))

### 2021-04-01 R/independent_modeling.R

@@ -1,4 +1,5 @@
 source(here::here("R", "cleaningScript.R"))
+# TODO: write all script files to csv and read in
 source(here::here("R", "helper_functions.R"))
 library(tidymodels)
 theme_set(theme_test())
@@ -18,6 +19,7 @@ ids <- crossing(teamid = team1id, other = team2id) %>%
   mutate(game = paste0(team, " vs. ", other_team)) %>% 
   select(team, other_team,game, teamid, otherid = other)
   
+
 start <- Sys.time()
 predicted_games <- ids %>% 
   mutate(predicted_probs = map2_dbl(.x = ids$teamid, 
@@ -37,37 +39,124 @@ predicted_games %>%
   )
 
 
+# poisson -----------------------------------------------------------------
 
-# fit <- glm(win ~ x3fg +
-#              opposingx3fg +
-#              # field goal pct
-#              fg_percent +
-#              opposingfg_percent +
-#              # free throws
-#              ft_percent +
-#              opposingft_percent +
-#              # rebound per game
-#              rpg + 
-#              opposingrpg +
-#              # steels
-#              st +
-#              opposingst +
-#              #turnover
-#              to +
-#              opposingto +
-#              # blocks
-#              opposingbkpg +
-#              bkpg,
-#              data = merged, family = "binomial")
+
+fit_poisson <- glm(team_score ~ x3fg +
+             opposingx3fg +
+             # field goal pct
+             fg_percent +
+             opposingfg_percent +
+             # free throws
+             ft_percent +
+             opposingft_percent +
+             # rebound per game
+             rpg +
+             opposingrpg +
+             # steels
+             st +
+             opposingst +
+             #turnover
+             to +
+             opposingto +
+             # blocks
+             opposingbkpg +
+             bkpg,
+             data = merged, family = poisson(link = "log"))
+
+summary(fit_poisson)
+
+
+poisson_predictor_fn <- function (team_1_id, team_2_id) {
+  stats_team_1 <- filter(s2021, TeamID == team_1_id)
+  stats_team_2 <- filter(s2021, TeamID == team_2_id) %>%
+    rename_all(~ paste0("opposing",.))
+  bound_cols <- bind_cols(stats_team_1, stats_team_2)
+  pred <- predict(fit_poisson, newdata = bound_cols, type = "response")
+
+  return(pred)
+}
 # 
[…]

### 2021-04-01 R/independent_modeling.R
put note to step_norm...

@@ -5,6 +5,8 @@ library(tidymodels)
 theme_set(theme_test())
 options(tibble.print_min = 20)
 
+# TODO: recipes::step_normalize() try that see if there is a difference
+
 team1id <- pull(s2021, TeamID)
 team2id <- pull(s2021, TeamID) 
 

### 2021-04-02 R/cleaningScript.R
cashed larger computations

@@ -56,7 +56,7 @@ nameandID <- list_of_dfs$`rawdata/MTeamSpellings.csv` %>%
   filter(Season == 2019) %>% 
   select(-Season) 
 
-team_stats <- sapply(stats, selector) %>% 
+sapply(stats, selector) %>% 
   bind_rows(.id = "id") %>% 
   select(-contains("opp"), -`REB MAR`, -ORebs, -DRebs, -TOPG) %>% 
   mutate(season = str_extract(id,"([0-9]+).*$"), .before = id) %>% 
@@ -68,7 +68,11 @@ team_stats <- sapply(stats, selector) %>%
   # Drops 8 teams, all with W-L ratios < .5
   drop_na() %>% 
   mutate(TeamID = as.character(TeamID)) %>% 
-  tibble()
+  tibble() %>% 
+  write_csv(here::here("data", "team_stats.csv"))
+
+team_stats <- read_csv(here::here("team_stats.csv"))
+
 # Here is complete df of team stats from 2019 to 2021
 
 team_stats
@@ -116,16 +120,12 @@ statoutcome %>%
 
 merged <- readr::read_csv(here::here("data", "merged.csv"))
 
-merged %>% 
-  skimr::skim()
-
-
 # bracket no pockets ------------------------------------------------------
 
 begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
 team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
 
-begining_bracket <- begining_bracket %>% 
+begining_bracket %>% 
   select(2, 35) %>% 
   stack() %>% 
   drop_na() %>% 
@@ -144,42 +144,37 @@ begining_bracket <- begining_bracket %>%
   rename(otherteamid = TeamID) %>% 
   mutate(teamid = replace(teamid, team == "app st.", 1111)) %>% 
   mutate_if(is.numeric, as.character) %>% 
-  select(-ind, -game)
+  select(-ind, -game) %>% 
+  write_csv(here::here("data", "begining_bracket.csv"))
 
-begining_bracket
+begining_bracket <- read_csv(here::here("data", "begining_bracket.csv"))
 
 # 2021 stats --------------------------------------------------------------
 
-
-
-s2021 <- team_stats %>% 
+team_stats %>% 
   filter(season == 2021, 
[…]

### 2021-04-02 R/independent_modeling.R

@@ -1,37 +1,51 @@
-source(here::here("R", "cleaningScript.R"))
-# TODO: write all script files to csv and read in
-source(here::here("R", "helper_functions.R"))
+# source(here::here("R", "cleaningScript.R"))
+
 library(tidymodels)
-theme_set(theme_test())
+source(here::here("R", "helper_functions.R"))
+ids <- readr::read_csv(here::here("data", "ids.csv"))
+s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+team_stats <- readr::read_csv(here::here("data", "team_stats.csv"))
+begining_bracket <- readr::read_csv(here::here("data", "begining_bracket.csv"))
 options(tibble.print_min = 20)
 
 # TODO: recipes::step_normalize() try that see if there is a difference
 
-team1id <- pull(s2021, TeamID)
-team2id <- pull(s2021, TeamID) 
+# logsitic fit ------------------------------------------------------------
 
-ids <- crossing(teamid = team1id, other = team2id) %>% 
-  distinct() %>%
-  filter(teamid != other) %>% 
-  top_n(2016) %>%   # choose(64,2) = 2016 (removing double counts)
-  left_join(s2021, by = c("other" = "TeamID")) %>% 
-  rename(other_team = team) %>% 
-  select(teamid:other_team) %>% 
-  left_join(s2021, by = c("teamid" = "TeamID")) %>% 
-  mutate(game = paste0(team, " vs. ", other_team)) %>% 
-  select(team, other_team,game, teamid, otherid = other)
-  
+# logistic_fit <- glm(win ~ x3fg +
+#                       opposingx3fg +
+#                       # field goal pct
+#                       fg_percent +
+#                       opposingfg_percent +
+#                       # free throws
+#                       ft_percent +
+#                       opposingft_percent +
+#                       # rebound per game
+#                       rpg +
+#                       opposingrpg +
+#                       # steels
+#                       st +
+#                       opposingst +
+#                       #turnover
+#                       to +
+#                       opposingto +
+#                       # blocks
+#                       opposingbkpg +
+#                       bkpg,
+#                     data = merged, family = "binomial")
+# 
+# start <- Sys.time()
+# logistic_predictions <- ids %>%
+#   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
[…]

### 2021-04-02 R/logistic_dependent_model.R
rm todos

@@ -114,13 +114,3 @@ championship_probs <- map2_dbl(.x = championship_game$teamid,
 championship_prediction  <- add_probs(championship_game, championship_probs)
 
 nice_format(championship_prediction)
-
-# TODO: How do we score this model? train and test data set
-# check previous year 19
-# corss valad
-# caret package
-# weight prediction score pred_winner_prob pred_loser_prob
-# TODO: Run make it independent
-
-# multinomeal category low close high
-# problem with right or wrong...

### 2021-04-02 paper/paper.Rmd
addining paper

---
title             : "Using GLMs to Predict Basketball Games"
shorttitle        : "GLM"

author: 
  - name          : "Sabrina Ball"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
  - name          : "Joseph Despres"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Michigan State University \n "

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

__Introduction__

testing citat [@R-base]
<!-- (a) introduces the background -->
<!-- (b) states the scientific questions you aim to answer -->
<!-- (c) describes the data set that you will use for the analysis (d) summarizes your major findings. -->
__Exploratory Data Analysis__  
<!-- • Perform exploratory data analysis (with summary statistics or exploratory graphs) that reveals some basic and general information of the data set. -->
__Description__  
<!-- • Provide a description and explanation of the steps that you have taken to obtain the best model (or models). -->
__Results__  
<!-- • Give the results of relevant parameter estimation and statistical inference under your best model (or models). Discuss and interpret important features of your best model (or models), and state your conclusions in the context of the scientific questions that you have been focused on. -->
__Diagnostic Methods__  
<!-- • Check the goodness of fit of your model (or models), including a description of the model diagnostic methods and test procedures you used, and what the results were. -->
__Conculsion__  
<!-- • Conclude the report with a discussion section that -->
<!-- (a) recaps your scientific findings -->
<!-- (b) discusses possible limitations and improvements of your methods -->
<!-- (c) outlines additional questions that are worth studying as a future work. -->

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2021-04-02 paper/paper.tex
addining paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Using GLMs to Predict Basketball Games},
  pdfauthor={Sabrina Ball1 \& Joseph Despres1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{GLM}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Using GLMs to Predict Basketball Games}
\author{Sabrina Ball\textsuperscript{1} \& Joseph Despres\textsuperscript{1}}
\date{}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Michigan State University}

\begin{document}
\maketitle

\textbf{Introduction}

testing citat (R Core Team, 2020)
\textbf{Exploratory Data Analysis}\\
\textbf{Description}\\
\textbf{Results}\\
\textbf{Diagnostic Methods}\\
\textbf{Conculsion}\\

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\end{CSLReferences}

\endgroup

\end{document}

### 2021-04-05 proposal/proposal.Rmd
finished proposal

@@
 biblio-style: apsr
-title: "Using Statistical Methods to Predict the Outcome of Basketball Games"
+title: "Predicting the Outcome of Basketball Games with GLMs"
 author:
-- name: Sabrina Ball
+- name: Sabrina Ball and Joseph Despres
   affiliation:
-- name: Joseph Despres
-  affiliation:
-abstract: "Quick Summary"
+abstract: "Every year top university basketball teams compete in the March Madness tournament.
+Casual fans and Enthuasists submit predictions gambling small sums of money on the outcomes.
+This study will determine if which GLM performs best in predicting the results."
 geometry: margin=1in
@@
 
-# 1. Data
- ^[https://plexkits.com/march-madness-bracket/].
+# Data
+Our data comes from three different sources Kaggle, NCAA, and the tournament results.
+Kagggle provides a comprehensive dataset^[https://plexkits.com/march-madness-bracket/] for all NCAA in season (non-turniment) basektball games from 2001 to 2020.
+The NCAA provides team-level statistics^[https://stats.ncaa.org/rankings/change_sport_year_div] for each team.
+We will use those two datasets to train a model that uses team-level statistics to predict basketball game outcomes.
+We will test our model on the actual tournament results after it is concluded on April 5th.
+The objective is to predict the individual game outcomes as accurately as possible.
 
-## Variables
+Our response is the outcome *win/loss* as a function of team statistics.
+Win or loss coded 0 or 1.
+Then we will look at both team's statistics and determine which is more important.
+In our dataset we have: *field goal percentage* which is the rate attempt to score vs actually scoring.
+*Free-throw percentage*,  is the rate that the team has the opportunity to take an unguarded shot.
+Cumulative *Three-point goals made*.
+*Rebounds per game*, counts the times the team recovered the ball after a missed shot.
+*Steals*, the times a team was able to remove possession of the ball from the opposing team.
+*Turnover*, the number of times the team lost possession.
+*Blocks*, The number of times the team was able to block a shot made by the opposing team.
 
-Our response is the outcome win/loss as a function of team statistics. Win or loss coded 0 or 1. field goal precentage, 
-win,fg_percent, opposingfg_percent, ft_percent, opposingft_percent, rpg, opposingrpg, st, opposingst, to, opposingto, opposingbkpg, bkpg
+# Learning Objectives
 
-# 2. Questions
+We want to learn how to make and validate predictions with GLMs.
+Additionally, we want an appropriate scoring metric that can compare the performance of various methods.
+We want to know which GLM performs the best for this particular problem.
 
-Can we use statistical methods to predict outcomes of the March Madness basketball games better than chance? Better than seed? Better than betting markets?
+# Methods and Difficulties
 
-# 3. Learning
+We will use three regression methods, logistic, Poisson, and multinominal, to compare the results to determine which is most accurate.
+Logistic regression most naturally fits this problem because games do not tie so there is a 50/50 split of wins and losses.
+Also, we will try Poisson regression to predict the number of points scored by each team.
+Since there are many close basketball games we will use multinomial regression to predict if the game score is going to be a large difference one way, small difference, or a large difference the other way.
 
-We would like to learn how to make predictions with logistic regression.
-We would like to come up with a good way to evaluate our predictions.
-This will be dificult since the outcome is binary so we would need to weight our predictions somehow and index our socres.
-
-# 4. Methods
-
-Logistic regression with a response prediction function.
-
-# 5. Issues
-
-The seeds are likely derrived from the predictors we are using. Multicolinearity, will be something to watchout for because better teams probally have better statistics.
+For the most part, assembling the data was not an issue yet I was worried about that.
+An issue we will face is scoring this model.
+We switched from making predictions depending on the previous round to being independent of the previous round.
+We come up with a simple scoring metric then, treat games played as a random sample from all possible games, and use basic statistical methods to infer the prediction accuracy.

### 2021-04-05 R/cleaningScript.R
fixed a bug, where reader converted id to numeric and messed up joins

@@ -1,6 +1,5 @@
 library(tidyverse) 
 # options(tibble.print_min = 35)
-
 list_of_dfs <- sapply(paste0("rawdata/", dir("rawdata")), read_csv, USE.NAMES = TRUE)
 
 games <- list_of_dfs$`rawdata/MConferenceTourneyGames.csv` %>% 
@@ -67,17 +66,21 @@ sapply(stats, selector) %>%
   left_join(nameandID, by = c("team" = "sp")) %>% 
   # Drops 8 teams, all with W-L ratios < .5
   drop_na() %>% 
-  mutate(TeamID = as.character(TeamID)) %>% 
   tibble() %>% 
-  write_csv(here::here("data", "team_stats.csv"))
+  write_csv(here::here("data", "team_stats.csv")) 
 
-team_stats <- read_csv(here::here("team_stats.csv"))
+
+team_stats <- read_csv(here::here("data", "team_stats.csv")) %>% 
+    mutate(TeamID = as.character(TeamID),
+           season = as.character(season))
 
 # Here is complete df of team stats from 2019 to 2021
 
-team_stats
+team_stats 
 
 statoutcome <- seasionsoutcomes %>% 
+  mutate(TeamID = as.character(TeamID),
+         season = as.character(season)) %>% 
   # this is where we lose 2021
   left_join(team_stats, by = c('TeamID', 'season')) %>% 
   drop_na() %>% 

### 2021-04-05 paper/paper.Rmd
update paper

@@
 title             : "Using GLMs to Predict Basketball Games"
-shorttitle        : "GLM"
+shorttitle        : "Using GLMs to Predict Basketball Games"
 
 author: 
+  - name          : "Joe Despres"
+    corresponding : yes    # Define only one corresponding author
   - name          : "Sabrina Ball"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-  - name          : "Joseph Despres"
-    affiliation   : "1"
 
 affiliation:
-  - id            : "1"
-    institution   : "Michigan State University \n "
+  - id            : ""
+    institution   : "Michigan State University"
 
-keywords          : "keywords"
+keywords          : "keyword"
 wordcount         : "X"
@@
 
-floatsintext      : no
+floatsintext      : yes
 figurelist        : no
@@
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : no
@@
 output            : papaja::apa6_pdf
+urlcolor: blue
 ---
@@
 library("papaja")
+library(tidymodels)
+library(patchwork)
+library(kableExtra)
+theme_set(theme_test())
 r_refs("r-references.bib")
+
+merged <- readr::read_csv("data/merged.csv")
+team_stats <- readr::read_csv("data/team_stats.csv")
 ```
@@
 
-testing citat [@R-base]
+Every year top university basketball teams compete in the March Madness tournament.
+Casual fans and enthusiasts submit predictions gambling small sums of money on the outcomes.
+This study will determine if which GLM performs best in predicting the results.
+Our data comes from three different sources Kaggle, NCAA, and the tournament results.
+Kagggle provides a comprehensive dataset^[CITE THIS: https://plexkits.com/march-madness-bracket/] for all NCAA in season (non-turniment) basektball games from 2001 to 2020.
+The NCAA provides team-level statistics^[CITE THIS: https://stats.ncaa.org/rankings/change_sport_year_div] for each team.
+We will use those two datasets to train a model that uses team-level statistics to predict basketball game outcomes.
+We will test our model on the actual tournament results after it is concluded on April 5th.
+The objective is to predict the individual game outcomes as accurately as possible.
+
+Our response is the outcome *win/loss* as a function of team statistics.
+Win or loss coded 0 or 1.
+Then we will look at both team's statistics and determine which is more important.
+In our dataset we have: *field goal percentage* which is the rate attempt to score vs actually scoring.
+*Free-throw percentage*,  is the rate that the team has the opportunity to take an unguarded shot.
+Cumulative *Three-point goals made*.
+*Rebounds per game*, counts the times the team recovered the ball after a missed shot.
+*Steals*, the times a team was able to remove possession of the ball from the opposing team.
+*Turnover*, the number of times the team lost possession.
+*Blocks*, The number of times the team was able to block a shot made by the opposing team.
+
+We will use three regression methods, logistic, Poisson, and multinominal, to compare the results to determine which is most accurate.
+Logistic regression most naturally fits this problem because games do not tie so there is a 50/50 split of wins and losses.
+Also, we will try Poisson regression to predict the number of points scored by each team.
+Since there are many close basketball games we will use multinomial regression to predict if the game score is going to be a large difference one way, small difference, or a large difference the other way.
+
+We Found....
+
 <!-- (a) introduces the background -->
@@
 <!-- (c) describes the data set that you will use for the analysis (d) summarizes your major findings. -->
+
 __Exploratory Data Analysis__
+
+taking a look at our variables we see that they all have a reasionably normal distribution.
+
+```{r}
+hists <- merged %>% 
+  select(x3fg, fg_percent, rpg, st, to,  bkpg)
+
+histogram <- function(x, t){
+  ggplot(data = hists, aes(x)) + 
+    geom_histogram(bins = 30, color = "black", fill = "gray98") + 
+    theme_light(base_size = 5) + 
+    labs(title = paste0("Distribution of ", t), x = t, y = "") + 
+    NULL
+
+}
+
+a <- histogram(merged$x3fg,       t = "Three-Points goals Scored")
+b <- histogram(merged$fg_percent, t = "Percentage of Field Goals Scored")
+c <- histogram(merged$rpg,        t = "Rebounds Per Game")
+d <- histogram(merged$st,         t = "Steels Per Game")
+e <- histogram(merged$to,         t = "Turn Overs Per Game")
+f <- histogram(merged$bkpg,       t = "Blocks Per game")
+(a + b + c) / (d + e + f)
+```
+
+```{r}
+stat <- function(x, df = hists, rounding_digits = 2){
+    x <- enquo(x)
+    df %>%
+      summarise(
+                       Mean  = mean( !! x),
+                     Median  = median( !! x),
+                         Std = sd  ( !! x),
+                         Min = min ( !! x),
+                         Max = max ( !! x),
+                       Range = max(!! x) - min(!! x) ) %>%
+      mutate_if(is.numeric, round, rounding_digits)
+}
+
+variables <- c(
+"Three-Points goals Scored",
+"Field Goals Scored Percentage",
+"Rebounds Per Game",
+"Steels Per Game",
+"Turn Overs Per Game",
+"Blocks Per game")
+
+hists %>% 
+  map_df(stat) %>% 
+  mutate(Variable = variables , .before = Mean) %>% 
+  kable(
+  format = "latex",
+  booktabs = TRUE,
+  escape = FALSE,
+  longtable = TRUE,
+  caption = "Descriptive Statistics")
+```
+
 <!-- • Perform exploratory data analysis (with summary statistics or exploratory graphs) that reveals some basic and general information of the data set. -->
 __Description__  
+To gather these data, we used the tidyr, [@R-tidyr] dplyr, [@R-dplyr] and purrr [@R-purrr] package to read-in clean, and prepare our data for analysis.
+Since this is a predictive model we used all the data we had at our disposal that was not colinear.
+
 <!-- • Provide a description and explanation of the steps that you have taken to obtain the best model (or models). -->

### 2021-04-05 paper/paper.tex
update paper

@@
 \PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
 %
@@
   english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
   pdftitle={Using GLMs to Predict Basketball Games},
-  pdfauthor={Sabrina Ball1 \& Joseph Despres1},
+  pdfauthor={Joe Despres \& Sabrina Ball},
   pdflang={en-EN},
-  pdfkeywords={keywords},
-  hidelinks,
+  pdfkeywords={keyword},
+  colorlinks=true,
+  linkcolor=Maroon,
+  filecolor=Maroon,
+  citecolor=Blue,
+  urlcolor=blue,
   pdfcreator={LaTeX via pandoc}}
@@
 
+% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
+% See https://github.com/crsh/papaja/issues/292
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
+
 % \usepackage{etoolbox}
@@
 \makeatother
-\shorttitle{GLM}
-\keywords{keywords\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
-\usepackage{lineno}
-
-\linenumbers
+\shorttitle{Using GLMs to Predict Basketball Games}
+\keywords{keyword\newline\indent Word count: X}
 \usepackage{csquotes}
@@
 \title{Using GLMs to Predict Basketball Games}
-\author{Sabrina Ball\textsuperscript{1} \& Joseph Despres\textsuperscript{1}}
+\author{Joe Despres\textsuperscript{} \& Sabrina Ball\textsuperscript{}}
 \date{}
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Michigan State University}
+\affiliation{\vspace{0.5cm}\textsuperscript{} Michigan State University}
 
@@
 
-testing citat (R Core Team, 2020)
-\textbf{Exploratory Data Analysis}\\
+Every year top university basketball teams compete in the March Madness tournament.
+Casual fans and enthusiasts submit predictions gambling small sums of money on the outcomes.
+This study will determine if which GLM performs best in predicting the results.
+Our data comes from three different sources Kaggle, NCAA, and the tournament results.
+Kagggle provides a comprehensive dataset\footnote{CITE THIS: \url{https://plexkits.com/march-madness-bracket/}} for all NCAA in season (non-turniment) basektball games from 2001 to 2020.
+The NCAA provides team-level statistics\footnote{CITE THIS: \url{https://stats.ncaa.org/rankings/change_sport_year_div}} for each team.
+We will use those two datasets to train a model that uses team-level statistics to predict basketball game outcomes.
+We will test our model on the actual tournament results after it is concluded on April 5th.
+The objective is to predict the individual game outcomes as accurately as possible.
+
+Our response is the outcome \emph{win/loss} as a function of team statistics.
+Win or loss coded 0 or 1.
+Then we will look at both team's statistics and determine which is more important.
+In our dataset we have: \emph{field goal percentage} which is the rate attempt to score vs actually scoring.
+\emph{Free-throw percentage}, is the rate that the team has the opportunity to take an unguarded shot.
+Cumulative \emph{Three-point goals made}.
+\emph{Rebounds per game}, counts the times the team recovered the ball after a missed shot.
+\emph{Steals}, the times a team was able to remove possession of the ball from the opposing team.
+\emph{Turnover}, the number of times the team lost possession.
+\emph{Blocks}, The number of times the team was able to block a shot made by the opposing team.
+
+We will use three regression methods, logistic, Poisson, and multinominal, to compare the results to determine which is most accurate.
+Logistic regression most naturally fits this problem because games do not tie so there is a 50/50 split of wins and losses.
+Also, we will try Poisson regression to predict the number of points scored by each team.
+Since there are many close basketball games we will use multinomial regression to predict if the game score is going to be a large difference one way, small difference, or a large difference the other way.
+
+We Found\ldots.
+
+\textbf{Exploratory Data Analysis}
+
+taking a look at our variables we see that they all have a reasionably normal distribution.
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+
+\begin{longtable}[t]{lrrrrrr}
+\caption{\label{tab:unnamed-chunk-2}Descriptive Statistics}\\
+\toprule
+Variable & Mean & Median & Std & Min & Max & Range\\
+\midrule
+Three-Points goals Scored & 242.08 & 241.00 & 48.18 & 107.0 & 454.00 & 347.00\\
+Field Goals Scored Percentage & 44.03 & 44.10 & 2.46 & 35.4 & 52.60 & 17.20\\
+Rebounds Per Game & 35.48 & 35.35 & 2.49 & 27.6 & 43.81 & 16.21\\
+Steels Per Game & 206.34 & 202.00 & 40.59 & 116.0 & 369.00 & 253.00\\
+Turn Overs Per Game & 421.84 & 419.00 & 49.01 & 290.0 & 604.00 & 314.00\\
+\addlinespace
+Blocks Per game & 3.30 & 3.20 & 0.94 & 1.3 & 6.80 & 5.50\\
+\bottomrule
+\end{longtable}
+
 \textbf{Description}\\
+To gather these data, we used the tidyr, (Wickham, 2021) dplyr, (Wickham, François, Henry, \& Müller, 2021) and purrr (Henry \& Wickham, 2020) package to read-in clean, and prepare our data for analysis.
+Since this is a predictive model we used all the data we had at our disposal that was not colinear.
+
 \textbf{Results}\\
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
+\leavevmode\hypertarget{ref-R-purrr}{}%
+Henry, L., \& Wickham, H. (2020). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+Wickham, H. (2021). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}
+
+\leavevmode\hypertarget{ref-R-dplyr}{}%
+Wickham, H., François, R., Henry, L., \& Müller, K. (2021). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://CRAN.R-project.org/package=dplyr}
 

### 2021-04-05 scrape_finals.R
added scraping script

@@ -1,0 +1,57 @@
+library(rvest)
+
+url <- "https://www.ncaa.com/news/basketball-men/mml-official-bracket/2021-04-05/2021-ncaa-bracket-printable-march-madness-bracket-pdf"
+webpage <- read_html(url)
+
+rank_data_html <- html_nodes(webpage,'.embedded-entity a')
+
+rank_data_html %>% 
+  html_text(trim = TRUE) %>% 
+  html_table()
+
+links <- rank_data_html %>% 
+  html_attr("href")
+
+links
+
+
+
+url <- "https://www.ncaa.com/march-madness-live/game/208?cid=mml2021_editorial_gamecenter"
+webpage <- read_html(url)
+
+score_getter <- function (link) {
+  webpage <- read_html(link)
+  rank_data_html <- html_nodes(webpage,'.color_lvl_5')
+  teams <- html_nodes(webpage,'.h4')
+  
+  teams <- teams %>% 
+    html_text(trim = TRUE)
+  
+  raw <- rank_data_html %>% 
+    html_text(trim = TRUE)
+  team <- raw[1:2]
+  team2 <- raw[5:6]
+  output <- c(raw, teams)
+  return(output)
+    
+}
+
+links <- subset(links, grepl("https://www.ncaa.com/game/", links) )
+
+lapply(links, score_getter)
+
+score_getter(links[1])
+
+
+# this works the best -----------------------------------------------------
+
+
+url <- "https://www.ncaa.com/march-madness-live/scores"
+webpage <- read_html(url)
+
+rank_data_html <- html_nodes(webpage,'.lvp')
+
+data <- rank_data_html %>% 
+  html_text(trim = TRUE)
+
+tibble::tibble(data[12:277])

### 2021-04-06 R/scrape_finals.R
added scraping script

@@ -1,57 +1,54 @@
+
+# Scrape finals scores ----------------------------------------------------
+
 library(rvest)
+library(dplyr)
 
-url <- "https://www.ncaa.com/news/basketball-men/mml-official-bracket/2021-04-05/2021-ncaa-bracket-printable-march-madness-bracket-pdf"
-webpage <- read_html(url)
-
-rank_data_html <- html_nodes(webpage,'.embedded-entity a')
-
-rank_data_html %>% 
-  html_text(trim = TRUE) %>% 
-  html_table()
-
-links <- rank_data_html %>% 
-  html_attr("href")
-
-links
-
-
-
-url <- "https://www.ncaa.com/march-madness-live/game/208?cid=mml2021_editorial_gamecenter"
-webpage <- read_html(url)
-
-score_getter <- function (link) {
-  webpage <- read_html(link)
-  rank_data_html <- html_nodes(webpage,'.color_lvl_5')
-  teams <- html_nodes(webpage,'.h4')
-  
-  teams <- teams %>% 
-    html_text(trim = TRUE)
-  
-  raw <- rank_data_html %>% 
-    html_text(trim = TRUE)
-  team <- raw[1:2]
-  team2 <- raw[5:6]
-  output <- c(raw, teams)
-  return(output)
-    
-}
-
-links <- subset(links, grepl("https://www.ncaa.com/game/", links) )
-
-lapply(links, score_getter)
-
-score_getter(links[1])
-
-
-# this works the best -----------------------------------------------------
-
+options(tibble.print_max = 50, tibble.print_min = 50)
 
 url <- "https://www.ncaa.com/march-madness-live/scores"
-webpage <- read_html(url)
 
-rank_data_html <- html_nodes(webpage,'.lvp')
-
-data <- rank_data_html %>% 
+scraped_final_scores <- read_html(url) %>% 
+  html_nodes('.lvp') %>% 
   html_text(trim = TRUE)
 
-tibble::tibble(data[12:277])
+games  <- tibble::tibble(scraped_final_scores[12:275]) %>% 
+  rename(game = 1) %>% 
+  mutate(id = rep(c("a", "b"), 132)) # this will break when game is done
+
+spliter <- function(df, tag){
+  df  %>% 
+    filter(id == tag)
+}
+
+a <- spliter(games, "a")
+b <- spliter(games, "b")
+
[…]

### 2021-04-06 R/scrape_finals.R
scrape complete adding data and final script

@@ -12,9 +12,11 @@ scraped_final_scores <- read_html(url) %>%
   html_nodes('.lvp') %>% 
   html_text(trim = TRUE)
 
-games  <- tibble::tibble(scraped_final_scores[12:275]) %>% 
-  rename(game = 1) %>% 
-  mutate(id = rep(c("a", "b"), 132)) # this will break when game is done
+scraped_final_scores[1:300]
+
+games  <- tibble::tibble(scraped_final_scores[12:279]) %>% 
+    rename(game = 1) %>% 
+  mutate(id = rep(c("a", "b"), 268/2)) # this will break when game is done
 
 spliter <- function(df, tag){
   df  %>% 
@@ -34,7 +36,7 @@ scores <- bind_cols(a, b) %>%
   mutate(team = stringr::str_to_lower(team)) %>% 
   left_join(MTeamSpellings, by = "team") %>% 
   select(1, 2, 4) %>% 
-  mutate(id = rep(c("a", "b"), 132/2)) # this will break when game is done
+  mutate(id = rep(c("a", "b"), 134/2)) # this will break when game is done
   
 a <- spliter(scores, "a")
 b <- spliter(scores, "b")
@@ -49,6 +51,6 @@ bind_cols(a, b) %>%
   select(1:3, 5:7) %>% 
   mutate(game = paste0(team, " vs ", otherteam),
          winner = if_else(teamscore > otherteamscore, team, otherteam)) %>% 
-  select(game, winner, teamscore, otherteamscore) %>% 
+  select(game, winner, teamscore, otherteamscore) %>%
   readr::write_csv(here::here("data", "results.csv"))
   

### 2021-04-07 R/independent_modeling.R
cleaned up script and and re-ran models

@@ -12,7 +12,6 @@ options(tibble.print_min = 20)
 # TODO: recipes::step_normalize() try that see if there is a difference
 
 # logsitic fit ------------------------------------------------------------
-
 # logistic_fit <- glm(win ~ x3fg +
 #                       opposingx3fg +
 #                       # field goal pct
@@ -34,26 +33,26 @@ options(tibble.print_min = 20)
 #                       opposingbkpg +
 #                       bkpg,
 #                     data = merged, family = "binomial")
-# 
+# # 
 # start <- Sys.time()
 # logistic_predictions <- ids %>%
 #   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
 #                                     .y = ids$otherid,
-#                                     .f = logistic_predictor))
+#                                     .f = logistic_predictor)) %>%
+# # This takes about 30 seconds to run..not if we cache it 
+#   mutate(
+#     predicted_winner = if_else(predicted_probs > .5, team, other_team),
+#     predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
+#     predicted_loser_probs = 1 - predicted_winner_probs,
+#     prediction_points = predicted_winner_probs - predicted_loser_probs,
+#     .after = game
+#   )
 # paste0("runtime: ", round(Sys.time() - start, 2))
-# # This takes about 30 seconds to run..not if we cache it
+
 # saveRDS(logistic_predictions, here::here("cache", "logistic_predictions.RDS"))
 
-logistic_predictions <- 
-  readRDS(here::here("cache", "logistic_predictions.RDS")) %>% 
-  mutate(  
-    predicted_winner = if_else(predicted_probs > .5, team, other_team),
-    predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
-    predicted_loser_probs = 1 - predicted_winner_probs,
-    prediction_points = predicted_winner_probs - predicted_loser_probs,
-    .after = game
-  )
 
+logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
 logistic_predictions
 
 
@@ -111,7 +110,6 @@ logistic_predictions
 # saveRDS(posson_prediction, here::here("cache", "posson_prediction.RDS"))
 
[…]

### 2021-04-07 R/model_valadation.R
added valadation file

@@ -1,0 +1,10 @@
+library(tidymodels)
+
+
+logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
+
+posson_prediction <- readRDS(here::here("cache", "posson_prediction.RDS"))
+
+multinom_prediction <- readRDS(here::here("cache", "multinom_prediction.RDS"))
+
+results <- read_csv(here::here("data", "data/results.csv"))

### 2021-04-07 R/scrape_finals.R
added all columns to csv

@@ -12,8 +12,6 @@ scraped_final_scores <- read_html(url) %>%
   html_nodes('.lvp') %>% 
   html_text(trim = TRUE)
 
-scraped_final_scores[1:300]
-
 games  <- tibble::tibble(scraped_final_scores[12:279]) %>% 
     rename(game = 1) %>% 
   mutate(id = rep(c("a", "b"), 268/2)) # this will break when game is done
@@ -51,6 +49,5 @@ bind_cols(a, b) %>%
   select(1:3, 5:7) %>% 
   mutate(game = paste0(team, " vs ", otherteam),
          winner = if_else(teamscore > otherteamscore, team, otherteam)) %>% 
-  select(game, winner, teamscore, otherteamscore) %>%
   readr::write_csv(here::here("data", "results.csv"))
   

### 2021-04-07 R/cleaningScript.R
removed restriction on combinations

@@ -171,7 +171,7 @@ team2id <- pull(s2021, TeamID)
 crossing(teamid = team1id, other = team2id) %>%
   distinct() %>%
   filter(teamid != other) %>%
-  top_n(2016) %>%   # choose(64,2) = 2016 (removing double counts)
+  top_n(2016) %>% 
   left_join(s2021, by = c("other" = "TeamID")) %>%
   rename(other_team = team) %>%
   select(teamid:other_team) %>%

### 2021-04-07 R/scrape_finals.R
updated scrape

@@ -46,8 +46,11 @@ bind_cols(a, b) %>%
             otherteamscore = 6, 
             teamid = 3, 
          otherteamid = 7) %>% 
-  select(1:3, 5:7) %>% 
-  mutate(game = paste0(team, " vs ", otherteam),
+  select(1:3, 5:7) %>%
+  
+  mutate(team = gsub("[^[:alnum:]]", "", team),
+         otherteam = gsub("[^[:alnum:]]", "", otherteam),
+    game = paste0(team, " vs. ", otherteam),
          winner = if_else(teamscore > otherteamscore, team, otherteam)) %>% 
   readr::write_csv(here::here("data", "results.csv"))
   

### 2021-04-07 R/cleaningScript.R
update files

@@ -171,7 +171,7 @@ team2id <- pull(s2021, TeamID)
 crossing(teamid = team1id, other = team2id) %>%
   distinct() %>%
   filter(teamid != other) %>%
-  top_n(2016) %>% 
+  # top_n(2016) %>% 
   left_join(s2021, by = c("other" = "TeamID")) %>%
   rename(other_team = team) %>%
   select(teamid:other_team) %>%

### 2021-04-07 R/independent_modeling.R

@@ -12,6 +12,7 @@ options(tibble.print_min = 20)
 # TODO: recipes::step_normalize() try that see if there is a difference
 
 # logsitic fit ------------------------------------------------------------
+
 # logistic_fit <- glm(win ~ x3fg +
 #                       opposingx3fg +
 #                       # field goal pct
@@ -33,14 +34,14 @@ options(tibble.print_min = 20)
 #                       opposingbkpg +
 #                       bkpg,
 #                     data = merged, family = "binomial")
-# # 
+# 
 # start <- Sys.time()
+# 
 # logistic_predictions <- ids %>%
 #   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
 #                                     .y = ids$otherid,
 #                                     .f = logistic_predictor)) %>%
-# # This takes about 30 seconds to run..not if we cache it 
-#   mutate(
+#    mutate(
 #     predicted_winner = if_else(predicted_probs > .5, team, other_team),
 #     predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
 #     predicted_loser_probs = 1 - predicted_winner_probs,
@@ -48,9 +49,9 @@ options(tibble.print_min = 20)
 #     .after = game
 #   )
 # paste0("runtime: ", round(Sys.time() - start, 2))
-
+# 
 # saveRDS(logistic_predictions, here::here("cache", "logistic_predictions.RDS"))
-
+# 
 
 logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
 logistic_predictions
@@ -106,14 +107,14 @@ logistic_predictions
 #                                     .y = ids$teamid,
 #                                     .f = poisson_predictor))
 # paste0("runtime: ", round(Sys.time() - start, 2))
-## this takes about a 160 secs to run
+# # this takes about a 160 secs to run
 # saveRDS(posson_prediction, here::here("cache", "posson_prediction.RDS"))
 
 posson_prediction <- readRDS(here::here("cache", "posson_prediction.RDS"))
 posson_prediction
 
 # mulitnom ----------------------------------------------------------------
-# 
+#
 # fit_multinom <- VGAM::vglm(x_tile ~ x3fg +
 #                  opposingx3fg +
 #                  # field goal pct
@@ -172,12 +173,12 @@ posson_prediction
 #   return(x)
 # }
 # 
[…]

### 2021-04-07 R/scrape_finals.R

@@ -53,4 +53,3 @@ bind_cols(a, b) %>%
     game = paste0(team, " vs. ", otherteam),
          winner = if_else(teamscore > otherteamscore, team, otherteam)) %>% 
   readr::write_csv(here::here("data", "results.csv"))
-  

### 2021-04-07 R/model_valadation.R
scored models.

@@ -1,10 +1,72 @@
 library(tidymodels)
 
+options(tibble.print_max = 50, tibble.print_min = 30)
 
+results <- readr::read_csv(here::here("data", "results.csv"))
 logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
-
 posson_prediction <- readRDS(here::here("cache", "posson_prediction.RDS"))
-
 multinom_prediction <- readRDS(here::here("cache", "multinom_prediction.RDS"))
 
-results <- read_csv(here::here("data", "data/results.csv"))
+
+gameplayed <- results %>% 
+  mutate(gameid = paste0(teamid, "vs", otherteamid))
+
+
+# logistic scoring --------------------------------------------------------
+
+toscore <- function (df) {
+  df %>% 
+    mutate(gameid = paste0(teamid, "vs", otherid)) %>% 
+    filter(gameid %in% pull(gameplayed, gameid)) %>% 
+    left_join(gameplayed, by = c("team", "game", "teamid", "gameid"))
+  
+}
+
+logistic_model_score <- toscore(logistic_predictions) %>% 
+  mutate(score = if_else(predicted_winner == winner, predicted_winner_probs, -1*predicted_winner_probs)) 
+
+logistic_model_score %>% 
+summarise(mean = mean(score, na.rm = TRUE),
+             sd = sd(score, na.rm = TRUE),
+            t_stat = mean / sd, 
+            p = pt(t_stat, n() - 1, lower.tail = FALSE))
+
+
+# poisson scoring ---------------------------------------------------------
+
+posson_prediction  <- toscore(posson_prediction) %>% 
+  mutate(predicted_difference = predicted_score - other_predicted_score,
+         actual_difference = teamscore - otherteamscore,
+         score = predicted_difference - actual_difference) 
+
+posson_model_score %>% 
+  summarise(mean = mean(score, na.rm = TRUE),
+            sd = sd(score, na.rm = TRUE),
+            t_stat = mean / sd, 
+            p = pt(t_stat, n() - 1, lower.tail = FALSE))
+
+# multinomeal scoring -----------------------------------------------------
+
+xtile_matcher <- function (x) {
+  ifelse(x > 12 , test$`<-12` , 
+         ifelse(x >= -12 & x < -4, test$`-12:-4`,
+                ifelse(-4 & x <  4, test$`-4:4`,
[…]

### 2021-04-08 R/model_valadation.R
added multinomeal score

@@ -47,14 +47,7 @@ posson_model_score %>%
 
 # multinomeal scoring -----------------------------------------------------
 
-xtile_matcher <- function (x) {
-  ifelse(x > 12 , test$`<-12` , 
-         ifelse(x >= -12 & x < -4, test$`-12:-4`,
-                ifelse(-4 & x <  4, test$`-4:4`,
-                       ifelse(x >= 4 & x < 13, test$`4:13`, test$`>13`))))
-}
-
-test <- toscore(multinom_prediction) %>% 
+multinomeal_model_score  <- toscore(multinom_prediction) %>% 
   mutate(diff = teamscore - otherteamscore,
   fifth_xtile = case_when(diff <  -12             ~ '<-12',
                           diff >= -12 & diff < -4 ~ '-12:-4',
@@ -68,5 +61,8 @@ test <- toscore(multinom_prediction) %>%
                     diff >=   4 & diff < 13  ~ `4:13`,
                     diff >= 13               ~ `>13`)) %>% 
   select(`<-12`:`>13`, fifth_xtile, probs) %>%  # TODO: come up with a way to score this
-  mutate(score = if_else(probs > 0.25, probs * 4, 1-probs))
+  mutate(score = if_else(probs > 0.25, (probs * 4), (1 - probs))) 
 
+multinomeal_model_score %>% 
+  pull(score) %>% 
+  hist()

### 2021-04-08 proposal/proposal.Rmd
added link and changed font

@@
 This study will determine if which GLM performs best in predicting the results."
-geometry: margin=1in
+geometry: margin=0.5in
 fontfamily: mathpazo
-fontsize: 11pt
+fontsize: 10pt
 # spacing: double
 endnote: no
+
 ---
@@
 We will use those two datasets to train a model that uses team-level statistics to predict basketball game outcomes.
-We will test our model on the actual tournament results after it is concluded on April 5th.
+We will test our model on the actual tournament results which we obtained with a scraping script^[https://github.com/despresj/March-Madness/blob/main/R/scrape_finals.R].
   The objective is to predict the individual game outcomes as accurately as possible.
-
 Our response is the outcome *win/loss* as a function of team statistics.

### 2021-04-08 slides/slides.Rmd
adding slides

---
title: "March Madness Basketball Turniment"
author: "Sabrina Ball and Joe Despres"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(https://upload.wikimedia.org/wikipedia/en/2/28/March_Madness_logo.svg)

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---

https://github.com/EvaMaeRey/flipbookr

---

### 2021-04-09 proposal/proposal.Rmd
made edits

@@
 title: "Predicting the Outcome of Basketball Games with GLMs"
+header-includes:
+   - \setlength\parindent{24pt}
 author:
@@
   affiliation:
-abstract: "Every year top university basketball teams compete in the March Madness tournament.
-Casual fans and Enthuasists submit predictions gambling small sums of money on the outcomes.
-This study will determine if which GLM performs best in predicting the results."
+abstract: \hspace{\parindent}Every year top university basketball teams compete in the March Madness tournament.
+Casual fans and enthuasists submit predictions gambling small sums of money on the outcomes.
+This study will determine which GLM performs best in predicting the results.
 geometry: margin=0.5in
@@
 endnote: no
+indent: true
 
@@
 # Data
-Our data comes from three different sources Kaggle, NCAA, and the tournament results.
-Kagggle provides a comprehensive dataset^[https://plexkits.com/march-madness-bracket/] for all NCAA in season (non-turniment) basektball games from 2001 to 2020.
+
+\hspace{\parindent} Our data comes from three different sources Kaggle, NCAA, and the tournament results.
+Kaggle provides a comprehensive dataset^[https://plexkits.com/march-madness-bracket/] for all NCAA in season (non-tournament) basektball games from 2001 to 2020.
 The NCAA provides team-level statistics^[https://stats.ncaa.org/rankings/change_sport_year_div] for each team.
@@
 We will test our model on the actual tournament results which we obtained with a scraping script^[https://github.com/despresj/March-Madness/blob/main/R/scrape_finals.R].
-  The objective is to predict the individual game outcomes as accurately as possible.
+
+\hspace{\parindent}The objective is to predict the individual game outcomes as accurately as possible.
 Our response is the outcome *win/loss* as a function of team statistics.
@@
 *Free-throw percentage*,  is the rate that the team has the opportunity to take an unguarded shot.
-Cumulative *Three-point goals made*.
+Cumulative *three-point goals made*.
 *Rebounds per game*, counts the times the team recovered the ball after a missed shot.
@@
 *Turnover*, the number of times the team lost possession.
-*Blocks*, The number of times the team was able to block a shot made by the opposing team.
+*Blocks*, the number of times the team was able to block a shot made by the opposing team.
 
@@
 
-We want to learn how to make and validate predictions with GLMs.
+\hspace{\parindent}We want to learn how to make and validate predictions with GLMs.
 Additionally, we want an appropriate scoring metric that can compare the performance of various methods.
@@
 
-We will use three regression methods, logistic, Poisson, and multinominal, to compare the results to determine which is most accurate.
+\hspace{\parindent}We will use three regression methods, logistic, poisson, and multinominal, to compare the results to determine which is most accurate.
 Logistic regression most naturally fits this problem because games do not tie so there is a 50/50 split of wins and losses.
-Also, we will try Poisson regression to predict the number of points scored by each team.
+Also, we will try poisson regression to predict the number of points scored by each team.
 Since there are many close basketball games we will use multinomial regression to predict if the game score is going to be a large difference one way, small difference, or a large difference the other way.
 
-For the most part, assembling the data was not an issue yet I was worried about that.
+\hspace{\parindent}For the most part, assembling the data was not an issue yet I was worried about that.
 An issue we will face is scoring this model.

### 2021-04-09 R/model_valadation.R
update files

@@ -63,6 +63,4 @@ multinomeal_model_score  <- toscore(multinom_prediction) %>%
   select(`<-12`:`>13`, fifth_xtile, probs) %>%  # TODO: come up with a way to score this
   mutate(score = if_else(probs > 0.25, (probs * 4), (1 - probs))) 
 
-multinomeal_model_score %>% 
-  pull(score) %>% 
-  hist()
+multinomeal_model_score

### 2021-04-09 slides/slides.Rmd
update slides

@@
 ```{r setup, include=FALSE}
-options(htmltools.dir.version = FALSE)
+knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
+library(flipbookr)
+library(tidyverse)
+
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+ids <- readr::read_csv(here::here("data", "ids.csv"))
+s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
+```
+
+<style type="text/css">
+.remark-slide-content {
+    font-size: 30px;
+    padding: 1em 4em 1em 4em;
+}
+</style>
+
+---
+```{r}
+logistic_predictor_fn <- function (team_1_id, team_2_id) {
+  stats_team_1 <- filter(s2021, TeamID == team_1_id)
+  stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
+                         rename_all(~ paste0("opposing",.)) 
+  bound_cols <- bind_cols(stats_team_1, stats_team_2)
+  pred <- predict(logistic_fit, newdata = bound_cols, type = "response")
+
+  return(pred)
+}
+
+logistic_predictor <- purrr::possibly(.f = logistic_predictor_fn, otherwise = "ERROR")
 ```
@@
 
-https://github.com/EvaMaeRey/flipbookr
+```{r}
+logistic_fit <- glm(win ~ x3fg + opposingx3fg +
+                      # field goal pct
+                      fg_percent + opposingfg_percent +
+                      # free throws
+                      ft_percent + opposingft_percent +
+                      # rebound per game
+                      rpg + opposingrpg +
+                      # steels
+                      st + opposingst +
+                      #turnover
+                      to + opposingto +
+                      # blocks
+                      opposingbkpg + bkpg,
+                    data = merged, family = "binomial")
+```
+---
+`r chunk_reveal("logistic")`
+
+```{r logistic}
+ids %>%
+  mutate(predicted_probs = map2_dbl(.x = ids$teamid,
+                                    .y = ids$otherid,
+                                    .f = logistic_predictor)) %>% 
+   mutate(
+    predicted_winner = if_else(predicted_probs > .5, team, other_team),
+    predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
+    predicted_loser_probs = 1 - predicted_winner_probs,
+    prediction_points = predicted_winner_probs - predicted_loser_probs,
+    .after = game
+  )
+
+```
 
 ---
+
+`r chunk_reveal("my_cars")`
+
+```{r my_cars, include = FALSE}
+cars %>%
+  filter(speed > 4) %>%
+  ggplot() +
+  aes(x = speed) + #BREAK
+  aes(y = dist) + #BREAK
+  geom_point(
+    alpha = .8, 
+    color = "blue" 
+    ) + 
+  aes(size = speed) #BREAK
+```

### 2021-04-09 R/independent_modeling.R
added files

@@ -13,45 +13,45 @@ options(tibble.print_min = 20)
 
 # logsitic fit ------------------------------------------------------------
 
-# logistic_fit <- glm(win ~ x3fg +
-#                       opposingx3fg +
-#                       # field goal pct
-#                       fg_percent +
-#                       opposingfg_percent +
-#                       # free throws
-#                       ft_percent +
-#                       opposingft_percent +
-#                       # rebound per game
-#                       rpg +
-#                       opposingrpg +
-#                       # steels
-#                       st +
-#                       opposingst +
-#                       #turnover
-#                       to +
-#                       opposingto +
-#                       # blocks
-#                       opposingbkpg +
-#                       bkpg,
-#                     data = merged, family = "binomial")
-# 
-# start <- Sys.time()
-# 
-# logistic_predictions <- ids %>%
-#   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
-#                                     .y = ids$otherid,
-#                                     .f = logistic_predictor)) %>%
-#    mutate(
-#     predicted_winner = if_else(predicted_probs > .5, team, other_team),
-#     predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
-#     predicted_loser_probs = 1 - predicted_winner_probs,
-#     prediction_points = predicted_winner_probs - predicted_loser_probs,
-#     .after = game
-#   )
-# paste0("runtime: ", round(Sys.time() - start, 2))
-# 
-# saveRDS(logistic_predictions, here::here("cache", "logistic_predictions.RDS"))
-# 
+logistic_fit <- glm(win ~ x3fg +
+                      opposingx3fg +
+                      # field goal pct
+                      fg_percent +
+                      opposingfg_percent +
+                      # free throws
+                      ft_percent +
+                      opposingft_percent +
+                      # rebound per game
+                      rpg +
+                      opposingrpg +
+                      # steels
+                      st +
+                      opposingst +
+                      #turnover
+                      to +
+                      opposingto +
+                      # blocks
+                      opposingbkpg +
+                      bkpg,
+                    data = merged, family = "binomial")
+
+start <- Sys.time()
+
+logistic_predictions <- ids %>%
+  mutate(predicted_probs = map2_dbl(.x = ids$teamid,
+                                    .y = ids$otherid,
[…]

### 2021-04-11 slides/slides.Rmd
worked on slides

--- 2021-04-08 slides/slides.Rmd
@@
 background-image: url(https://upload.wikimedia.org/wikipedia/en/2/28/March_Madness_logo.svg)
+---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(fig.width = 6, echo = FALSE, message = FALSE, warning = FALSE, comment = "", cache = F)
 
-```{r setup, include=FALSE}
-options(htmltools.dir.version = FALSE)
+library(tidyverse)
+library(kableExtra)
+library(patchwork)
+
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+ids <- readr::read_csv(here::here("data", "ids.csv"))
+s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
+source(here::here("R", "fitting.R"))
+```
+
+<!-- --- -->
+<!-- # Backround on the turniment -->
+<!-- quick facts... -->
+
+<!-- --- -->
+
+<!-- ![](images/empty_bracket.jpg) -->
+
+<!-- --- -->
+<!-- class: center, middle -->
+<!-- # Zoomed in -->
+
+<!-- ![](images/zoomed.png) -->
+
+<!-- --- -->
+<!-- # Goal -->
+<!-- Can we use basketball seasion data to predict this turniment outcomes? -->
+
+---
+
+# Training Data
+
+```{r}
+descriptive_statistics <- function(x){
+  output <- c(mean(x), median(x), sd(x), min(x), max(x))
+  output <- round(output, digits = 2)
+  return(output)
+}
+
+desc_stats <- merged %>% 
+  select(fg_percent, ft_percent, rpg,st, to, bkpg) %>% 
+  map_df(descriptive_statistics) %>% 
+  t()
+
+colnames(desc_stats) <- c("Mean", "Median", "Standard DEv", "Min", "Max")
+rownames(desc_stats) <- c("Field Goal Percentage", "Free Throw Percentage", "Rebounds Per Game", "Steals", "Turnovers", "Blocked Shots Per Game")
+
+kable(desc_stats) %>%
+  kable_styling(bootstrap_options = c("striped", "hover")) 
+```
+
+---
+class: center
+
+# Distribution of Predictors
+
+```{r fig.width=15, fig.height=10}
+hists <- merged %>% 
+  select(x3fg, fg_percent, rpg, st, to,  bkpg)
+
+histogram <- function(x, t){
+  ggplot(data = hists, aes(x)) + 
+    geom_histogram(bins = 50, color = "black", fill = "gray80") + 
+    theme_light(base_size = 20) + 
+    labs(title = paste0("", t), x = "", y = "") + 
+    NULL
+
+}
+
+a <- histogram(merged$x3fg,       t = "Three-Points Goals")
+b <- histogram(merged$fg_percent, t = "Percentage of Goals Scored")
+c <- histogram(merged$rpg,        t = "Rebounds Per Game")
+d <- histogram(merged$st,         t = "Steals Per Game")
+e <- histogram(merged$to,         t = "Turnovers Per Game")
+f <- histogram(merged$bkpg,       t = "Blocks Per game")
+# (a + b + c) / (d + e + f)
+(a + b) / (c + d) /(e + f)
 ```
@@
 
-https://github.com/EvaMaeRey/flipbookr
+# Model
+
+```{r}
+
+equatiomatic::extract_eq(model = logistic_fit, 
+                         use_coefs = FALSE, 
+                         show_distribution = TRUE, 
+                         wrap = TRUE, 
+                         terms_per_line = 2, 
+                         intercept = "beta")
+# results='asis' in code chunk to render as LaTex
+```
 
 ---
+
+```{r}
+
+equatiomatic::extract_eq(model = fit_poisson, 
+                         use_coefs = FALSE, 
+                         show_distribution = TRUE, 
+                         wrap = TRUE, 
+                         terms_per_line = 2, 
+                         intercept = "beta")
+```
+---
+
+# Results
+
+```{r}
+# require(memisc)
+# mtable('Model 1' = logistic_fit,
+#             'Model 2' = fit_poisson,
+#             summary.stats = c('R-squared','F','p','N'), )
+
+```

### 2021-04-11 R/fitting.R

@@ -1,0 +1,65 @@
+logistic_fit <- glm(win ~ x3fg +
+                       opposingx3fg +
+                       # field goal pct
+                       fg_percent +
+                       opposingfg_percent +
+                       # free throws
+                       ft_percent +
+                       opposingft_percent +
+                       # rebound per game
+                       rpg +
+                       opposingrpg +
+                       # steels
+                       st +
+                       opposingst +
+                       #turnover
+                       to +
+                       opposingto +
+                       # blocks
+                       opposingbkpg +
+                       bkpg,
+                     data = merged, family = "binomial")
+ 
+ fit_poisson <- glm(team_score ~ x3fg +
+              opposingx3fg +
+              # field goal pct
+              fg_percent +
+              opposingfg_percent +
+              # free throws
+              ft_percent +
+              opposingft_percent +
+              # rebound per game
+              rpg +
+              opposingrpg +
+              # steels
+              st +
+              opposingst +
+              #turnover
+              to +
+              opposingto +
+              # blocks
+              opposingbkpg +
+              bkpg,
+              data = merged, family = poisson(link = "log"))
+ 
+ fit_multinom <- VGAM::vglm(x_tile ~ x3fg +
+                  opposingx3fg +
+                  # field goal pct
+                  fg_percent +
+                  opposingfg_percent +
+                  # free throws
+                  ft_percent +
+                  opposingft_percent +
+                  # rebound per game
+                  rpg +
+                  opposingrpg +
+                  # steels
+                  st +
+                  opposingst +
+                  #turnover
+                  to +
+                  opposingto +
+                  # blocks
+                  opposingbkpg +
+                  bkpg, family = VGAM::cumulative(parallel=TRUE),
+                  data = merged)

### 2021-04-13 slides/slides.Rmd
added files changed in meeting

--- 2021-04-08 slides/slides.Rmd
@@
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(fig.width = 6, echo = FALSE, message = FALSE, warning = FALSE, comment = "", cache = F)
+
+library(tidyverse)
+library(kableExtra)
+library(patchwork)
+
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+ids <- readr::read_csv(here::here("data", "ids.csv"))
+s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
+source(here::here("R", "fitting.R"))
+```
 
@@
 
-```{r setup, include=FALSE}
-options(htmltools.dir.version = FALSE)
+---
+
+# Backround
+
+--
+
+- First tournament was in 1939 only 8 teams
+
+--
+
+- Currently, the top 64 teams qualify
+
+--
+
+- Only one team has ever gone undefeated thought a season and the tournament
+
+--
+
+- 16.9 Million Viewers
+
+--
+
+- MSU won in 1979 and 2000
+
+--
+
+- maybe 2022...
+
+--
+
+- When the turiment starts we form small groups and gamble on the outcomes. Usually by selecting winners and loser from the backet until we have a completely filled out turniment
+
+--
+
+- Nobody has ever made a perfect bracket there are $2^{63}$ possible outcomes.
+
+---
+
+![](images/empty_bracket.jpg)
+
+---
+class: center, middle
+# Zoomed in
+
+![](images/zoomed.png)
+
+---
+
+class: inverse, center, middle
+
+# Can we use basketball seasion data to predict this turniment outcomes?
+
+---
+
+# Predictors
+
+```{r}
+descriptive_statistics <- function(x){
+  output <- c(mean(x), median(x), sd(x), min(x), max(x))
+  output <- round(output, digits = 2)
+  return(output)
+}
+
+desc_stats <- merged %>%
+  select(fg_percent, ft_percent, rpg,st, to, bkpg) %>%
+  map_df(descriptive_statistics) %>%
+  t()
+
+colnames(desc_stats) <- c("Mean", "Median", "Std. Dev", "Min", "Max")
+rownames(desc_stats) <- c("Field Goal Percentage", "Free Throw Percentage", "Rebounds Per Game", "Steals", "Turnovers", "Blocked Shots Per Game")
+
+kable(desc_stats) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"))
+
+seasion_seasion <- merged %>%
+  nrow()
+
+# n teams
+unique_teams <- length(unique(merged$team_id))
+
+# 64 teams in the turniment.
+
+cbind(
+    `Games Played` = seasion_seasion,
+      Teams = unique_teams,
+      Turniment = 64) %>%
+  kable() %>%
+  kable_styling(bootstrap_options = "striped",
+                full_width = F,
+                position = "center")
+```
+
+---
+class: center
+
+# Distribution of Predictors
+
+```{r fig.width=15, fig.height=10}
+hists <- merged %>%
+  select(x3fg, fg_percent, rpg, st, to,  bkpg)
+
+histogram <- function(x, t){
+  ggplot(data = hists, aes(x)) +
+    geom_histogram(bins = 50, color = "black", fill = "gray80") +
+    theme_light(base_size = 20) +
+    labs(title = paste0("", t), x = "", y = "") +
+    NULL
+
+}
+
+a <- histogram(merged$x3fg,       t = "Three-Points Goals")
+b <- histogram(merged$fg_percent, t = "Percentage of Goals Scored")
+c <- histogram(merged$rpg,        t = "Rebounds Per Game")
+d <- histogram(merged$st,         t = "Steals Per Game")
+e <- histogram(merged$to,         t = "Turnovers Per Game")
+f <- histogram(merged$bkpg,       t = "Blocks Per game")
+# (a + b + c) / (d + e + f)
+(a + b) / (c + d) /(e + f)
 ```
@@
 
-https://github.com/EvaMaeRey/flipbookr
+class: center
+
+# Logistic Model
+
+```{r}
+equatiomatic::extract_eq(model = logistic_fit,
+                         use_coefs = FALSE,
+                         show_distribution = TRUE,
+                         wrap = TRUE,
+                         terms_per_line = 2,
+                         intercept = "beta")
+```
 
 ---
+
+# Poisson Model
+
+```{r}
+equatiomatic::extract_eq(model = fit_poisson,
+                         use_coefs = FALSE,
+                         show_distribution = TRUE,
+                         wrap = TRUE,
+                         terms_per_line = 2,
+                         intercept = "beta")
+```
+
+---
+
+# Multinomeal Model
+
+---
+class: center 
+### Output and Interpretation
+
+```{r}
+table_output <- function(obj, rows){
+  broom::tidy(obj) %>% 
+  mutate(p.value = as.character(p.value)) %>%
+  mutate_if(is.numeric, round, 3) %>% 
+  mutate(p.value = as.numeric(p.value))  %>% 
+  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
+  kable_styling(font_size = 15) %>% 
+  row_spec(rows, color = "white", background = "#025cba")
+}
+
+table_output(obj = logistic_fit, rows = 2:3)
+```
+
+---
+class: center 
+## Output and Interpretation
+
+```{r}
+table_output(obj = logistic_fit, rows = 4:5)
+```
+
+---
+class: center 
+## Output and Interpretation
+
+```{r}
+table_output(obj = logistic_fit, rows = 6:7)
+```
+
+---
+class: center 
+## Output and Interpretation
+
+```{r}
+table_output(obj = logistic_fit, rows = 8:9)
+```
+
+---
+class: center 
+## Output and Interpretation
+
+```{r}
+table_output(obj = logistic_fit, rows = 10:11)
+```
+
+---
+class: center 
+## Output and Interpretation
+
+```{r}
+table_output(obj = logistic_fit, rows = 12:13)
+```
+
+---
+class: center 
+## Output and Interpretation
+
+```{r}
+table_output(obj = logistic_fit, rows = 14:15)
+```
+
+---
+class: center
+# Scoring
+
+$$\LARGE\sum_{i=1}^{63} (p_i * \mathds{1}_{correct prediction}  -p_i * \mathds{1}_{incorrect perdiction}) = score$$
+
+---
+
+# Performance
+
+---
+
+# Conclusion

### 2021-04-13 R/independent_modeling.R

@@ -13,48 +13,50 @@ options(tibble.print_min = 20)
 
 # logsitic fit ------------------------------------------------------------
 
-logistic_fit <- glm(win ~ x3fg +
-                      opposingx3fg +
-                      # field goal pct
-                      fg_percent +
-                      opposingfg_percent +
-                      # free throws
-                      ft_percent +
-                      opposingft_percent +
-                      # rebound per game
-                      rpg +
-                      opposingrpg +
-                      # steels
-                      st +
-                      opposingst +
-                      #turnover
-                      to +
-                      opposingto +
-                      # blocks
-                      opposingbkpg +
-                      bkpg,
-                    data = merged, family = "binomial")
-
-start <- Sys.time()
-
-logistic_predictions <- ids %>%
-  mutate(predicted_probs = map2_dbl(.x = ids$teamid,
-                                    .y = ids$otherid,
-                                    .f = logistic_predictor)) %>%
-   mutate(
-    predicted_winner = if_else(predicted_probs > .5, team, other_team),
-    predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
-    predicted_loser_probs = 1 - predicted_winner_probs,
-    prediction_points = predicted_winner_probs - predicted_loser_probs,
-    .after = game
-  )
-paste0("runtime: ", round(Sys.time() - start, 2))
-
-saveRDS(logistic_predictions, here::here("cache", "logistic_predictions.RDS"))
-
+# logistic_fit <- glm(win ~ x3fg +
+#                       opposingx3fg +
+#                       # field goal pct
+#                       fg_percent +
+#                       opposingfg_percent +
+#                       # free throws
+#                       ft_percent +
+#                       opposingft_percent +
+#                       # rebound per game
+#                       rpg +
+#                       opposingrpg +
+#                       # steels
+#                       st +
+#                       opposingst +
+#                       #turnover
+#                       to +
+#                       opposingto +
+#                       # blocks
+#                       opposingbkpg +
+#                       bkpg,
+#                     data = merged, family = "binomial")
+# 
+# 
+# start <- Sys.time()
+# 
+# logistic_predictions <- ids %>%
+#   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
+#                                     .y = ids$otherid,
[…]

### 2021-04-13 R/model_valadation.R

@@ -32,6 +32,7 @@ summarise(mean = mean(score, na.rm = TRUE),
             p = pt(t_stat, n() - 1, lower.tail = FALSE))
 
 
+
 # poisson scoring ---------------------------------------------------------
 
 posson_prediction  <- toscore(posson_prediction) %>% 
@@ -60,7 +61,8 @@ multinomeal_model_score  <- toscore(multinom_prediction) %>%
                     diff >=  -4 & diff <  4  ~ `-4:4`,
                     diff >=   4 & diff < 13  ~ `4:13`,
                     diff >= 13               ~ `>13`)) %>% 
-  select(`<-12`:`>13`, fifth_xtile, probs) %>%  # TODO: come up with a way to score this
-  mutate(score = if_else(probs > 0.25, (probs * 4), (1 - probs))) 
+  select(`<-12`:`>13`, fifth_xtile, probs) %>%
+  mutate(score = if_else(probs > 0.25, (probs * 4), (-1/4)*(1 - probs))) 
+multinomeal_model_score 
+sum(multinomeal_model_score$score, na.rm = T)
 
-multinomeal_model_score

### 2021-04-14 R/model_valadation.R
cached result

@@ -25,6 +25,9 @@ toscore <- function (df) {
 logistic_model_score <- toscore(logistic_predictions) %>% 
   mutate(score = if_else(predicted_winner == winner, predicted_winner_probs, -1*predicted_winner_probs)) 
 
+# saveRDS(logistic_model_score, here::here("cache", "logistic_model_score.RDS"))
+
+
 logistic_model_score %>% 
 summarise(mean = mean(score, na.rm = TRUE),
              sd = sd(score, na.rm = TRUE),

### 2021-04-15 slides/slides.Rmd
worked on slides

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(fig.width = 6, echo = FALSE, message = FALSE, warning = FALSE, comment = "", cache = F)
+knitr::opts_chunk$set(fig.width=15,
+                      fig.height=12, 
+                      echo = FALSE, 
+                      message = FALSE,
+                      warning = FALSE, 
+                      comment = "")
 
@@
 source(here::here("R", "fitting.R"))
+theme_set( theme_light(base_size = 25))
 ```
@@
 
-- Currently, the top 64 teams qualify
+- Now, its restricted to the top 64 NCAA Division I Teams
 
@@
 
-- maybe 2022...
-
---
-
 - When the turiment starts we form small groups and gamble on the outcomes. Usually by selecting winners and loser from the backet until we have a completely filled out turniment
@@
 
-- Nobody has ever made a perfect bracket there are $2^{63}$ possible outcomes.
+- Nobody has ever made a perfect bracket and there are $2^{63}$ possible outcomes.
 
@@
 
-class: inverse, center, middle
+class: inverse
 
-# Can we use basketball seasion data to predict this turniment outcomes?
+# Using basketball seasion data to predict this year's turniment outcomes requires us to:
+
+### Design predive models to make out of sample predictions
+
+--
+
+### Derrive a scoring metric
+
+--
+
+### Evaluate and tune the model
+
+--
+
+### We made several models, In the interstst of time we are going to discuss the logistic because its the most straight-forward
+
+---
+class: center
+
+```{r fig.width=15, fig.height=12}
+hists <- merged %>%
+  select(x3fg, fg_percent, rpg, st, to,  bkpg)
+
+histogram <- function(x, t){
+  ggplot(data = hists, aes(x)) +
+    geom_histogram(bins = 55, color = "white", fill = "black") +
+    theme_light(base_size = 25) +
+    labs(title = paste0("", t), x = "", y = "") +
+    NULL
+
+}
+
+a <- histogram(merged$x3fg,       t = "Three-Points Goals")
+b <- histogram(merged$fg_percent, t = "Percentage of Goals Scored")
+c <- histogram(merged$rpg,        t = "Rebounds Per Game")
+d <- histogram(merged$st,         t = "Steals Per Game")
+e <- histogram(merged$to,         t = "Turnovers Per Game")
+f <- histogram(merged$bkpg,       t = "Blocks Per game")
+# (a + b + c) / (d + e + f)
+(a + b) / (c + d) /(e + f)
+```
 
@@
 
-# Predictors
+# Real Quick Descriptive Statisticcs
 
@@
 
----
-class: center
+- Normal-ish distributions
 
-# Distribution of Predictors
+- No catastrophic skews or heavy tails
 
-```{r fig.width=15, fig.height=10}
-hists <- merged %>%
-  select(x3fg, fg_percent, rpg, st, to,  bkpg)
+--
 
-histogram <- function(x, t){
-  ggplot(data = hists, aes(x)) +
-    geom_histogram(bins = 50, color = "black", fill = "gray80") +
-    theme_light(base_size = 20) +
-    labs(title = paste0("", t), x = "", y = "") +
-    NULL
-
-}
-
-a <- histogram(merged$x3fg,       t = "Three-Points Goals")
-b <- histogram(merged$fg_percent, t = "Percentage of Goals Scored")
-c <- histogram(merged$rpg,        t = "Rebounds Per Game")
-d <- histogram(merged$st,         t = "Steals Per Game")
-e <- histogram(merged$to,         t = "Turnovers Per Game")
-f <- histogram(merged$bkpg,       t = "Blocks Per game")
-# (a + b + c) / (d + e + f)
-(a + b) / (c + d) /(e + f)
-```
+- Ideally we would have had steaks and turn-overs per game rather than season total. Not all played the same game
 
@@
 ---
+class: center
 
-# Poisson Model
+Summary
 
 ```{r}
-equatiomatic::extract_eq(model = fit_poisson,
-                         use_coefs = FALSE,
-                         show_distribution = TRUE,
-                         wrap = TRUE,
-                         terms_per_line = 2,
-                         intercept = "beta")
+row_name <- c("Intercept",
+  "Three Pointers", "O* Three Pointers", 
+          "Field Goal Percentage", "O* Field Goal Percentage",
+          "Free-throw percentage", "O* Free-throw percentage",
+          "Rebounds", "O* Rebounds",
+          "Steals", "O* Steals",
+          "Turnovers", "O* Turnovers",
+          "Blocks", "O* Blocks"
+          )
+
+cols <- c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")
+
+table_output <- function(obj, rows){
+  broom::tidy(obj) %>%
+  mutate(p.value = as.character(p.value), term = row_name) %>%
+  mutate_if(is.numeric, round, 3) %>%
+  mutate(p.value = as.numeric(p.value))  %>%
+  kable(col.names = cols) %>%
+  kable_styling(font_size = 15) %>%
+  row_spec(rows, color = "white", background = "#025cba") %>% 
+  add_footnote("Where O* is the Mertric associated with their opposing team", notation = "none")
+}
+  broom::tidy(logistic_fit) %>%
+  mutate(p.value = as.character(p.value), term = row_name) %>%
+  mutate_if(is.numeric, round, 3) %>%
+  mutate(p.value = as.numeric(p.value))  %>%
+  kable(col.names = cols) %>%
+  kable_styling(font_size = 15) %>%
+  # row_spec(rows, color = "white", background = "#025cba") %>% 
+  add_footnote("Where O* is the Mertric associated with their opposing team", 
+               notation = "none")
+
 ```
@@
 ---
-
-# Multinomeal Model
-
----
 class: center
-### Output and Interpretation
+Summary
 
 ```{r}
-table_output <- function(obj, rows){
-  broom::tidy(obj) %>% 
-  mutate(p.value = as.character(p.value)) %>%
-  mutate_if(is.numeric, round, 3) %>% 
-  mutate(p.value = as.numeric(p.value))  %>% 
-  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
-  kable_styling(font_size = 15) %>% 
-  row_spec(rows, color = "white", background = "#025cba")
-}
 
@@
 class: center
-## Output and Interpretation
+Summary
 
@@
 ---
-class: center
-# Scoring
+## Scoring
 
-$$\LARGE\sum_{i=1}^{63} (p_i * \mathds{1}_{correct prediction}  -p_i * \mathds{1}_{incorrect perdiction}) = score$$
+$$\sum_{i=1}^{n} [\mathbf {1} _{\{correct\ prediction\}} \hat{p_i} -\mathbf {1} _{incorrect\ perdiction}\hat{p_i} ] = Model\ Score$$
+
+--
+
+- Basketball games are often quite close and odds in betting markets are often close to 1:1.
+
+--
+
+- Our goal Goal is to weight correct prediction higher when it assigns a higher probability. Conversely a lower score to an incorrect prediction.
+
+--
+
+- $E(Model\ Score) = 0$ if our predictions are no better than randomly guessing.
 
@@
 
-# Performance
+## Scoring
+
+$$\sum_{i=1}^{n} [\mathbf {1} _{\{correct\ prediction\}} \hat{p_i} -\mathbf {1} _{incorrect\ perdiction}\hat{p_i} ] = Model\ Score$$
+
+Minimal Example
+
+```{r}
+col_fix <- function(x){
+  x <- str_replace(x, "_", " ")
+  x <- str_to_title(x)
+  x
+}
+
+logistic_model_score <- readRDS(here::here("cache", "logistic_model_score.RDS")) %>% 
+  drop_na() %>% 
+  head(10) %>% 
+  mutate_if(is.character, str_to_title) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(phat = if_else(predicted_winner == winner, 
+                          predicted_winner_probs, predicted_loser_probs)) %>% 
+  rename_all(col_fix) %>%
+  rename( `$\\hat{p}$` = Phat) %>% 
+  select(Game, `Predicted Winner`, `Winner`, `$\\hat{p}$`, Score)
+```
+
+```{r}
+logistic_model_score %>% 
+  kable(align = "r") %>% 
+  kable_styling(font_size = 15) 
+```
 
@@
 
+## Scoring
+
+$$\sum_{i=1}^{n} [\mathbf {1} _{\{correct\ prediction\}} \hat{p_i} -\mathbf {1} _{incorrect\ perdiction}\hat{p_i} ] = Model\ Score$$
+
+Correct Predictions
+
+```{r}
+logistic_model_score %>% 
+  kable(align = "r") %>% 
+  kable_styling(font_size = 15) %>%
+  column_spec(c(1:5), color =  ifelse(logistic_model_score$Score > 0, "white", "black"), 
+              background = ifelse(logistic_model_score$Score > 0, "#025cba", "#f0f0f0"))
+```
+
+---
+
+## Scoring
+
+$$\sum_{i=1}^{n} [\mathbf {1} _{\{correct\ prediction\}} \hat{p_i} -\mathbf {1} _{incorrect\ perdiction}\hat{p_i} ] = Model\ Score$$
+
+Score
+
+```{r}
+logistic_model_score %>% 
+  kable(align = "r") %>% 
+  kable_styling(font_size = 15) %>% 
+  column_spec(c(5), color =  ifelse(logistic_model_score$Score > 0, "white", "black"), 
+              background = ifelse(logistic_model_score$Score > 0, "#025cba", "#ff3700"))
+```
+
+```{r fig.align='right'}
+paste0("Model Score = ", sum(as.numeric(logistic_model_score$Score))) %>% 
+  kable(col.names = "$\\sum_{i}Score$") %>% 
+  kable_styling(font_size = 15) %>% 
+  row_spec(1, color = "white", background = "#025cba")
+```
+
+---
+
+```{r}
+perform <- readRDS(here::here("cache", "logistic_model_score.RDS")) %>% 
+  mutate(outcome = if_else(score > 0, "Correct", "Incorrect")) %>% 
+  drop_na() 
+```
+
+```{r fig.height=10}
+
+a <- perform %>%
+  ggplot(aes(x = abs(score), fill = outcome)) + 
+   geom_histogram(alpha=0.7, position="identity", bins = 10) + 
+  labs(title = "Histogram of Scores", x = "Absolute Value of Score") +
+  scale_fill_manual(values = c("#025cba", "#ff3700"))
+
+b <- perform %>% 
+  ggplot(aes(as.factor(outcome), fill = outcome)) + 
+  geom_bar(alpha = 0.7) + 
+  labs(title = "", x = "") + 
+  scale_fill_manual(values = c("#025cba", "#ff3700")) +
+  geom_text(size=10, color = "grey80", stat = 'count', aes(label =..count.., hjust = 1.5)) + 
+  coord_flip()
+
+a/b
+```
+
+```{r}
+t2 <- perform %>% 
+  summarise(totalscore = sum(score),
+            se = sd(score), 
+            t = totalscore/se,
+              p = pt(t, nrow(perform)-1, lower.tail = F)) %>% 
+  t()
+
+score <- perform %>% 
+  group_by(outcome) %>% 
+  summarise(totalscore = sum(score)) %>% 
+  mutate(totalscore = round(totalscore, 2))
+
+t1 <- rbind(score, c("Total", sum(score$totalscore))) 
+```
+
+```{r}
+knitr::kable(list(t1, t2)) %>% 
+  kable_styling(font_size = 10)
+
+```
+
+---
+class: inverse
+
 # Conclusion
+
+### Logistic Regression performs well. Given measured statists.
+
+### However out model is independent of the pervious game which would aplify shortcommings.

### 2021-04-15 R/fitting.R
adding valadation

@@ -19,7 +19,10 @@ logistic_fit <- glm(win ~ x3fg +
                        opposingbkpg +
                        bkpg,
                      data = merged, family = "binomial")
- 
+
+logistic_fit_null <- glm(win ~ 1,
+                    data = merged, family = "binomial") 
+
  fit_poisson <- glm(team_score ~ x3fg +
               opposingx3fg +
               # field goal pct

### 2021-04-15 R/model_valadation.R

@@ -38,10 +38,16 @@ summarise(mean = mean(score, na.rm = TRUE),
 
 # poisson scoring ---------------------------------------------------------
 
-posson_prediction  <- toscore(posson_prediction) %>% 
+posson_model_score <- toscore(posson_prediction) %>% 
   mutate(predicted_difference = predicted_score - other_predicted_score,
          actual_difference = teamscore - otherteamscore,
-         score = predicted_difference - actual_difference) 
+         score = predicted_difference - actual_difference) %>% 
+  mutate(predicted_winner = if_else(predicted_difference > 0, team, other_team),
+         outcome = if_else(predicted_winner == winner, "Correct", "Incorrect")) 
+
+
+# saveRDS(posson_model_score, here::here("cache", "posson_model_score.RDS"))
+
 
 posson_model_score %>% 
   summarise(mean = mean(score, na.rm = TRUE),
@@ -64,8 +70,19 @@ multinomeal_model_score  <- toscore(multinom_prediction) %>%
                     diff >=  -4 & diff <  4  ~ `-4:4`,
                     diff >=   4 & diff < 13  ~ `4:13`,
                     diff >= 13               ~ `>13`)) %>% 
-  select(`<-12`:`>13`, fifth_xtile, probs) %>%
+  # select(`<-12`:`>13`, fifth_xtile, probs) %>%
   mutate(score = if_else(probs > 0.25, (probs * 4), (-1/4)*(1 - probs))) 
 multinomeal_model_score 
 sum(multinomeal_model_score$score, na.rm = T)
+
+colnames(multinomeal_model_score)
+  
+# saveRDS(logistic_model_score, here::here("cache", "logistic_model_score.RDS"))
+
+multinomeal_model_score <- multinomeal_model_score %>% 
+  mutate(predicted_winner = if_else(`<-12` + `-12:-4` > `other4:13` + `>13`, 
+                                  team, other_team),
+         outcome = if_else(predicted_winner == winner, "Correct", "Incorrect"))
+
+# saveRDS(multinomeal_model_score, here::here("cache", "multinomeal_model_score.RDS"))
 

### 2021-04-15 slides/slides.Rmd
add finished slides

@@
 
-# Backround
+# Backround on March Madness Tournament
 
@@
 
-- First tournament was in 1939 only 8 teams
+- First hosted in 1939 with 8 participating teams.
 
@@
 
-- Now, its restricted to the top 64 NCAA Division I Teams
+- Now, restricted to the top 64 qualifying NCAA division I teams.
 
@@
 
-- Only one team has ever gone undefeated thought a season and the tournament
+- Currently 16.9 million viewers.
 
@@
 
-- 16.9 Million Viewers
+- Michigan State won in 1979 and 2000.
 
@@
 
-- MSU won in 1979 and 2000
+- Offices, organizations, and friend groups make and submit a total of 40 Million predictions each year.
 
@@
 
-- When the turiment starts we form small groups and gamble on the outcomes. Usually by selecting winners and loser from the backet until we have a completely filled out turniment
+- At the beginning small groups form and gamble on the outcomes by selecting winners and losers.
 
@@
 
-- Nobody has ever made a perfect bracket and there are $2^{63}$ possible outcomes.
+- Nobody has ever made a perfect bracket prediction and there are $2^{63}$ possible outcomes.
 
@@
 ---
+
+![](images/results.jpg)
+
+---
+
 class: center, middle
@@
 
-# Using basketball seasion data to predict this year's turniment outcomes requires us to:
-
-### Design predive models to make out of sample predictions
+## We use basketball season game scores and statistics to predict this year's tournament outcome.
 
@@
 
-### Derrive a scoring metric
+## We need to derive a scoring metric.
 
@@
 
-### Evaluate and tune the model
+## Evaluate and tune the model.
 
@@
 
-### We made several models, In the interstst of time we are going to discuss the logistic because its the most straight-forward
+## Compare the model's performance.
+
+--
+
+## We made several models, in the interest of time we are going to discuss the logistic because its the most straight-forward.
 
@@
   ggplot(data = hists, aes(x)) +
-    geom_histogram(bins = 55, color = "white", fill = "black") +
+    geom_histogram(bins = 55, color = "white", fill = "#025cba") +
     theme_light(base_size = 25) +
@@
     NULL
-
 }
@@
 e <- histogram(merged$to,         t = "Turnovers Per Game")
-f <- histogram(merged$bkpg,       t = "Blocks Per game")
+f <- histogram(merged$bkpg,       t = "Blocks Per Game")
 # (a + b + c) / (d + e + f)
@@
 
-- Normal-ish distributions
+- Normal-ish distributions.
 
-- No catastrophic skews or heavy tails
+- No catastrophic skews or heavy tails.
 
@@
 
-- Ideally we would have had steaks and turn-overs per game rather than season total. Not all played the same game
+- Ideally we would have had steaks and turn-overs per game rather than season total. Not all played the same game.
 
@@
 
+lrt_stat <- 2*(as.numeric(logLik(logistic_fit)) -
+  as.numeric(logLik(logistic_fit_null)))
+
+pval_lrt <- pchisq(lrt_stat, df = 1, lower.tail = FALSE)
+
+dev <- deviance(logistic_fit)
+
+pval_dev <- pchisq(dev, df.residual(logistic_fit), 
+       lower.tail = FALSE)
+
+tests <- cbind(
+LRT = c(lrt_stat, pval_lrt),
+DEViance = c(dev, pval_dev)
+)
+
+rownames(tests) <- c("Stat", "P Value")
+lrt_dev <- kable(tests) %>% 
+  kable_styling(font_size = 10)
+lrt_dev
 ```
@@
 ```{r}
-
 table_output(obj = logistic_fit, rows = 2:3)
+lrt_dev
 ```
@@
 table_output(obj = logistic_fit, rows = 4:5)
+lrt_dev
 ```
@@
 table_output(obj = logistic_fit, rows = 6:7)
+lrt_dev
 ```
@@
 table_output(obj = logistic_fit, rows = 8:9)
+lrt_dev
 ```
@@
 table_output(obj = logistic_fit, rows = 10:11)
+lrt_dev
 ```
@@
 table_output(obj = logistic_fit, rows = 12:13)
+lrt_dev
 ```
@@
 table_output(obj = logistic_fit, rows = 14:15)
+lrt_dev
 ```
@@
 
-- Our goal Goal is to weight correct prediction higher when it assigns a higher probability. Conversely a lower score to an incorrect prediction.
+- Our goal is to weight correct prediction higher when it assigns a higher probability. Conversely, a lower score to an incorrect prediction.
 
@@
 
-- $E(Model\ Score) = 0$ if our predictions are no better than randomly guessing.
+- If our predictions are no better than randomly guessing $E(Model\ Score) = 0$.
 
@@
 
+Incorrect Predictions
+
+```{r}
+logistic_model_score %>% 
+  kable(align = "r") %>% 
+  kable_styling(font_size = 15) %>% 
+  column_spec(c(1:5), color =  ifelse(logistic_model_score$Score > 0, "black", "white"), 
+              background = ifelse(logistic_model_score$Score > 0, "white", "#ff3700"))
+```
+
+---
+
+## Scoring
+
+$$\sum_{i=1}^{n} [\mathbf {1} _{\{correct\ prediction\}} \hat{p_i} -\mathbf {1} _{incorrect\ perdiction}\hat{p_i} ] = Model\ Score$$
+
 Score
@@
   kable(align = "r") %>% 
-  kable_styling(font_size = 15) %>% 
-  column_spec(c(5), color =  ifelse(logistic_model_score$Score > 0, "white", "black"), 
-              background = ifelse(logistic_model_score$Score > 0, "#025cba", "#ff3700"))
+  kable_styling(font_size = 15)
 ```
@@
   scale_fill_manual(values = c("#025cba", "#ff3700")) +
-  geom_text(size=10, color = "grey80", stat = 'count', aes(label =..count.., hjust = 1.5)) + 
+  geom_text(size=10, color = "grey80", stat = 'count',
+            aes(label =..count.., hjust = 1.5)) + 
   coord_flip()
@@
               p = pt(t, nrow(perform)-1, lower.tail = F)) %>% 
-  t()
+  t() %>% 
+  tibble() %>% 
+  mutate_if(is.numeric, round, 2)
+
+rownames(t2) <- c("Total Score", "Standard Error", "T-Statistic", "P-value")
+colnames(t2) <- NULL
 
@@
 t1 <- rbind(score, c("Total", sum(score$totalscore))) 
+colnames(t1) <- c("Result", "Score")
+
 ```
@@
   kable_styling(font_size = 10)
+```
+---
+class: center
+Model Performance
 
+```{r}
+a <- perform %>% 
+  ggplot(aes(as.factor(outcome), fill = outcome)) + 
+  geom_bar(alpha = 0.7) + 
+  labs(title = "#2 Logistic Model", x = "", y = "") + 
+  scale_fill_manual(values = c("#025cba", "#ff3700")) +
+  geom_text(size=10, color = "grey80", stat = 'count',
+            aes(label =..count.., hjust = 1.5)) + 
+  coord_flip()
+```
+
+```{r}
+posson_model_score <- readRDS(here::here("cache","posson_model_score.RDS"))
+
+b <- posson_model_score %>% 
+  drop_na() %>% 
+  ggplot(aes(as.factor(outcome), fill = outcome)) + 
+  geom_bar(alpha = 0.7) + 
+  labs(title = "#1 Poisson Model", x = "", y = "") + 
+  scale_fill_manual(values = c("#025cba", "#ff3700")) +
+  geom_text(size=10, color = "grey80", stat = 'count',
+            aes(label =..count.., hjust = 1.5)) + 
+  coord_flip()
+```
+
+```{r}
+multinomeal_model_score <- readRDS(here::here("cache", "multinomeal_model_score.RDS"))
+
+c <- multinomeal_model_score %>% 
+  drop_na() %>% 
+  ggplot(aes(as.factor(outcome), fill = outcome)) + 
+  geom_bar(alpha = 0.7) + 
+  labs(title = "#3 Multinomeal Model", x = "", y = "") + 
+  scale_fill_manual(values = c("#025cba", "#ff3700")) +
+  geom_text(size=10, color = "grey80", stat = 'count',
+            aes(label =..count.., hjust = 1.5)) + 
+  coord_flip()
+```
+
+```{r fig.height=11}
+b/a/c
 ```
@@
 
-### Logistic Regression performs well. Given measured statists.
+### Both Logistic and Poisson Regression perform well.
 
-### However out model is independent of the pervious game which would aplify shortcommings.
+--
+
+### However, we are scoring the model based on independent predictions. While real-world predictions are made dependent on the previous round.
+
+--
+
+### These predictions are not wildly different than betting markets and seeds because they use similar statistics to derive their predictions.

### 2021-04-16 slides/slides.Rmd
fixed typo

@@
 source(here::here("R", "fitting.R"))
-theme_set( theme_light(base_size = 25))
+theme_set(theme_light(base_size = 25))
 ```

### 2021-04-16 R/model_valadation.R
fix error

@@ -80,7 +80,7 @@ colnames(multinomeal_model_score)
 # saveRDS(logistic_model_score, here::here("cache", "logistic_model_score.RDS"))
 
 multinomeal_model_score <- multinomeal_model_score %>% 
-  mutate(predicted_winner = if_else(`<-12` + `-12:-4` > `other4:13` + `>13`, 
+  mutate(predicted_winner = if_else(`<-12` + `-12:-4` < `other4:13` + `>13`, 
                                   team, other_team),
          outcome = if_else(predicted_winner == winner, "Correct", "Incorrect"))
 

### 2021-04-16 paper/paper.Rmd
worked on paper

@@
 
-__Introduction__
+# Introduction
+<!-- (a) introduces the background -->
+<!-- (b) states the scientific questions you aim to answer -->
+<!-- (c) describes the data set that you will use for the analysis (d) summarizes your major findings. -->
 
@@
 Casual fans and enthusiasts submit predictions gambling small sums of money on the outcomes.
-This study will determine if which GLM performs best in predicting the results.
+This study will compare which GLM performs best in predicting the turniment results.
 Our data comes from three different sources Kaggle, NCAA, and the tournament results.
@@
 The NCAA provides team-level statistics^[CITE THIS: https://stats.ncaa.org/rankings/change_sport_year_div] for each team.
-We will use those two datasets to train a model that uses team-level statistics to predict basketball game outcomes.
+We will use those two datasets to train a model using team-level statistics then use that model to predict basketball game outcomes.
 We will test our model on the actual tournament results after it is concluded on April 5th.
@@
 
-Our response is the outcome *win/loss* as a function of team statistics.
-Win or loss coded 0 or 1.
-Then we will look at both team's statistics and determine which is more important.
-In our dataset we have: *field goal percentage* which is the rate attempt to score vs actually scoring.
+Our response terms is the outcome of the game *win or loss* with a 0 possibility of a tie.
+Our aim is to derive a function that will accurately predict this using team-level statistics.
+In our dataset we have many predictors however, we only selected the predictors that were not co-linear.
+First, *field goal percentage* which is the rate attempt to score vs actually scoring.
 *Free-throw percentage*,  is the rate that the team has the opportunity to take an unguarded shot.
@@
 
-We will use three regression methods, logistic, Poisson, and multinominal, to compare the results to determine which is most accurate.
-Logistic regression most naturally fits this problem because games do not tie so there is a 50/50 split of wins and losses.
-Also, we will try Poisson regression to predict the number of points scored by each team.
-Since there are many close basketball games we will use multinomial regression to predict if the game score is going to be a large difference one way, small difference, or a large difference the other way.
+March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
+Therefore, making accurate predictions will be depending on predicting the previous round correctly.
+We did make predictions in that fashion, however our models will be compared by taking independent predictions.
+Meaning, we will filter our data down to using only the 64 teams that qualify, predict every single possible combination of games ${64\choose 2} = 2016$ then use the 63 games played as a sample from the population of all possible games.
+Then use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
+Then we use the results to determine which regression method performs the best.
 
-We Found....
+We will use three regression methods, logistic, poisson, and multinominal, to generate a prediction models, to compare the results, and make a determination as to which is best suited to this problem.
+Logistic regression most naturally suits this suits this problem because games do not tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
+This problem could also be suited to poison regression because the number of points scored is poisson distrubuted.
+After fitting a poisson model we will predict the amount of points Team A will score, then predict the number of points Team B will score then take whichever team was predicted to score more and record that as the prediction.
 
-<!-- (a) introduces the background -->
-<!-- (b) states the scientific questions you aim to answer -->
-<!-- (c) describes the data set that you will use for the analysis (d) summarizes your major findings. -->
-
-__Exploratory Data Analysis__
+# Exploratory Data Analysis
 

### 2021-04-17 paper/paper.Rmd
worked on paper first draft complete

@@
 shorttitle        : "Using GLMs to Predict Basketball Games"
-
 author: 
@@
   - name          : "Sabrina Ball"
-
 affiliation:
@@
 wordcount         : "X"
-
 bibliography      : ["r-references.bib"]
-
 floatsintext      : yes
@@
 draft             : no
-
 documentclass     : "apa6"
@@
 urlcolor: blue
+editor_options: 
+  chunk_output_type: inline
+language:
+  label:
+    fig: "Abbildung "
 ---
@@
 
-merged <- readr::read_csv("data/merged.csv")
-team_stats <- readr::read_csv("data/team_stats.csv")
+merged <- readr::read_csv("https://raw.githubusercontent.com/despresj/March-Madness/main/data/merged.csv")
+team_stats <- readr::read_csv("https://raw.githubusercontent.com/despresj/March-Madness/main/data/team_stats.csv")
 ```
 
-# Introduction
-<!-- (a) introduces the background -->
-<!-- (b) states the scientific questions you aim to answer -->
-<!-- (c) describes the data set that you will use for the analysis (d) summarizes your major findings. -->
+<!-- All the files for this project is in a github repo found at https://github.com/despresj/March-Madness -->
+## Introduction
 
-Every year top university basketball teams compete in the March Madness tournament.
+Every year top division I basketball teams compete in the annual March Madness tournament.
 Casual fans and enthusiasts submit predictions gambling small sums of money on the outcomes.
-This study will compare which GLM performs best in predicting the turniment results.
-Our data comes from three different sources Kaggle, NCAA, and the tournament results.
-Kagggle provides a comprehensive dataset^[CITE THIS: https://plexkits.com/march-madness-bracket/] for all NCAA in season (non-turniment) basektball games from 2001 to 2020.
-The NCAA provides team-level statistics^[CITE THIS: https://stats.ncaa.org/rankings/change_sport_year_div] for each team.
-We will use those two datasets to train a model using team-level statistics then use that model to predict basketball game outcomes.
-We will test our model on the actual tournament results after it is concluded on April 5th.
-The objective is to predict the individual game outcomes as accurately as possible.
-
+This study uses logistic, poisson, and multinominal regression models in R [@R-base] to predict the March Madness tournament outcomes.
+Our primary objective is to determine which GLM is the most accurate when predicting the results.
+The data we use come from three different sources Kaggle [@R-Kaggle], NCAA [@R-NCAA], and the tournament results [@R-NCAA].
+Kaggle [@R-Kaggle] provides a comprehensive dataset including all NCAA in-season (non-turniment) basketball games from 2001 to 2020.
+The NCAA [@R-NCAA] provides team-level statistics for each team.
+We filter, clean, and combine these data using the tidyverse package [@R-tidyverse].
+Then use the combination of these datasets to fit our models.
+The objective is to predict the individual game outcomes as accurately as possible and determine which GLM is the most accurate.
 Our response terms is the outcome of the game *win or loss* with a 0 possibility of a tie.
@@
 *Blocks*, The number of times the team was able to block a shot made by the opposing team.
-
 March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
@@
 Then we use the results to determine which regression method performs the best.
-
 We will use three regression methods, logistic, poisson, and multinominal, to generate a prediction models, to compare the results, and make a determination as to which is best suited to this problem.
@@
 
-# Exploratory Data Analysis
+We find that these models have an accurate between 59% and 65% which is well over 50%.
+Therefore, GLM's are helpful in predicting March madness tournament outcomes.
+We find that the team-level statistics recorded by the NCAA are highly significant and helpful for predicting the tournament outcomes however, come up short of producing truely fantastic results.
+Also, we find that when predicting the probability of a win or a loss we find that the coefficients are symmetric because of the nature of the zero-sum-contest.
 
-taking a look at our variables we see that they all have a reasionably normal distribution.
+From these results we see that logistic and posisson model was the best at predicting wins and losses.
+All three models were well over 50% accurate, so this the research goal of having GLM's predict March Madness Outcomes.
+Also, we confirmed the suspicion that basketball statistics are symmetrical in relation to the probability of winning.
+Yet are not when we are predicting the score.
 
-```{r}
+## Exploratory Data Analysis
+
+```{r fig.height=2}
 hists <- merged %>% 
@@
 
-histogram <- function(x, t){
-  ggplot(data = hists, aes(x)) + 
-    geom_histogram(bins = 30, color = "black", fill = "gray98") + 
-    theme_light(base_size = 5) + 
-    labs(title = paste0("Distribution of ", t), x = t, y = "") + 
-    NULL
-
-}
-
-a <- histogram(merged$x3fg,       t = "Three-Points goals Scored")
-b <- histogram(merged$fg_percent, t = "Percentage of Field Goals Scored")
-c <- histogram(merged$rpg,        t = "Rebounds Per Game")
-d <- histogram(merged$st,         t = "Steels Per Game")
-e <- histogram(merged$to,         t = "Turn Overs Per Game")
-f <- histogram(merged$bkpg,       t = "Blocks Per game")
-(a + b + c) / (d + e + f)
-```
-
-```{r}
-stat <- function(x, df = hists, rounding_digits = 2){
+stat <- function(x, df = hists, rounding_digits = 2) {
     x <- enquo(x)
     df %>%
-      summarise(
-                       Mean  = mean( !! x),
-                     Median  = median( !! x),
-                         Std = sd  ( !! x),
-                         Min = min ( !! x),
-                         Max = max ( !! x),
-                       Range = max(!! x) - min(!! x) ) %>%
+      summarise(Mean = mean(!!x),
+              Median = median(!!x),
+                 Std = sd(!!x),
+                 Min = min(!!x),
+                 Max = max(!!x),
+               Range = max(!!x) - min(!!x)) %>%
       mutate_if(is.numeric, round, rounding_digits)
@@
 "Steels Per Game",
-"Turn Overs Per Game",
+"Turn Overs",
 "Blocks Per game")
@@
 
-<!-- • Perform exploratory data analysis (with summary statistics or exploratory graphs) that reveals some basic and general information of the data set. -->
-__Description__  
-To gather these data, we used the tidyr, [@R-tidyr] dplyr, [@R-dplyr] and purrr [@R-purrr] package to read-in clean, and prepare our data for analysis.
-Since this is a predictive model we used all the data we had at our disposal that was not colinear.
+```{r }
+histogram <- function(x, t) {
+  ggplot(data = hists, aes(x)) + 
+    geom_histogram(bins = 50, color = "black", fill = "white") + 
+    theme_light(base_size = 7) + 
+    labs(title = paste0(t), x = "", y = "") + 
+    NULL
+}
 
-<!-- • Provide a description and explanation of the steps that you have taken to obtain the best model (or models). -->
-__Results__  
-<!-- • Give the results of relevant parameter estimation and statistical inference under your best model (or models).
-Discuss and interpret important features of your best model (or models), and state your conclusions in the context of the scientific questions that you have been focused on.
--->
-__Diagnostic Methods__  
-<!-- • Check the goodness of fit of your model (or models), including a description of the model diagnostic methods and test procedures you used, and what the results were. -->
-__Conculsion__  
-<!-- • Conclude the report with a discussion section that -->
-<!-- (a) recaps your scientific findings -->
-<!-- (b) discusses possible limitations and improvements of your methods -->
-<!-- (c) outlines additional questions that are worth studying as a future work. -->
+a <- histogram(merged$x3fg, t = "Three-Points Goals")
+b <- histogram(merged$fg_percent, t = "Percentage of Field Goals Scored")
+c <- histogram(merged$rpg, t = "Rebounds Per Game")
+d <- histogram(merged$st, t = "Steels Per Game")
+e <- histogram(merged$to, t = "Turn Overs Per Game")
+f <- histogram(merged$bkpg, t = "Blocks Per game")
+(a + b + c + d + e + f)
+```
+
+As mentioned above, there are substantially more variables in the dataset we used, however we omitted the ones that were co-linear.
+When making a decision we selected the variable using deviance tests and likelihood ratios.
+For our modeling and prediction we are going to use these statistics associated with both teams.
+Basic descriptive statistics can be found in Table 1.
+We notice that the mean of the predictor is relatively close to the median.
+The standard deviations can be high, but are not wildly so and the range is reasionable for the data we have.
+In the histograms below we remark the predictors roughly normal distribution with no substantial skew, heavy tails, or slow decay.
+
+Now we turn our attention to the outcomes.
+The logistic model is straight forward all we needed to do was code a win to be 1 and a loss to be 0 then fit the model and make a prediction.
+The possion model will work if we to taking score as a count.
+This is not perfectly suited to poisson because a winning team's score has a mean of `r mean(merged$w_score) %>% round(2)` and a variance of `r var(merged$w_score) %>% round(2)`.
+Regardless we will see how it performs, when we will predict the score for Team A and predict the score for Team B, then take who ever has a higher predictions.
+
+```{r warning=FALSE}
+w <- merged %>% 
+  ggplot(aes(x = w_score)) + 
+  geom_histogram(bins = 50, color = "black", fill = "white") +
+  geom_vline(xintercept = mean(merged$w_score), color = "red", size = 0.25) + 
+   labs(title = "Winning Team's Score", x = "", y = "",
+       subtitle = paste0("Mean = ", mean(merged$w_score) %>% round(2),
+                         " Variance = ", var(merged$w_score) %>% round(2)))
+
+l <- merged %>% 
+  ggplot(aes(x = l_score)) + 
+  geom_histogram(bins = 50, color = "black", fill = "white")  +
+  geom_vline(xintercept = mean(merged$l_score), color = "red", size = 0.25) + 
+  labs(title = "Losing Team's Score", x = "", y = "",
+       subtitle = paste0("Mean = ", mean(merged$l_score) %>% round(2), 
+                        " Variance = ", var(merged$l_score) %>% round(2)))
+
+d <-  merged %>% 
+  ggplot(aes(x = diff)) + 
+  geom_histogram(bins = 100, color = "black", fill = "white") + 
+  geom_vline(xintercept = c(-12, -4, 4, 12), color = "red", size = 0.25) +
+  labs(title = "Differences in Score", x = "", y = "") + 
+  scale_x_continuous(limits = c(-50, 50), 
+                     breaks = c(-28, -20, -12, -4, 4, 12, 20, 28))
+
+(w + l) / d
+```
+The multinomeal case is not obviously suited to this problem however we can adapt it.
+We will take a look at the difference in score between winner and loser.
+This is a common thing in betting markets to gamble on.
+Therefore we mapped the difference in score into 5 categories using the 0.2, 0.4, 0.6, and 0.8 quantifies.
+First, the team losing with by more than 12.
+Second, the team losing by between 12 and 4 points.
+Third, a very close game between losing by 4 and winning by 4 (This is fairly common as there is really good competition between teams).
+Fourth, winning by more than 4 and less than 12.
+In the fifth category, winning by more than 12.
+We will use the probabilities to predict a winner by suming the proabilities of losing by more than 12 and losing by between 12 and four and summing the probability of winning by more than 12 and winning by between four and 12.
+Then taking whichever is larger as a prediction.
+
+## Description
+
+To make an accurate compartment, we use the same formula for the three models with only the dependent variable being different.
+Win or loss for our logistic, the amount Team A scores for the poisson and the difference in score for the multinomeal.
+Now we tried normalizing the data where we subtracted the mean and devided by standard deviation for all out predictors.
+However, that really did not have that much effect.
+Also we individually tested each predictor using likelyhood ratio tests to add each term.
+These all came significant.
+Also when we were adding terms we found many were co-linear, therefore when we found co-linear predictors we omitted the one with a smaller likelihood ratio.
+
+```{r, out.width="100%", fig.cap = ""}
+knitr::include_graphics("/Users/josephdespres/Documents/MSU/STT-864_Statistical_Methods_II/March-Madness/paper/paper_files/equ.png")
+```
+
+## Results
+
+In the Table 2 made manually with the kableExtra package [@R-kableExtra], you can see that we have the results for poisson, logistic, and multinomeal models.
+We have a sample of over 35,000 and all of the coeffecients are highly significant.
+This is to be expected because these are the statistics that the NCAA records and ones they have determined to being useful to measure a team's performance.
+The aim of this study is to compare predictive models so we will not cover it exhaustively or include individual z-statistics and p-values.
+First, notice is that in the case of the logistic and multinomeal models we see that when comparing a factors that affect a team's probability of winning we see the associated coefficient is quite similar for the factor capturing the opposing teams.
+That is because basketball is a zero-sum-game, and anything good for team A is proportionately bad for team B.
+Note this is not true of the  multinomeal and poisson models.
+The reasion that is is because the winner picked by thoes models is a function of the score.
+Therefore, thoes models are looking at coeffecients that are oging to increase the scores.
+Take three pointers in the poisson model, for instance, a team where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
+More points scored is not deterministic of winning, however, it is an indicator.
+
+```{r models}
+predictors <- "x3fg + opposingx3fg + # field goal pct
+              fg_percent + opposingfg_percent + # free throws
+              ft_percent + opposingft_percent +# rebound per game
+              rpg + opposingrpg + # steels
+              st + opposingst + #turnover
+              to +  opposingto + # blocks
+              opposingbkpg + bkpg"
+
+logistic_fit <- glm(paste0("win ~ ", predictors), data = merged, family = "binomial")
+fit_poisson <- glm(paste0("team_score ~ ", predictors), data = merged, family = poisson(link = "log"))
+fit_multinom <- VGAM::vglm(paste0("x_tile ~ ", predictors), family = VGAM::cumulative(parallel=TRUE), data = merged)
+```
+
+```{r}
+row_name <- c("Three Pointers", "O* Three Pointers", "Field Goals", "O* Field Goals",
+          "Free-throws", "O* Free-throws", "Rebounds", "O* Rebounds", "Steals", "O* Steals",
+          "Turnovers", "O* Turnovers", "Blocks", "O* Blocks")
+
+binder <- function(obj){
+  obj %>% tidy() %>% filter(term != "(Intercept)")
+}
+
+multi_terms <- c('<-12', '-12:-4', '-4:4', '4:13', row_name)
+
+logistic_pos <- binder(logistic_fit) %>% 
+  rename_all(~paste0("L", .x)) %>% 
+  bind_cols(binder(fit_poisson) %>% rename_all(~paste0("p", .x))) %>% 
+  transmute(`Logistic Model` = Lestimate, `Poisson Model` = pestimate)
+
+tibble(`Logistic Model` = rep(NA, 4),`Poisson Model` = rep(NA, 4)) %>% 
+  bind_rows(logistic_pos) %>% 
+  bind_cols(tibble(`Multinomial Model` = fit_multinom@coefficients), 
+            tibble(Terms = multi_terms)) %>% 
+  select(Terms, `Multinomial Model`, `Logistic Model`:`Poisson Model`) %>% 
+  mutate_if(is.numeric, round, 5) %>% 
+  mutate_at(vars(Terms:`Poisson Model`), replace_na, '.') %>% 
+  kable(format = "latex", booktabs = TRUE,
+  escape = FALSE,longtable = TRUE, caption = "Regression Output") %>% 
+  footnote(number = c("Sample size 35248 games.",
+                      "O* is the term asscoiated with the opposing team.",
+                      "All the above terms are significant at P < 0.01.",
+                      "Multinomeal Intercepts are differences in predicted score."))
+
+```
+
+## Goodness of Fit
+
+Now that we have fitted the model, lets take a look at how well it performed.
+Using the tidyr [@-tidyr] package, we make every combination of teams and associated statistics in one dataframe.
+Then wrote custom functions that would take two of the 64 teams as inputs and output the probabilities of winning, predicted amounts of points scored, or probabilities of the score resulting in one of the above mentioned multinomeal categories.
+After that we used the purrr [@R-purrr] package to iterate over all possible games that could be played.
+From there, wrote and ran a webscraping script  to obtain a dataframe of the results.
+Then counted the games each model predicted correctly and incorrectly.
+In this case our population is all  ${64\choose 2}$ possible games in this turniment and a sample of the games that were played.
+We are not assuming that we have a random sample from the whole population as these are the best teams in the league.
+Also, there is a selection bias twards games that were played.
+Therefore, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams only are represented once.
+Therefore, we do not claim that these results will hold for games in general.
+However, for the purpose of comparing models for this specific turniment this is approprite.
+
+```{r}
+bar <- function (obj, title, sub) {
+  obj %>% 
+  drop_na() %>% 
+  ggplot(aes(as.factor(outcome), fill = outcome)) + 
+  geom_bar(alpha = 0.7) + 
+  labs(title = title, subtitle = paste("Accuracy:", sub, "%"), x = "", y = "") + 
+  scale_fill_manual(name = "", values = c("grey20", "grey80"))  +
+  geom_text(size=5, color = "black", stat = 'count',
+            aes(label =..count.., hjust = 1.5)) + 
+  coord_flip()
+}
+
+multinomeal_model_score <- readRDS(here::here("cache", "multinomeal_model_score.RDS"))
+posson_model_score <- readRDS(here::here("cache","posson_model_score.RDS"))
+logistic_model_score <- readRDS(here::here("cache", "logistic_model_score.RDS")) %>% 
+  mutate(outcome = if_else(winner == predicted_winner, "Correct", "Incorrect"))
+
+a <- bar(obj = posson_model_score, title = "#1 Poisson Model", sub = "64.5")
+b <- bar(obj = logistic_model_score, title = "#2 Logistic Model", sub = "63.1")
+c <- bar(obj = multinomeal_model_score , title = "#3 Multinomeal Model", sub = "59.6")
+
+a/b/c
+```
+
+```{r}
+# plot(residuals(fit_poisson))
+# 
+# plot(residuals(logistic_fit))
+```
+## Conculsion
+
+From these results we see that logistic and posisson model was the best at predicting wins and losses.
+All three models were well over 50% accurate, so this the research goal of having GLM's predict March Madness Outcomes.
+Also, we confirmed the suspicion that basketball statistics are symmetrical in relation to the probability of winning.
+Yet are not when we are predicting the score.
+
+The limits of this study have a lot to do with data limitations.
+Ideally, we would have more tournaments playing with lesser skilled teams.
+would be much more robust if the tournament was much larger.
+We had a lot of data points over 35,000 which I think was more than sufficient, however, we lacking non co-linear covariates.
+We only had 7 predictors.
+An accuracy of 64.5% is not superb considering betting markets take these exact factors into account.
+
+Future studies could be focused on fitting more models and getting a better understanding of season games before making predictions like this.
+I would like to get more robust predictions, therefore, I think it could be beneficial to go back through the seasion games data and draw out the times each team in the March Madness tournament played each other and test our model against that.
+(Well see if we feel like doing that).
+While exploring, I found that teams from different parts of the country tended to have a different coefficients relationship between statistics and outcomes.
+I think the team's conferences are a random effect that should be taken into account.
+I think the season games could be modeled very nicely using linear mixed models.
 
@@
 \setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\setlength{\leftskip}{0.0in}
 
@@
 \endgroup
+
+```{r}
+beepr::beep()
+```

### 2021-04-17 paper/paper.tex
worked on paper first draft complete

--- 2021-04-02 paper/paper.tex
@@
 \PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
 %
 \documentclass[
-  english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
   pdftitle={Using GLMs to Predict Basketball Games},
-  pdfauthor={Sabrina Ball1 \& Joseph Despres1},
-  pdflang={en-EN},
-  pdfkeywords={keywords},
-  hidelinks,
+  pdfauthor={Joe Despres \& Sabrina Ball},
+  pdfkeywords={keyword},
+  colorlinks=true,
+  linkcolor=Maroon,
+  filecolor=Maroon,
+  citecolor=Blue,
+  urlcolor=blue,
   pdfcreator={LaTeX via pandoc}}
@@
 \makeatother
-\shorttitle{GLM}
-\keywords{keywords\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
-\usepackage{lineno}
-
-\linenumbers
+\shorttitle{Using GLMs to Predict Basketball Games}
+\keywords{keyword\newline\indent Word count: X}
 \usepackage{csquotes}
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
 \ifluatex
@@
 
-\textbf{Introduction}
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
-testing citat (R Core Team, 2020)
-\textbf{Exploratory Data Analysis}\\
-\textbf{Description}\\
-\textbf{Results}\\
-\textbf{Diagnostic Methods}\\
-\textbf{Conculsion}\\
+Every year top division I basketball teams compete in the annual March Madness tournament.
+Casual fans and enthusiasts submit predictions gambling small sums of money on the outcomes.
+This study uses logistic, poisson, and multinominal regression models in R (R Core Team, 2020) to predict the March Madness tournament outcomes.
+Our primary objective is to determine which GLM is the most accurate when predicting the results.
+The data we use come from three different sources Kaggle (Kaggle, 2021), NCAA (NCAA, 2021), and the tournament results (NCAA, 2021).
+Kaggle (Kaggle, 2021) provides a comprehensive dataset including all NCAA in-season (non-turniment) basketball games from 2001 to 2020.
+The NCAA (NCAA, 2021) provides team-level statistics for each team.
+We filter, clean, and combine these data using the tidyverse package (Wickham et al., 2019).
+Then use the combination of these datasets to fit our models.
+The objective is to predict the individual game outcomes as accurately as possible and determine which GLM is the most accurate.\\
+Our response terms is the outcome of the game \emph{win or loss} with a 0 possibility of a tie.
+Our aim is to derive a function that will accurately predict this using team-level statistics.
+In our dataset we have many predictors however, we only selected the predictors that were not co-linear.
+First, \emph{field goal percentage} which is the rate attempt to score vs actually scoring.
+\emph{Free-throw percentage}, is the rate that the team has the opportunity to take an unguarded shot.
+Cumulative \emph{Three-point goals made}.
+\emph{Rebounds per game}, counts the times the team recovered the ball after a missed shot.
+\emph{Steals}, the times a team was able to remove possession of the ball from the opposing team.
+\emph{Turnover}, the number of times the team lost possession.
+\emph{Blocks}, The number of times the team was able to block a shot made by the opposing team.\\
+March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
+Therefore, making accurate predictions will be depending on predicting the previous round correctly.
+We did make predictions in that fashion, however our models will be compared by taking independent predictions.
+Meaning, we will filter our data down to using only the 64 teams that qualify, predict every single possible combination of games \({64\choose 2} = 2016\) then use the 63 games played as a sample from the population of all possible games.
+Then use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
+Then we use the results to determine which regression method performs the best.\\
+We will use three regression methods, logistic, poisson, and multinominal, to generate a prediction models, to compare the results, and make a determination as to which is best suited to this problem.
+Logistic regression most naturally suits this suits this problem because games do not tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
+This problem could also be suited to poison regression because the number of points scored is poisson distrubuted.
+After fitting a poisson model we will predict the amount of points Team A will score, then predict the number of points Team B will score then take whichever team was predicted to score more and record that as the prediction.
+
+We find that these models have an accurate between 59\% and 65\% which is well over 50\%.
+Therefore, GLM's are helpful in predicting March madness tournament outcomes.
+We find that the team-level statistics recorded by the NCAA are highly significant and helpful for predicting the tournament outcomes however, come up short of producing truely fantastic results.
+Also, we find that when predicting the probability of a win or a loss we find that the coefficients are symmetric because of the nature of the zero-sum-contest.
+
+From these results we see that logistic and posisson model was the best at predicting wins and losses.
+All three models were well over 50\% accurate, so this the research goal of having GLM's predict March Madness Outcomes.
+Also, we confirmed the suspicion that basketball statistics are symmetrical in relation to the probability of winning.
+Yet are not when we are predicting the score.
+
+\hypertarget{exploratory-data-analysis}{%
+\subsection{Exploratory Data Analysis}\label{exploratory-data-analysis}}
+
+\begin{longtable}[t]{lrrrrrr}
+\caption{\label{tab:unnamed-chunk-1}Descriptive Statistics}\\
+\toprule
+Variable & Mean & Median & Std & Min & Max & Range\\
+\midrule
+Three-Points goals Scored & 242.08 & 241.00 & 48.18 & 107.0 & 454.00 & 347.00\\
+Field Goals Scored Percentage & 44.03 & 44.10 & 2.46 & 35.4 & 52.60 & 17.20\\
+Rebounds Per Game & 35.48 & 35.35 & 2.49 & 27.6 & 43.81 & 16.21\\
+Steels Per Game & 206.34 & 202.00 & 40.59 & 116.0 & 369.00 & 253.00\\
+Turn Overs & 421.84 & 419.00 & 49.01 & 290.0 & 604.00 & 314.00\\
+\addlinespace
+Blocks Per game & 3.30 & 3.20 & 0.94 & 1.3 & 6.80 & 5.50\\
+\bottomrule
+\end{longtable}
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+As mentioned above, there are substantially more variables in the dataset we used, however we omitted the ones that were co-linear.
+When making a decision we selected the variable using deviance tests and likelihood ratios.
+For our modeling and prediction we are going to use these statistics associated with both teams.
+Basic descriptive statistics can be found in Table 1.
+We notice that the mean of the predictor is relatively close to the median.
+The standard deviations can be high, but are not wildly so and the range is reasionable for the data we have.
+In the histograms below we remark the predictors roughly normal distribution with no substantial skew, heavy tails, or slow decay.
+
+Now we turn our attention to the outcomes.
+The logistic model is straight forward all we needed to do was code a win to be 1 and a loss to be 0 then fit the model and make a prediction.
+The possion model will work if we to taking score as a count.
+This is not perfectly suited to poisson because a winning team's score has a mean of 77.13 and a variance of 113.22.
+Regardless we will see how it performs, when we will predict the score for Team A and predict the score for Team B, then take who ever has a higher predictions.
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+The multinomeal case is not obviously suited to this problem however we can adapt it.
+We will take a look at the difference in score between winner and loser.
+This is a common thing in betting markets to gamble on.
+Therefore we mapped the difference in score into 5 categories using the 0.2, 0.4, 0.6, and 0.8 quantifies.
+First, the team losing with by more than 12.
+Second, the team losing by between 12 and 4 points.
+Third, a very close game between losing by 4 and winning by 4 (This is fairly common as there is really good competition between teams).
+Fourth, winning by more than 4 and less than 12.
+In the fifth category, winning by more than 12.
+We will use the probabilities to predict a winner by suming the proabilities of losing by more than 12 and losing by between 12 and four and summing the probability of winning by more than 12 and winning by between four and 12.
+Then taking whichever is larger as a prediction.
+
+\hypertarget{description}{%
+\subsection{Description}\label{description}}
+
+To make an accurate compartment, we use the same formula for the three models with only the dependent variable being different.
+Win or loss for our logistic, the amount Team A scores for the poisson and the difference in score for the multinomeal.
+Now we tried normalizing the data where we subtracted the mean and devided by standard deviation for all out predictors.
+However, that really did not have that much effect.
+Also we individually tested each predictor using likelyhood ratio tests to add each term.
+These all came significant.
+Also when we were adding terms we found many were co-linear, therefore when we found co-linear predictors we omitted the one with a smaller likelihood ratio.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/josephdespres/Documents/MSU/STT-864_Statistical_Methods_II/March-Madness/paper/paper_files/equ} \end{figure}
+
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
+
+In the Table 2 made manually with the kableExtra package (Zhu, 2020), you can see that we have the results for poisson, logistic, and multinomeal models.
+We have a sample of over 35,000 and all of the coeffecients are highly significant.
+This is to be expected because these are the statistics that the NCAA records and ones they have determined to being useful to measure a team's performance.
+The aim of this study is to compare predictive models so we will not cover it exhaustively or include individual z-statistics and p-values.
+First, notice is that in the case of the logistic and multinomeal models we see that when comparing a factors that affect a team's probability of winning we see the associated coefficient is quite similar for the factor capturing the opposing teams.
+That is because basketball is a zero-sum-game, and anything good for team A is proportionately bad for team B.
+Note this is not true of the multinomeal and poisson models.
+The reasion that is is because the winner picked by thoes models is a function of the score.
+Therefore, thoes models are looking at coeffecients that are oging to increase the scores.
+Take three pointers in the poisson model, for instance, a team where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
+More points scored is not deterministic of winning, however, it is an indicator.
+
+\begin{longtable}[t]{llll}
+\caption{\label{tab:unnamed-chunk-5}Regression Output}\\
+\toprule
+Terms & Multinomial Model & Logistic Model & Poisson Model\\
+\midrule
+<-12 & -1.2516 & . & .\\
+-12:-4 & -0.28546 & . & .\\
+-4:4 & 0.54181 & . & .\\
+4:13 & 1.54645 & . & .\\
+Three Pointers & -0.00061 & 0.00243 & 4e-04\\
+\addlinespace
+O* Three Pointers & 0.00067 & -0.00255 & 0.00013\\
+Field Goals & -0.07605 & 0.18322 & 0.01421\\
+O* Field Goals & 0.0752 & -0.17632 & -0.0052\\
+Free-throws & -0.01624 & 0.03242 & 0.00424\\
+O* Free-throws & 0.01418 & -0.03059 & 0.00078\\
+\addlinespace
+Rebounds & -0.06812 & 0.14321 & 0.01286\\
+O* Rebounds & 0.06823 & -0.14476 & -0.00234\\
+Steals & -0.00287 & 0.00655 & 0.00039\\
+O* Steals & 0.00276 & -0.0064 & -0.00021\\
+Turnovers & 0.00234 & -0.00603 & -0.00021\\
+\addlinespace
+O* Turnovers & -0.00251 & 0.0058 & 0.00045\\
+Blocks & 0.06195 & -0.14284 & -0.01497\\
+O* Blocks & -0.05622 & 0.16488 & 0.00263\\
+\bottomrule
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} Sample size 35248 games.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{2} O* is the term asscoiated with the opposing team.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{3} All the above terms are significant at P < 0.01.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{4} Multinomeal Intercepts are differences in predicted score.}\\
+\end{longtable}
+
+\hypertarget{goodness-of-fit}{%
+\subsection{Goodness of Fit}\label{goodness-of-fit}}
+
+Now that we have fitted the model, lets take a look at how well it performed.
+Using the tidyr {[}@-tidyr{]} package, we make every combination of teams and associated statistics in one dataframe.
+Then wrote custom functions that would take two of the 64 teams as inputs and output the probabilities of winning, predicted amounts of points scored, or probabilities of the score resulting in one of the above mentioned multinomeal categories.
+After that we used the purrr (Henry \& Wickham, 2020) package to iterate over all possible games that could be played.
+From there, wrote and ran a webscraping script to obtain a dataframe of the results.
+Then counted the games each model predicted correctly and incorrectly.
+In this case our population is all \({64\choose 2}\) possible games in this turniment and a sample of the games that were played.
+We are not assuming that we have a random sample from the whole population as these are the best teams in the league.
+Also, there is a selection bias twards games that were played.
+Therefore, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams only are represented once.
+Therefore, we do not claim that these results will hold for games in general.
+However, for the purpose of comparing models for this specific turniment this is approprite.
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+\hypertarget{conculsion}{%
+\subsection{Conculsion}\label{conculsion}}
+
+From these results we see that logistic and posisson model was the best at predicting wins and losses.
+All three models were well over 50\% accurate, so this the research goal of having GLM's predict March Madness Outcomes.
+Also, we confirmed the suspicion that basketball statistics are symmetrical in relation to the probability of winning.
+Yet are not when we are predicting the score.
+
+The limits of this study have a lot to do with data limitations.
+Ideally, we would have more tournaments playing with lesser skilled teams.
+would be much more robust if the tournament was much larger.
+We had a lot of data points over 35,000 which I think was more than sufficient, however, we lacking non co-linear covariates.
+We only had 7 predictors.
+An accuracy of 64.5\% is not superb considering betting markets take these exact factors into account.
+
+Future studies could be focused on fitting more models and getting a better understanding of season games before making predictions like this.
+I would like to get more robust predictions, therefore, I think it could be beneficial to go back through the seasion games data and draw out the times each team in the March Madness tournament played each other and test our model against that.
+(Well see if we feel like doing that).
+While exploring, I found that teams from different parts of the country tended to have a different coefficients relationship between statistics and outcomes.
+I think the team's conferences are a random effect that should be taken into account.
+I think the season games could be modeled very nicely using linear mixed models.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-R-purrr}{}%
+Henry, L., \& Wickham, H. (2020). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+
+\leavevmode\hypertarget{ref-R-Kaggle}{}%
+Kaggle. (2021). \emph{March machine learning mania 2021}. Retrieved from \url{https://www.kaggle.com/c/ncaam-march-mania-2021}
+
+\leavevmode\hypertarget{ref-R-NCAA}{}%
+NCAA. (2021). \emph{Men's basketball}. Retrieved from \url{https://www.ncaa.com/stats/basketball-men/d1}
+
 \leavevmode\hypertarget{ref-R-base}{}%
 R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
+D., François, R., \ldots{} Yutani, H.
+(2019).
+Welcome to the {tidyverse}.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\hypertarget{ref-R-kableExtra}{}%
+Zhu, H. (2020). \emph{kableExtra: Construct complex table with 'kable' and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}
 

### 2021-04-17 paper/paper.Rmd
add paper

@@
     institution   : "Michigan State University"
-
 keywords          : "keyword"
@@
 footnotelist      : no
-linenumbers       : no
+linenumbers       : yes
 mask              : no
@@
 
-Every year top division I basketball teams compete in the annual March Madness tournament.
-Casual fans and enthusiasts submit predictions gambling small sums of money on the outcomes.
-This study uses logistic, poisson, and multinominal regression models in R [@R-base] to predict the March Madness tournament outcomes.
-Our primary objective is to determine which GLM is the most accurate when predicting the results.
-The data we use come from three different sources Kaggle [@R-Kaggle], NCAA [@R-NCAA], and the tournament results [@R-NCAA].
-Kaggle [@R-Kaggle] provides a comprehensive dataset including all NCAA in-season (non-turniment) basketball games from 2001 to 2020.
+Every year qualifying Division I basketball teams compete in the annual March Madness tournament.
+Casual fans and enthusiasts submit predictions gambling small sums of money in hopes of completing a perfect bracket.
+A perfect bracket is when one correctly predicts the outcome of all 63 games.
+This study uses logistic, poisson, and multinomial regression models fitted with R [@R-base] to predict the outcomes of the March Madness tournament games.
+Our primary objective is to determine which of these GLMs make the most accurate predictions.
+To begin, we collect data from three different sources Kaggle [@R-Kaggle], NCAA [@R-NCAA], and the tournament results [@R-NCAA].
+Kaggle [@R-Kaggle], provides a comprehensive dataset including all NCAA in-season basketball games from 2001 to 2020.
 The NCAA [@R-NCAA] provides team-level statistics for each team.
@@
 The objective is to predict the individual game outcomes as accurately as possible and determine which GLM is the most accurate.
-Our response terms is the outcome of the game *win or loss* with a 0 possibility of a tie.
+Our response terms is the outcome of the game, *win or loss*, without the possibility of a tie.
 Our aim is to derive a function that will accurately predict this using team-level statistics.
-In our dataset we have many predictors however, we only selected the predictors that were not co-linear.
-First, *field goal percentage* which is the rate attempt to score vs actually scoring.
-*Free-throw percentage*,  is the rate that the team has the opportunity to take an unguarded shot.
-Cumulative *Three-point goals made*.
-*Rebounds per game*, counts the times the team recovered the ball after a missed shot.
-*Steals*, the times a team was able to remove possession of the ball from the opposing team.
+In our dataset, we have many predictors however, we only selected the predictors that are not co-linear.
+First, *field goal percentage* which is the ratio of attempted scores to made scores.
+*Free-throw percentage*,  is the the rate of successfully scoring a penalty shot.
+Cumulative *Three-point goals made*, field goals made from a sufficient distance.
+*Rebounds per game*, counts the amount of times a team recovers the ball after a missed shot.
+*Steals* is when a team was able to remove possession of the ball from the opposing team.
 *Turnover*, the number of times the team lost possession.
-*Blocks*, The number of times the team was able to block a shot made by the opposing team.
+*Blocks*, the number of times the team was able to block a shot made by the opposing team.
 March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
-Therefore, making accurate predictions will be depending on predicting the previous round correctly.
-We did make predictions in that fashion, however our models will be compared by taking independent predictions.
-Meaning, we will filter our data down to using only the 64 teams that qualify, predict every single possible combination of games ${64\choose 2} = 2016$ then use the 63 games played as a sample from the population of all possible games.
+Therefore, making accurate predictions will be dependent on predicting the previous round correctly.
+We did make predictions using that method, however our models will be compared by treating the games as independent of the previous round.
+Meaning, we will filter our data down to include only the 64 teams that qualify, predict every single possible combination of games ${64\choose 2} = 2016$, then use the 63 games played as a sample from the population of all possible games.
 Then use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
 Then we use the results to determine which regression method performs the best.
-We will use three regression methods, logistic, poisson, and multinominal, to generate a prediction models, to compare the results, and make a determination as to which is best suited to this problem.
-Logistic regression most naturally suits this suits this problem because games do not tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
-This problem could also be suited to poison regression because the number of points scored is poisson distrubuted.
-After fitting a poisson model we will predict the amount of points Team A will score, then predict the number of points Team B will score then take whichever team was predicted to score more and record that as the prediction.
+We will use three regression methods, logistic, poisson, and multinomial, to generate prediction models, compare the results, and make a determination as to which is best suited to this problem.
+Logistic regression most naturally suits this problem because games cannot tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
+This problem could also be suited to poison regression because the number of points scored is roughly poisson distributed.
+After fitting a poisson model, we will predict the number of points scored by Team A, predict the number of points scored by Team B, and take the team with the highest predicted score to be the winner.
+A multinomial model is less naturally suited to this problem, however it may address the shortcomings of the logistic model.
+In particular, many games are close scoring and are won by merely a few points.
+Therefore, we will use multinomeal regression to predict the difference in game score by assigning the differences into five categories.
+From there, assign a predicted winner based on which outcome the model considers to be the most likely.
 
-We find that these models have an accurate between 59% and 65% which is well over 50%.
-Therefore, GLM's are helpful in predicting March madness tournament outcomes.
-We find that the team-level statistics recorded by the NCAA are highly significant and helpful for predicting the tournament outcomes however, come up short of producing truely fantastic results.
-Also, we find that when predicting the probability of a win or a loss we find that the coefficients are symmetric because of the nature of the zero-sum-contest.
-
-From these results we see that logistic and posisson model was the best at predicting wins and losses.
-All three models were well over 50% accurate, so this the research goal of having GLM's predict March Madness Outcomes.
-Also, we confirmed the suspicion that basketball statistics are symmetrical in relation to the probability of winning.
-Yet are not when we are predicting the score.
+These models predict the correct outcomes with an accuracy between 59% and 65%.
+GLMs are accurate relative to randomly guessing March Madness tournament outcomes.
+Team-level statistics, recorded by the NCAA, are highly significant and helpful predicting the tournament outcomes but do not yield fantastic results.
+Also, when predicting the probability of a win or a loss using a logit link, the coefficients are symmetric because basketball games are a zero-sum-contest.
+From our results, we see that logistic and poisson models are the best at predicting wins and losses.
+All three GLM models are well over 50% accurate, therefore, we claim that using GLM models are more accurate than random chance.
 
@@
 
-```{r }
+```{r, fig.align="center"}
 histogram <- function(x, t) {
@@
 c <- histogram(merged$rpg, t = "Rebounds Per Game")
-d <- histogram(merged$st, t = "Steels Per Game")
+d <- histogram(merged$st, t = "Steals Per Game")
 e <- histogram(merged$to, t = "Turn Overs Per Game")
@@
 
-As mentioned above, there are substantially more variables in the dataset we used, however we omitted the ones that were co-linear.
-When making a decision we selected the variable using deviance tests and likelihood ratios.
-For our modeling and prediction we are going to use these statistics associated with both teams.
+As mentioned above, there are substantially more variables in the dataset than we used, however we omitted the ones that were co-linear.
+When making a decision on which to keep, we selected the variable using deviance and likelihood ratio tests.
+For our modeling and prediction we are going to use the statistics associated with both teams.
 Basic descriptive statistics can be found in Table 1.
 We notice that the mean of the predictor is relatively close to the median.
-The standard deviations can be high, but are not wildly so and the range is reasionable for the data we have.
-In the histograms below we remark the predictors roughly normal distribution with no substantial skew, heavy tails, or slow decay.
+The standard deviations can be high, but not substantially high, and the ranges are reasonable for the data we have.
+We remark on the predictors being roughly normally distributed with no substantial skew, heavy tails, or slow decay as shown in Figure 1.
 
 Now we turn our attention to the outcomes.
-The logistic model is straight forward all we needed to do was code a win to be 1 and a loss to be 0 then fit the model and make a prediction.
-The possion model will work if we to taking score as a count.
+The logistic model is straight forward, only needing a win to be coded as 1 and a loss to be coded as 0.
+We fit a poisson model by taking a score as a count.
 This is not perfectly suited to poisson because a winning team's score has a mean of `r mean(merged$w_score) %>% round(2)` and a variance of `r var(merged$w_score) %>% round(2)`.
-Regardless we will see how it performs, when we will predict the score for Team A and predict the score for Team B, then take who ever has a higher predictions.
+Regardless we will see how it performs by predicting the score for Team A and Team B, then choose the one with the highest predicted score as the winner.
 
-```{r warning=FALSE}
+```{r warning=FALSE, fig.align="center"}
 w <- merged %>% 
@@
 ```
-The multinomeal case is not obviously suited to this problem however we can adapt it.
-We will take a look at the difference in score between winner and loser.
-This is a common thing in betting markets to gamble on.
-Therefore we mapped the difference in score into 5 categories using the 0.2, 0.4, 0.6, and 0.8 quantifies.
-First, the team losing with by more than 12.
-Second, the team losing by between 12 and 4 points.
-Third, a very close game between losing by 4 and winning by 4 (This is fairly common as there is really good competition between teams).
-Fourth, winning by more than 4 and less than 12.
-In the fifth category, winning by more than 12.
-We will use the probabilities to predict a winner by suming the proabilities of losing by more than 12 and losing by between 12 and four and summing the probability of winning by more than 12 and winning by between four and 12.
-Then taking whichever is larger as a prediction.
+
+The multinomeal case is not obviously suited to this problem, however, we can adapt it.
+Commonly seen in betting markets, we will take a look at the difference in score between winner and loser.
+Then assign it into five categories using the quintiles as shown in Figure 2.
+First category, is the probability of the team losing by more than twelve points.
+Second, the probability of the team losing between twelve and four points.
+Third, a very close game either losing by four or less or winning by four or less (this is fairly common as there is fierce competition between the teams).
+Fourth, winning by more than four and less than twelve points.
+In the fifth category, the probability of winning by more than twelve points.
+We will use the probabilities to predict a winner by comparing the sum of the predicted probabilities of being in the first and second quintiles to being in the fourth and fifth quintiles.
 
@@
 
-To make an accurate compartment, we use the same formula for the three models with only the dependent variable being different.
-Win or loss for our logistic, the amount Team A scores for the poisson and the difference in score for the multinomeal.
-Now we tried normalizing the data where we subtracted the mean and devided by standard deviation for all out predictors.
-However, that really did not have that much effect.
-Also we individually tested each predictor using likelyhood ratio tests to add each term.
-These all came significant.
-Also when we were adding terms we found many were co-linear, therefore when we found co-linear predictors we omitted the one with a smaller likelihood ratio.
+To make an accurate comparison, we use the same formula for the three models with only the dependent variable being different.
+Win or loss for our logistic model, the amount Team A scores for the poisson model, and the difference in score for the multinomeal model.
+We tried normalizing the data by subtracting the mean and dividing by the standard deviation for all predictors.
+However, that really did not have an effect on the prediction accuracy.
+Also, we individually tested each predictor using the likelihood ratio test before adding each term.
+When we were adding terms, we found many to be co-linear.
+When we found co-linear predictors, we omitted the one with a smaller likelihood ratio statistic.
 
@@
 
-In the Table 2 made manually with the kableExtra package [@R-kableExtra], you can see that we have the results for poisson, logistic, and multinomeal models.
-We have a sample of over 35,000 and all of the coeffecients are highly significant.
-This is to be expected because these are the statistics that the NCAA records and ones they have determined to being useful to measure a team's performance.
+In Table 2, made manually with the kableExtra package [@R-kableExtra], you can see that we have the results for the poisson, logistic, and multinomial models.
+We have a sample of over 35,000 and all of the coefficients are highly significant.
+Which is to be expected because these are the statistics that the NCAA collects as the metrics useful in measuring a team's ability to win games.
 The aim of this study is to compare predictive models so we will not cover it exhaustively or include individual z-statistics and p-values.
-First, notice is that in the case of the logistic and multinomeal models we see that when comparing a factors that affect a team's probability of winning we see the associated coefficient is quite similar for the factor capturing the opposing teams.
-That is because basketball is a zero-sum-game, and anything good for team A is proportionately bad for team B.
-Note this is not true of the  multinomeal and poisson models.
-The reasion that is is because the winner picked by thoes models is a function of the score.
-Therefore, thoes models are looking at coeffecients that are oging to increase the scores.
-Take three pointers in the poisson model, for instance, a team where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
+First notice, in the case of the logistic and multinomial models we see that when comparing factors that affect a team's probability of winning, the associated coefficient is nearly equal to the factor capturing the opposing teams metric.
+That is because basketball is a zero-sum-game, anything good for team A is proportionately bad for team B.
+Note this is not true for the poisson model because that is measuring points scored.
+Take three pointers in the poisson model, for instance, where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
 More points scored is not deterministic of winning, however, it is an indicator.
@@
   mutate_if(is.numeric, round, 5) %>% 
-  mutate_at(vars(Terms:`Poisson Model`), replace_na, '.') %>% 
+  mutate_at(vars(Terms:`Poisson Model`), replace_na, '--') %>% 
   kable(format = "latex", booktabs = TRUE,
@@
 After that we used the purrr [@R-purrr] package to iterate over all possible games that could be played.
-From there, wrote and ran a webscraping script  to obtain a dataframe of the results.
+From there, wrote and ran a web scraping script to obtain a dataframe of the results.
 Then counted the games each model predicted correctly and incorrectly.
-In this case our population is all  ${64\choose 2}$ possible games in this turniment and a sample of the games that were played.
+In this case our population is all ${64\choose 2}$ possible games in this tournament and a sample is the 63 games that were played.
 We are not assuming that we have a random sample from the whole population as these are the best teams in the league.
-Also, there is a selection bias twards games that were played.
+Also, there is a selection bias towards games that were played.
 Therefore, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams only are represented once.
 Therefore, we do not claim that these results will hold for games in general.
-However, for the purpose of comparing models for this specific turniment this is approprite.
+However, for the purpose of comparing models that make predictions in this specific tournament this is appropriate.
 
@@
 
-From these results we see that logistic and posisson model was the best at predicting wins and losses.
-All three models were well over 50% accurate, so this the research goal of having GLM's predict March Madness Outcomes.
-Also, we confirmed the suspicion that basketball statistics are symmetrical in relation to the probability of winning.
-Yet are not when we are predicting the score.
+From these results we see that logistic and poisson model are best at predicting wins and losses.
+All three models are well over 50% accurate, confirming the claim that GLMs can predict March Madness outcomes better than random chance.
+Also, we confirmed the suspicion that basketball statistics are symmetric in relation to the probability of winning and are not when we are predicting the game score.
 
 The limits of this study have a lot to do with data limitations.
-Ideally, we would have more tournaments playing with lesser skilled teams.
-would be much more robust if the tournament was much larger.
-We had a lot of data points over 35,000 which I think was more than sufficient, however, we lacking non co-linear covariates.
-We only had 7 predictors.
-An accuracy of 64.5% is not superb considering betting markets take these exact factors into account.
+Ideally, we would have more tournaments played with lower performing teams.
+Also, the results would be much more robust if the tournament was larger.
+We had a lot of data points.
+over 35,000, which is more than sufficient, however, we do not have an abundance of non co-linear covariates.
+An accuracy of 64.5% is not superb considering betting markets take these exact factors into account, however, being higher than random chance indicates that GLMs are useful in predicting March Madness outcomes.
 
-Future studies could be focused on fitting more models and getting a better understanding of season games before making predictions like this.
-I would like to get more robust predictions, therefore, I think it could be beneficial to go back through the seasion games data and draw out the times each team in the March Madness tournament played each other and test our model against that.
-(Well see if we feel like doing that).
-While exploring, I found that teams from different parts of the country tended to have a different coefficients relationship between statistics and outcomes.
-I think the team's conferences are a random effect that should be taken into account.
-I think the season games could be modeled very nicely using linear mixed models.
+Future studies could be focused on making a better fit and getting a better understanding of season games before making predictions like this.
+To get more robust predictions, it would be beneficial to go back through the season games data and draw out the times each team in the March Madness tournament played each other and test our model against that.
+While exploring, we found that teams from different parts of the country tended to have a different coefficient on statistics and outcomes.
+A team's conferences is likely a random effect that should be taken into account.
+Therefore, additional research into the random effect would certainly yield more robust predictions.
 

### 2021-04-17 paper/paper.tex
add paper

--- 2021-04-02 paper/paper.tex
@@
 \PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
 %
 \documentclass[
-  english,
-  man]{apa6}
-\usepackage{amsmath,amssymb}
+  man,floatsintext]{apa6}
 \usepackage{lmodern}
+\usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
@@
 \makeatother
-\shorttitle{GLM}
-\keywords{keywords\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
+\shorttitle{Using GLMs to Predict Basketball Games}
+\keywords{keyword\newline\indent Word count: X}
 \usepackage{lineno}
@@
 \usepackage{csquotes}
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
-\ifluatex
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
 \title{Using GLMs to Predict Basketball Games}
-\author{Sabrina Ball\textsuperscript{1} \& Joseph Despres\textsuperscript{1}}
+\author{Joe Despres\textsuperscript{} \& Sabrina Ball\textsuperscript{}}
 \date{}
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Michigan State University}
+\affiliation{\vspace{0.5cm}\textsuperscript{} Michigan State University}
 
@@
 
-\textbf{Introduction}
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
-testing citat (R Core Team, 2020)
-\textbf{Exploratory Data Analysis}\\
-\textbf{Description}\\
-\textbf{Results}\\
-\textbf{Diagnostic Methods}\\
-\textbf{Conculsion}\\
+Every year qualifying Division I basketball teams compete in the annual March Madness tournament.
+Casual fans and enthusiasts submit predictions gambling small sums of money in hopes of completing a perfect bracket.
+A perfect bracket is when one correctly predicts the outcome of all 63 games.
+This study uses logistic, poisson, and multinomial regression models fitted with R (R Core Team, 2020) to predict the outcomes of the March Madness tournament games.
+Our primary objective is to determine which of these GLMs make the most accurate predictions.
+To begin, we collect data from three different sources Kaggle (Kaggle, 2021), NCAA (NCAA, 2021), and the tournament results (NCAA, 2021).
+Kaggle (Kaggle, 2021), provides a comprehensive dataset including all NCAA in-season basketball games from 2001 to 2020.
+The NCAA (NCAA, 2021) provides team-level statistics for each team.
+We filter, clean, and combine these data using the tidyverse package (Wickham et al., 2019).
+Then use the combination of these datasets to fit our models.
+The objective is to predict the individual game outcomes as accurately as possible and determine which GLM is the most accurate.\\
+Our response terms is the outcome of the game, \emph{win or loss}, without the possibility of a tie.
+Our aim is to derive a function that will accurately predict this using team-level statistics.
+In our dataset, we have many predictors however, we only selected the predictors that are not co-linear.
+First, \emph{field goal percentage} which is the ratio of attempted scores to made scores.
+\emph{Free-throw percentage}, is the the rate of successfully scoring a penalty shot.
+Cumulative \emph{Three-point goals made}, field goals made from a sufficient distance.
+\emph{Rebounds per game}, counts the amount of times a team recovers the ball after a missed shot.
+\emph{Steals} is when a team was able to remove possession of the ball from the opposing team.
+\emph{Turnover}, the number of times the team lost possession.
+\emph{Blocks}, the number of times the team was able to block a shot made by the opposing team.\\
+March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
+Therefore, making accurate predictions will be dependent on predicting the previous round correctly.
+We did make predictions using that method, however our models will be compared by treating the games as independent of the previous round.
+Meaning, we will filter our data down to include only the 64 teams that qualify, predict every single possible combination of games \({64\choose 2} = 2016\), then use the 63 games played as a sample from the population of all possible games.
+Then use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
+Then we use the results to determine which regression method performs the best.\\
+We will use three regression methods, logistic, poisson, and multinomial, to generate prediction models, compare the results, and make a determination as to which is best suited to this problem.
+Logistic regression most naturally suits this problem because games cannot tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
+This problem could also be suited to poison regression because the number of points scored is roughly poisson distributed.
+After fitting a poisson model, we will predict the number of points scored by Team A, predict the number of points scored by Team B, and take the team with the highest predicted score to be the winner.
+A multinomial model is less naturally suited to this problem, however it may address the shortcomings of the logistic model.
+In particular, many games are close scoring and are won by merely a few points.
+Therefore, we will use multinomeal regression to predict the difference in game score by assigning the differences into five categories.
+From there, assign a predicted winner based on which outcome the model considers to be the most likely.
+
+These models predict the correct outcomes with an accuracy between 59\% and 65\%.
+GLMs are accurate relative to randomly guessing March Madness tournament outcomes.
+Team-level statistics, recorded by the NCAA, are highly significant and helpful predicting the tournament outcomes but do not yield fantastic results.
+Also, when predicting the probability of a win or a loss using a logit link, the coefficients are symmetric because basketball games are a zero-sum-contest.
+From our results, we see that logistic and poisson models are the best at predicting wins and losses.
+All three GLM models are well over 50\% accurate, therefore, we claim that using GLM models are more accurate than random chance.
+
+\hypertarget{exploratory-data-analysis}{%
+\subsection{Exploratory Data Analysis}\label{exploratory-data-analysis}}
+
+\begin{longtable}[t]{lrrrrrr}
+\caption{\label{tab:unnamed-chunk-1}Descriptive Statistics}\\
+\toprule{}
+Variable & Mean & Median & Std & Min & Max & Range\\
+\midrule{}
+Three-Points goals Scored & 242.08 & 241.00 & 48.18 & 107.0 & 454.00 & 347.00\\
+Field Goals Scored Percentage & 44.03 & 44.10 & 2.46 & 35.4 & 52.60 & 17.20\\
+Rebounds Per Game & 35.48 & 35.35 & 2.49 & 27.6 & 43.81 & 16.21\\
+Steels Per Game & 206.34 & 202.00 & 40.59 & 116.0 & 369.00 & 253.00\\
+Turn Overs & 421.84 & 419.00 & 49.01 & 290.0 & 604.00 & 314.00\\
+\addlinespace
+Blocks Per game & 3.30 & 3.20 & 0.94 & 1.3 & 6.80 & 5.50\\
+\bottomrule{}
+\end{longtable}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-2-1}
+
+}
+
+\caption{ }\label{fig:unnamed-chunk-2}
+\end{figure}
+
+As mentioned above, there are substantially more variables in the dataset than we used, however we omitted the ones that were co-linear.
+When making a decision on which to keep, we selected the variable using deviance and likelihood ratio tests.
+For our modeling and prediction we are going to use the statistics associated with both teams.
+Basic descriptive statistics can be found in Table 1.
+We notice that the mean of the predictor is relatively close to the median.
+The standard deviations can be high, but not substantially high, and the ranges are reasonable for the data we have.
+We remark on the predictors being roughly normally distributed with no substantial skew, heavy tails, or slow decay as shown in Figure 1.
+
+Now we turn our attention to the outcomes.
+The logistic model is straight forward, only needing a win to be coded as 1 and a loss to be coded as 0.
+We fit a poisson model by taking a score as a count.
+This is not perfectly suited to poisson because a winning team's score has a mean of 77.13 and a variance of 113.22.
+Regardless we will see how it performs by predicting the score for Team A and Team B, then choose the one with the highest predicted score as the winner.
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-3-1}
+
+}
+
+\caption{ }\label{fig:unnamed-chunk-3}
+\end{figure}
+
+The multinomeal case is not obviously suited to this problem, however, we can adapt it.
+Commonly seen in betting markets, we will take a look at the difference in score between winner and loser.
+Then assign it into five categories using the quintiles as shown in Figure 2.
+First category, is the probability of the team losing by more than twelve points.
+Second, the probability of the team losing between twelve and four points.
+Third, a very close game either losing by four or less or winning by four or less (this is fairly common as there is fierce competition between the teams).
+Fourth, winning by more than four and less than twelve points.
+In the fifth category, the probability of winning by more than twelve points.
+We will use the probabilities to predict a winner by comparing the sum of the predicted probabilities of being in the first and second quintiles to being in the fourth and fifth quintiles.
+
+\hypertarget{description}{%
+\subsection{Description}\label{description}}
+
+To make an accurate comparison, we use the same formula for the three models with only the dependent variable being different.
+Win or loss for our logistic model, the amount Team A scores for the poisson model, and the difference in score for the multinomeal model.
+We tried normalizing the data by subtracting the mean and dividing by the standard deviation for all predictors.
+However, that really did not have an effect on the prediction accuracy.
+Also, we individually tested each predictor using the likelihood ratio test before adding each term.
+When we were adding terms, we found many to be co-linear.
+When we found co-linear predictors, we omitted the one with a smaller likelihood ratio statistic.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/josephdespres/Documents/MSU/STT-864_Statistical_Methods_II/March-Madness/paper/paper_files/equ} \end{figure}
+
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
+
+In Table 2, made manually with the kableExtra package (Zhu, 2020), you can see that we have the results for the poisson, logistic, and multinomial models.
+We have a sample of over 35,000 and all of the coefficients are highly significant.
+Which is to be expected because these are the statistics that the NCAA collects as the metrics useful in measuring a team's ability to win games.
+The aim of this study is to compare predictive models so we will not cover it exhaustively or include individual z-statistics and p-values.
+First notice, in the case of the logistic and multinomial models we see that when comparing factors that affect a team's probability of winning, the associated coefficient is nearly equal to the factor capturing the opposing teams metric.
+That is because basketball is a zero-sum-game, anything good for team A is proportionately bad for team B.
+Note this is not true for the poisson model because that is measuring points scored.
+Take three pointers in the poisson model, for instance, where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
+More points scored is not deterministic of winning, however, it is an indicator.
+
+\begin{longtable}[t]{llll}
+\caption{\label{tab:unnamed-chunk-5}Regression Output}\\
+\toprule{}
+Terms & Multinomial Model & Logistic Model & Poisson Model\\
+\midrule{}
+<-12 & -1.2516 & -- & --\\
+-12:-4 & -0.28546 & -- & --\\
+-4:4 & 0.54181 & -- & --\\
+4:13 & 1.54645 & -- & --\\
+Three Pointers & -0.00061 & 0.00243 & 4e-04\\
+\addlinespace
+O* Three Pointers & 0.00067 & -0.00255 & 0.00013\\
+Field Goals & -0.07605 & 0.18322 & 0.01421\\
+O* Field Goals & 0.0752 & -0.17632 & -0.0052\\
+Free-throws & -0.01624 & 0.03242 & 0.00424\\
+O* Free-throws & 0.01418 & -0.03059 & 0.00078\\
+\addlinespace
+Rebounds & -0.06812 & 0.14321 & 0.01286\\
+O* Rebounds & 0.06823 & -0.14476 & -0.00234\\
+Steals & -0.00287 & 0.00655 & 0.00039\\
+O* Steals & 0.00276 & -0.0064 & -0.00021\\
+Turnovers & 0.00234 & -0.00603 & -0.00021\\
+\addlinespace
+O* Turnovers & -0.00251 & 0.0058 & 0.00045\\
+Blocks & 0.06195 & -0.14284 & -0.01497\\
+O* Blocks & -0.05622 & 0.16488 & 0.00263\\
+\bottomrule
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} Sample size 35248 games.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{2} O* is the term asscoiated with the opposing team.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{3} All the above terms are significant at P < 0.01.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{4} Multinomeal Intercepts are differences in predicted score.}\\{}
+\end{longtable}
+
+\hypertarget{goodness-of-fit}{%
+\subsection{Goodness of Fit}\label{goodness-of-fit}}
+
+Now that we have fitted the model, lets take a look at how well it performed.
+Using the tidyr {[}@-tidyr{]} package, we make every combination of teams and associated statistics in one dataframe.
+Then wrote custom functions that would take two of the 64 teams as inputs and output the probabilities of winning, predicted amounts of points scored, or probabilities of the score resulting in one of the above mentioned multinomeal categories.
+After that we used the purrr (Henry \& Wickham, 2020) package to iterate over all possible games that could be played.
+From there, wrote and ran a web scraping script to obtain a dataframe of the results.
+Then counted the games each model predicted correctly and incorrectly.
+In this case our population is all \({64\choose 2}\) possible games in this tournament and a sample is the 63 games that were played.
+We are not assuming that we have a random sample from the whole population as these are the best teams in the league.
+Also, there is a selection bias towards games that were played.
+Therefore, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams only are represented once.
+Therefore, we do not claim that these results will hold for games in general.
+However, for the purpose of comparing models that make predictions in this specific tournament this is appropriate.
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+\hypertarget{conculsion}{%
+\subsection{Conculsion}\label{conculsion}}
+
+From these results we see that logistic and poisson model are best at predicting wins and losses.
+All three models are well over 50\% accurate, confirming the claim that GLMs can predict March Madness outcomes better than random chance.
+Also, we confirmed the suspicion that basketball statistics are symmetric in relation to the probability of winning and are not when we are predicting the game score.
+
+The limits of this study have a lot to do with data limitations.
+Ideally, we would have more tournaments played with lower performing teams.
+Also, the results would be much more robust if the tournament was larger.
+We had a lot of data points.
+over 35,000, which is more than sufficient, however, we do not have an abundance of non co-linear covariates.
+An accuracy of 64.5\% is not superb considering betting markets take these exact factors into account, however, being higher than random chance indicates that GLMs are useful in predicting March Madness outcomes.
+
+Future studies could be focused on making a better fit and getting a better understanding of season games before making predictions like this.
+To get more robust predictions, it would be beneficial to go back through the season games data and draw out the times each team in the March Madness tournament played each other and test our model against that.
+While exploring, we found that teams from different parts of the country tended to have a different coefficient on statistics and outcomes.
+A team's conferences is likely a random effect that should be taken into account.
+Therefore, additional research into the random effect would certainly yield more robust predictions.
 
@@
 \setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\setlength{\leftskip}{0.0in}
 
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
+\begin{cslreferences}
+\leavevmode\hypertarget{ref-R-purrr}{}%
+Henry, L., \& Wickham, H. (2020). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+
+\leavevmode\hypertarget{ref-R-Kaggle}{}%
+Kaggle. (2021). \emph{March machine learning mania 2021}. Retrieved from \url{https://www.kaggle.com/c/ncaam-march-mania-2021}
+
+\leavevmode\hypertarget{ref-R-NCAA}{}%
+NCAA. (2021). \emph{Men's basketball}. Retrieved from \url{https://www.ncaa.com/stats/basketball-men/d1}
+
 \leavevmode\hypertarget{ref-R-base}{}%
@@
 
-\end{CSLReferences}
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
+D., François, R., \ldots{} Yutani, H.
+(2019).
+Welcome to the tidyverse.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\hypertarget{ref-R-kableExtra}{}%
+Zhu, H. (2020). \emph{KableExtra: Construct complex table with 'kable' and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}
+\end{cslreferences}
 

### 2021-04-17 R/bracket.R

@@ -1,89 +1,0 @@
-library(tidyverse)
-# https://plexkits.com/march-madness-bracket/ I think this will update
-# https://docs.google.com/spreadsheets/d/1zaWZ2Xh7sll-PZKA1DhKvS6TdfE0Yggk0m5OWY4aICc/edit#gid=0
-
-team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
-bracket <- readxl::read_excel(here::here("data", "prebracket.xlsx"))
-begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
-
-# Even better
-# bracket <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1zaWZ2Xh7sll-PZKA1DhKvS6TdfE0Yggk0m5OWY4aICc/edit#gid=0")
-# remember API times out...
-
-bracket <- bracket %>% 
-  select(2, 35)
-  
-colnames(bracket) <- c("x1", "x2")
-
-
-# bracket -----------------------------------------------------------------
-
-
-bracket <- stack(bracket) %>% 
-  select(values) %>% 
-  drop_na() %>% 
-  tibble() %>% 
-  rename(team = values) %>% 
-  mutate(team = tolower(team)) %>% 
-  mutate(game = 1:n(),
-         game = as.numeric(game),
-         otherteam  = if_else(game %% 2 == 1, team, lag(team)),
-         game = if_else(game %% 2 == 1, game + 1, game), .before = team) %>% 
-  filter(team != otherteam) %>% 
-  separate(team, into = c("team", "pocket"), "/") %>% 
-  filter(otherteam != "first round") %>% 
-  left_join(team_names, by = c("team" = "TeamNameSpelling")) %>% 
-  rename(teamid = TeamID) %>% 
-  left_join(team_names, by = c("otherteam" = "TeamNameSpelling")) %>%
-  rename(otherteamid = TeamID) %>% 
-  mutate(teamid = replace(teamid, team == "app st.", 1111)) %>% 
-  mutate_if(is.numeric, as.character)
-
-bracket
-
-# pockets -----------------------------------------------------------------
-
-pocket <- bracket %>% 
-  select(team, pocket) %>% 
-  drop_na() %>% 
-  left_join(team_names, by = c("team" = "TeamNameSpelling")) %>% 
-  rename(teamid = TeamID, otherteam = pocket) %>% 
[…]

### 2021-04-17 R/cleaningScript.R

@@ -1,183 +1,0 @@
-library(tidyverse) 
-# options(tibble.print_min = 35)
-list_of_dfs <- sapply(paste0("rawdata/", dir("rawdata")), read_csv, USE.NAMES = TRUE)
-
-games <- list_of_dfs$`rawdata/MConferenceTourneyGames.csv` %>% 
-  transmute(gameid = paste0(DayNum, WTeamID, LTeamID),
-            Season, ConfAbbrev) 
-
-seasionsoutcomes <- list_of_dfs$`rawdata/MRegularSeasonDetailedResults.csv` %>% 
-  left_join(list_of_dfs$`rawdata/MTeams.csv`, by = c("WTeamID" = "TeamID")) %>% 
-  filter(Season %in% c(2019, 2020, 2021)) %>% 
-  rename(WTeamName = TeamName) %>% 
-  left_join( list_of_dfs$`rawdata/MTeams.csv`, by = c("LTeamID" = "TeamID")) %>% 
-  select(1:7) %>% 
-  mutate(W = WTeamID, L =  LTeamID) %>% 
-  pivot_longer(cols = ends_with("ID"),
-               names_to = "outcome", 
-               values_to = "TeamID") %>% 
-  mutate(win = if_else(outcome == "WTeamID", 1 ,0),
-         TeamID = as.character(TeamID),
-         Season = as.character(Season)) %>% 
-  select(-outcome, -DayNum) %>% 
-  rename(season = Season)
-
-seasionsoutcomes
-
-stats_fn <- function (directory) {
-  stats <- sapply(paste0(directory, "/", dir(directory)),
-                  readxl::read_excel, USE.NAMES = TRUE)
-  stats <- plyr::join_all(stats, by= "Team")
-  stats <- stats[, !duplicated(colnames(stats), fromLast = TRUE)]
-  return(stats)
-}
-
-stats <- sapply(list.dirs("se_data")[-1], stats_fn)
-selector <- function (df, cols = co) {
-  grabbed <- select(df, contains(cols))
-  return(grabbed)
-}
-
-a <- colnames(stats$`se_data/se_2019`)
-b <- colnames(stats$`se_data/se_2020`)
-c <- colnames(stats$`se_data/se_2021`)
-
-co <- as.data.frame(table(c(a, b, c))) %>% 
-  filter(Freq == 3) %>% 
[…]

### 2021-04-17 R/explatoryAnalysis.R

@@ -1,77 +1,0 @@
-library(tidyverse)
-
-list_of_dfs <- sapply(paste0("rawdata/", dir("rawdata")), read_csv, USE.NAMES = TRUE)
-
-# https://www.kaggle.com/c/ncaam-march-mania-2021/data
-lapply(list_of_dfs, colnames)
-lapply(list_of_dfs, glimpse)
-
-# Keys --------------------------------------------------------------------
-# give us info about the data
-
-list_of_dfs$`rawdata/MSeasons.csv`
-# identifies the different seasons included in the historical data, 
-# along with certain season-level properties.
-
-list_of_dfs$`rawdata/MTeams.csv` # JOINED
-# identifies the id number of the team
-
-list_of_dfs$`rawdata/MTeamConferences.csv`
-# indicates the conference affiliations for 
-# each team during each season. 
-
-list_of_dfs$`rawdata/MTeamCoaches.csv`
-# indicates the head coach for each team in each season
-
-list_of_dfs$`rawdata/MNCAATourneySlots.csv`
-# bracket positions
-
-list_of_dfs$`rawdata/MNCAATourneySeedRoundSlots.csv`
-# seeds with slot and day and late
-
-list_of_dfs$`rawdata/MMasseyOrdinals.csv` 
-# lists out team rankings 
-
-list_of_dfs$`rawdata/Conferences.csv`
-# indicates the Division I conferences
-
-list_of_dfs$`rawdata/MTeamSpellings.csv`
-# indicates alternative spellings of many team names.
-
-list_of_dfs$`rawdata/Cities.csv`
-# This file provides a master list of cities that have
-# been locations for games played.
-
-# Data --------------------------------------------------------------------
-# This is information about games played
-
-list_of_dfs$`rawdata/MNCAATourneyDetailedResults.csv`
-# provides team-level box scores for many NCAA® tournaments
-
-list_of_dfs$`rawdata/MNCAATourneyCompactResults.csv`
-#identifies the game-by-game NCAA® tournament results 
-#for all seasons of historical data.
-
-list_of_dfs$`rawdata/MRegularSeasonDetailedResults.csv`
-# provides team-level box scores for many regular seasons 
-# of historical data, starting with the 2003 season. 
-
-list_of_dfs$`rawdata/MRegularSeasonCompactResults.csv`
-# identifies the game-by-game results for many 
-# seasons of historical data, starting with the 1985 season 
-
-# Probally not useful -----------------------------------------------------
[…]

### 2021-04-17 R/fitting.R

@@ -1,68 +1,0 @@
-logistic_fit <- glm(win ~ x3fg +
-                       opposingx3fg +
-                       # field goal pct
-                       fg_percent +
-                       opposingfg_percent +
-                       # free throws
-                       ft_percent +
-                       opposingft_percent +
-                       # rebound per game
-                       rpg +
-                       opposingrpg +
-                       # steels
-                       st +
-                       opposingst +
-                       #turnover
-                       to +
-                       opposingto +
-                       # blocks
-                       opposingbkpg +
-                       bkpg,
-                     data = merged, family = "binomial")
-
-logistic_fit_null <- glm(win ~ 1,
-                    data = merged, family = "binomial") 
-
- fit_poisson <- glm(team_score ~ x3fg +
-              opposingx3fg +
-              # field goal pct
-              fg_percent +
-              opposingfg_percent +
-              # free throws
-              ft_percent +
-              opposingft_percent +
-              # rebound per game
-              rpg +
-              opposingrpg +
-              # steels
-              st +
-              opposingst +
-              #turnover
-              to +
-              opposingto +
-              # blocks
-              opposingbkpg +
-              bkpg,
-              data = merged, family = poisson(link = "log"))
- 
- fit_multinom <- VGAM::vglm(x_tile ~ x3fg +
-                  opposingx3fg +
-                  # field goal pct
-                  fg_percent +
-                  opposingfg_percent +
-                  # free throws
-                  ft_percent +
-                  opposingft_percent +
-                  # rebound per game
-                  rpg +
-                  opposingrpg +
-                  # steels
-                  st +
-                  opposingst +
-                  #turnover
-                  to +
-                  opposingto +
-                  # blocks
-                  opposingbkpg +
-                  bkpg, family = VGAM::cumulative(parallel=TRUE),
-                  data = merged)

### 2021-04-17 R/helper_functions.R

@@ -1,55 +1,0 @@
-align_teams <- function(df){
-  
-  team <- filter(df, game != "team")
-  other <- filter(df, game == "team")
-  
-  output <- other %>% 
-    rename_all(~ paste0("other",.)) %>% 
-    bind_cols(team) 
-  
-  return(output)
-}
-
-advance_round <- function(df){
-  df <- df %>% 
-    select(predicted_winner, predicted_winnerid) %>% 
-    rename(team = predicted_winner, teamid = predicted_winnerid) %>% 
-    mutate(game = rep(c("team", "otherteam"), nrow(df)/2))
-  
-  return(df)
-}
-
-nice_format <- function(df){
-  output <- df %>%
-    mutate(pred_winner_prob = if_else(team_prob > 0.5, team_prob, otherteam_prob),
-           pred_loser_prob = if_else(team_prob < 0.5, team_prob, otherteam_prob)) %>% 
-    select(game, predicted_winner, 
-           pred_winner_prob, pred_loser_prob)
-  
-  return(output)
-}
-
-add_probs <- function(df, probs){
-  
-  df <- df %>% 
-    mutate(team_prob = probs,
-           otherteam_prob = 1 - team_prob, 
-           game = paste0(team, " vs. ", otherteam),
-           predicted_winner = if_else(team_prob > 0.5, team, otherteam),
-           predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid),
-           game = paste0(team, " vs. ", otherteam))
-  
-  return(df)
-}
-
-logistic_predictor_fn <- function (team_1_id, team_2_id) {
-  stats_team_1 <- filter(s2021, TeamID == team_1_id)
-  stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
-                         rename_all(~ paste0("opposing",.)) 
-  bound_cols <- bind_cols(stats_team_1, stats_team_2)
-  pred <- predict(logistic_fit, newdata = bound_cols, type = "response")
-  
-  return(pred)
-}
-
-logistic_predictor <- purrr::possibly(.f = logistic_predictor_fn, otherwise = "ERROR")

### 2021-04-17 R/independent_modeling.R

@@ -1,187 +1,0 @@
-# source(here::here("R", "cleaningScript.R"))
-
-library(tidymodels)
-source(here::here("R", "helper_functions.R"))
-ids <- readr::read_csv(here::here("data", "ids.csv"))
-s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
-merged <- readr::read_csv(here::here("data", "merged.csv"))
-team_stats <- readr::read_csv(here::here("data", "team_stats.csv"))
-begining_bracket <- readr::read_csv(here::here("data", "begining_bracket.csv"))
-options(tibble.print_min = 20)
-
-# TODO: recipes::step_normalize() try that see if there is a difference
-
-# logsitic fit ------------------------------------------------------------
-
-# logistic_fit <- glm(win ~ x3fg +
-#                       opposingx3fg +
-#                       # field goal pct
-#                       fg_percent +
-#                       opposingfg_percent +
-#                       # free throws
-#                       ft_percent +
-#                       opposingft_percent +
-#                       # rebound per game
-#                       rpg +
-#                       opposingrpg +
-#                       # steels
-#                       st +
-#                       opposingst +
-#                       #turnover
-#                       to +
-#                       opposingto +
-#                       # blocks
-#                       opposingbkpg +
-#                       bkpg,
-#                     data = merged, family = "binomial")
-# 
-# 
-# start <- Sys.time()
-# 
-# logistic_predictions <- ids %>%
-#   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
-#                                     .y = ids$otherid,
-#                                     .f = logistic_predictor)) %>%
-#    mutate(
-#     predicted_winner = if_else(predicted_probs > .5, team, other_team),
-#     predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
-#     predicted_loser_probs = 1 - predicted_winner_probs,
-#     prediction_points = predicted_winner_probs - predicted_loser_probs,
-#     .after = game
-#   )
-# paste0("runtime: ", round(Sys.time() - start, 2))
-# 
-# saveRDS(logistic_predictions, here::here("cache", "logistic_predictions.RDS"))
-# 
-
-logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
-logistic_predictions %>% 
[…]

### 2021-04-17 R/logistic_dependent_model.R

@@ -1,116 +1,0 @@
-source(here::here("R", 'cleaningScript.R'))
-source(here::here("R", "helper_functions.R"))
-theme_set(theme_test())
-options(tibble.print_min = 38)
-
-# merged <- readr::read_csv(here::here("data", "merged.csv"))
-
-# model selection ---------------------------------------------------------
-
-fit <- glm(win ~ x3fg +
-         opposingx3fg +
-# field goal pct
-           fg_percent +
-   opposingfg_percent +
-# free throws
-           ft_percent +
-   opposingft_percent +
-# rebound per game
-                  rpg + 
-          opposingrpg +
-# steels
-                   st +
-           opposingst +
-#turnover
-                   to +
-           opposingto +
-# blocks
-         opposingbkpg +
-                 bkpg,
-           data = merged, family = "binomial")
-
-# First round -------------------------------------------------------------
-
-begining_bracket
-
-probs <- map2_dbl(.x = begining_bracket$teamid, 
-                  .y = begining_bracket$otherteamid, 
-                  .f = logistic_predictor)
-
-
-begining_bracket <- add_probs(begining_bracket, probs = probs)
-
-nice_format(begining_bracket)
-
-# second round ------------------------------------------------------------
-
-second_round  <- advance_round(begining_bracket)
-
-second_round_games <- align_teams(second_round)
-
-probs <- map2_dbl(.x = second_round_games$teamid, 
-                  .y = second_round_games$otherteamid, 
-                  .f = logistic_predictor)
-
-second_round <- add_probs(second_round_games, probs)
-
-nice_format(second_round)
-
-# sweet sixteen -------------------------------------------------------------
-
-sweet_sixteen <- advance_round(second_round)
-
-sweet_sixteen_games <- align_teams(sweet_sixteen)
-
-sweet_sixteen_probs <- map2_dbl(.x = sweet_sixteen_games$teamid, 
-                                .y = sweet_sixteen_games$otherteamid, 
-                                .f = logistic_predictor)
-
-sweet_sixteen_prediction <- add_probs(sweet_sixteen_games, probs = sweet_sixteen_probs)
-
-nice_format(sweet_sixteen_prediction)
-
-
-# elite eight -------------------------------------------------------------
-
-elite_eight <- advance_round(sweet_sixteen_prediction)
-
-elite_eight_games <- align_teams(elite_eight)
-
-elite_eight_probs <- map2_dbl(.x = elite_eight_games$teamid, 
-                                .y = elite_eight_games$otherteamid, 
[…]

### 2021-04-17 R/model_valadation.R

@@ -1,88 +1,0 @@
-library(tidymodels)
-
-options(tibble.print_max = 50, tibble.print_min = 30)
-
-results <- readr::read_csv(here::here("data", "results.csv"))
-logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
-posson_prediction <- readRDS(here::here("cache", "posson_prediction.RDS"))
-multinom_prediction <- readRDS(here::here("cache", "multinom_prediction.RDS"))
-
-
-gameplayed <- results %>% 
-  mutate(gameid = paste0(teamid, "vs", otherteamid))
-
-
-# logistic scoring --------------------------------------------------------
-
-toscore <- function (df) {
-  df %>% 
-    mutate(gameid = paste0(teamid, "vs", otherid)) %>% 
-    filter(gameid %in% pull(gameplayed, gameid)) %>% 
-    left_join(gameplayed, by = c("team", "game", "teamid", "gameid"))
-  
-}
-
-logistic_model_score <- toscore(logistic_predictions) %>% 
-  mutate(score = if_else(predicted_winner == winner, predicted_winner_probs, -1*predicted_winner_probs)) 
-
-# saveRDS(logistic_model_score, here::here("cache", "logistic_model_score.RDS"))
-
-
-logistic_model_score %>% 
-summarise(mean = mean(score, na.rm = TRUE),
-             sd = sd(score, na.rm = TRUE),
-            t_stat = mean / sd, 
-            p = pt(t_stat, n() - 1, lower.tail = FALSE))
-
-
-
-# poisson scoring ---------------------------------------------------------
-
-posson_model_score <- toscore(posson_prediction) %>% 
-  mutate(predicted_difference = predicted_score - other_predicted_score,
-         actual_difference = teamscore - otherteamscore,
-         score = predicted_difference - actual_difference) %>% 
-  mutate(predicted_winner = if_else(predicted_difference > 0, team, other_team),
-         outcome = if_else(predicted_winner == winner, "Correct", "Incorrect")) 
-
-
-# saveRDS(posson_model_score, here::here("cache", "posson_model_score.RDS"))
-
-
-posson_model_score %>% 
-  summarise(mean = mean(score, na.rm = TRUE),
-            sd = sd(score, na.rm = TRUE),
-            t_stat = mean / sd, 
-            p = pt(t_stat, n() - 1, lower.tail = FALSE))
-
[…]

### 2021-04-17 R/scrape_finals.R

@@ -1,55 +1,0 @@
-
-# Scrape finals scores ----------------------------------------------------
-
-library(rvest)
-library(dplyr)
-
-options(tibble.print_max = 50, tibble.print_min = 50)
-
-url <- "https://www.ncaa.com/march-madness-live/scores"
-
-scraped_final_scores <- read_html(url) %>% 
-  html_nodes('.lvp') %>% 
-  html_text(trim = TRUE)
-
-games  <- tibble::tibble(scraped_final_scores[12:279]) %>% 
-    rename(game = 1) %>% 
-  mutate(id = rep(c("a", "b"), 268/2)) # this will break when game is done
-
-spliter <- function(df, tag){
-  df  %>% 
-    filter(id == tag)
-}
-
-a <- spliter(games, "a")
-b <- spliter(games, "b")
-
-MTeamSpellings <- readr::read_csv("rawdata/MTeamSpellings.csv") %>% 
-  mutate(TeamID = as.character(TeamID), 
-           team = TeamNameSpelling)
-
-scores <- bind_cols(a, b) %>% 
-  select(1, 3) %>% 
-  rename(team = 1, score = 2) %>% 
-  mutate(team = stringr::str_to_lower(team)) %>% 
-  left_join(MTeamSpellings, by = "team") %>% 
-  select(1, 2, 4) %>% 
-  mutate(id = rep(c("a", "b"), 134/2)) # this will break when game is done
-  
-a <- spliter(scores, "a")
-b <- spliter(scores, "b")
-
-bind_cols(a, b) %>% 
-  rename(team = 1, 
-            otherteam = 5,
-            teamscore = 2, 
-            otherteamscore = 6, 
-            teamid = 3, 
-         otherteamid = 7) %>% 
-  select(1:3, 5:7) %>%
-  
-  mutate(team = gsub("[^[:alnum:]]", "", team),
-         otherteam = gsub("[^[:alnum:]]", "", otherteam),
-    game = paste0(team, " vs. ", otherteam),
-         winner = if_else(teamscore > otherteamscore, team, otherteam)) %>% 
-  readr::write_csv(here::here("data", "results.csv"))

### 2021-04-17 R/testing_normalized.R

@@ -1,221 +1,0 @@
-# source(here::here("R", "cleaningScript.R"))
-
-library(tidymodels)
-source(here::here("R", "helper_functions.R"))
-ids <- readr::read_csv(here::here("data", "ids.csv"))
-s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
-merged <- readr::read_csv(here::here("data", "merged.csv"))
-team_stats <- readr::read_csv(here::here("data", "team_stats.csv"))
-begining_bracket <- readr::read_csv(here::here("data", "begining_bracket.csv"))
-options(tibble.print_min = 20)
-
-
-recipe(water + fat + protein ~ ., data = meat_tr) %>% 
-  step_normalize(all_predictors())
-
-
-normaliser <- function(x){
- output <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
- return(output)
-}
-
-normalized <- merged %>% 
-  select( x3fg ,
-            opposingx3fg ,
-            # field goal pct
-            fg_percent ,
-            opposingfg_percent ,
-            # free throws
-            ft_percent ,
-            opposingft_percent ,
-            # rebound per game
-            rpg ,
-            opposingrpg ,
-            # steels
-            st ,
-            opposingst ,
-            #turnover
-            to ,
-            opposingto ,
-            # blocks
-            opposingbkpg ,
-            bkpg) %>% 
-  map_df(normaliser) %>% 
-  bind_cols(win = merged$win)
-
-logistic_fit <- glm(win ~ x3fg +
-      opposingx3fg +
-      # field goal pct
-      fg_percent +
-      opposingfg_percent +
-      # free throws
-      ft_percent +
-      opposingft_percent +
-      # rebound per game
-      rpg +
-      opposingrpg +
-      # steels
-      st +
-      opposingst +
-      #turnover
-      to +
-      opposingto +
-      # blocks
-      opposingbkpg +
-      bkpg,
-    data = normalized, family = "binomial")
-  
-# TODO: recipes::step_normalize() try that see if there is a difference
-
-
-# logsitic fit ------------------------------------------------------------
-
-start <- Sys.time()
-
-logistic_predictions <- ids %>%
-  mutate(predicted_probs = map2_dbl(.x = ids$teamid,
-                                    .y = ids$otherid,
-                                    .f = logistic_predictor)) %>%
-   mutate(
-    predicted_winner = if_else(predicted_probs > .5, team, other_team),
[…]

### 2021-04-17 R/bracket.R
updating files from meeting

@@ -1,0 +1,89 @@
+library(tidyverse)
+# https://plexkits.com/march-madness-bracket/ I think this will update
+# https://docs.google.com/spreadsheets/d/1zaWZ2Xh7sll-PZKA1DhKvS6TdfE0Yggk0m5OWY4aICc/edit#gid=0
+
+team_names <- readr::read_csv(here::here("rawdata", "MTeamSpellings.csv"))
+bracket <- readxl::read_excel(here::here("data", "prebracket.xlsx"))
+begining_bracket <- readxl::read_excel(here::here("data", "begining_bracket.xlsx"))
+
+# Even better
+# bracket <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1zaWZ2Xh7sll-PZKA1DhKvS6TdfE0Yggk0m5OWY4aICc/edit#gid=0")
+# remember API times out...
+
+bracket <- bracket %>% 
+  select(2, 35)
+  
+colnames(bracket) <- c("x1", "x2")
+
+
+# bracket -----------------------------------------------------------------
+
+
+bracket <- stack(bracket) %>% 
+  select(values) %>% 
+  drop_na() %>% 
+  tibble() %>% 
+  rename(team = values) %>% 
+  mutate(team = tolower(team)) %>% 
+  mutate(game = 1:n(),
+         game = as.numeric(game),
+         otherteam  = if_else(game %% 2 == 1, team, lag(team)),
+         game = if_else(game %% 2 == 1, game + 1, game), .before = team) %>% 
+  filter(team != otherteam) %>% 
+  separate(team, into = c("team", "pocket"), "/") %>% 
+  filter(otherteam != "first round") %>% 
+  left_join(team_names, by = c("team" = "TeamNameSpelling")) %>% 
+  rename(teamid = TeamID) %>% 
+  left_join(team_names, by = c("otherteam" = "TeamNameSpelling")) %>%
+  rename(otherteamid = TeamID) %>% 
+  mutate(teamid = replace(teamid, team == "app st.", 1111)) %>% 
+  mutate_if(is.numeric, as.character)
+
+bracket
+
+# pockets -----------------------------------------------------------------
+
+pocket <- bracket %>% 
+  select(team, pocket) %>% 
+  drop_na() %>% 
+  left_join(team_names, by = c("team" = "TeamNameSpelling")) %>% 
+  rename(teamid = TeamID, otherteam = pocket) %>% 
[…]

### 2021-04-17 R/cleaningScript.R

@@ -1,0 +1,183 @@
+library(tidyverse) 
+# options(tibble.print_min = 35)
+list_of_dfs <- sapply(paste0("rawdata/", dir("rawdata")), read_csv, USE.NAMES = TRUE)
+
+games <- list_of_dfs$`rawdata/MConferenceTourneyGames.csv` %>% 
+  transmute(gameid = paste0(DayNum, WTeamID, LTeamID),
+            Season, ConfAbbrev) 
+
+seasionsoutcomes <- list_of_dfs$`rawdata/MRegularSeasonDetailedResults.csv` %>% 
+  left_join(list_of_dfs$`rawdata/MTeams.csv`, by = c("WTeamID" = "TeamID")) %>% 
+  filter(Season %in% c(2019, 2020, 2021)) %>% 
+  rename(WTeamName = TeamName) %>% 
+  left_join( list_of_dfs$`rawdata/MTeams.csv`, by = c("LTeamID" = "TeamID")) %>% 
+  select(1:7) %>% 
+  mutate(W = WTeamID, L =  LTeamID) %>% 
+  pivot_longer(cols = ends_with("ID"),
+               names_to = "outcome", 
+               values_to = "TeamID") %>% 
+  mutate(win = if_else(outcome == "WTeamID", 1 ,0),
+         TeamID = as.character(TeamID),
+         Season = as.character(Season)) %>% 
+  select(-outcome, -DayNum) %>% 
+  rename(season = Season)
+
+seasionsoutcomes
+
+stats_fn <- function (directory) {
+  stats <- sapply(paste0(directory, "/", dir(directory)),
+                  readxl::read_excel, USE.NAMES = TRUE)
+  stats <- plyr::join_all(stats, by= "Team")
+  stats <- stats[, !duplicated(colnames(stats), fromLast = TRUE)]
+  return(stats)
+}
+
+stats <- sapply(list.dirs("se_data")[-1], stats_fn)
+selector <- function (df, cols = co) {
+  grabbed <- select(df, contains(cols))
+  return(grabbed)
+}
+
+a <- colnames(stats$`se_data/se_2019`)
+b <- colnames(stats$`se_data/se_2020`)
+c <- colnames(stats$`se_data/se_2021`)
+
+co <- as.data.frame(table(c(a, b, c))) %>% 
+  filter(Freq == 3) %>% 
[…]

### 2021-04-17 R/explatoryAnalysis.R

@@ -1,0 +1,77 @@
+library(tidyverse)
+
+list_of_dfs <- sapply(paste0("rawdata/", dir("rawdata")), read_csv, USE.NAMES = TRUE)
+
+# https://www.kaggle.com/c/ncaam-march-mania-2021/data
+lapply(list_of_dfs, colnames)
+lapply(list_of_dfs, glimpse)
+
+# Keys --------------------------------------------------------------------
+# give us info about the data
+
+list_of_dfs$`rawdata/MSeasons.csv`
+# identifies the different seasons included in the historical data, 
+# along with certain season-level properties.
+
+list_of_dfs$`rawdata/MTeams.csv` # JOINED
+# identifies the id number of the team
+
+list_of_dfs$`rawdata/MTeamConferences.csv`
+# indicates the conference affiliations for 
+# each team during each season. 
+
+list_of_dfs$`rawdata/MTeamCoaches.csv`
+# indicates the head coach for each team in each season
+
+list_of_dfs$`rawdata/MNCAATourneySlots.csv`
+# bracket positions
+
+list_of_dfs$`rawdata/MNCAATourneySeedRoundSlots.csv`
+# seeds with slot and day and late
+
+list_of_dfs$`rawdata/MMasseyOrdinals.csv` 
+# lists out team rankings 
+
+list_of_dfs$`rawdata/Conferences.csv`
+# indicates the Division I conferences
+
+list_of_dfs$`rawdata/MTeamSpellings.csv`
+# indicates alternative spellings of many team names.
+
+list_of_dfs$`rawdata/Cities.csv`
+# This file provides a master list of cities that have
+# been locations for games played.
+
+# Data --------------------------------------------------------------------
+# This is information about games played
+
+list_of_dfs$`rawdata/MNCAATourneyDetailedResults.csv`
+# provides team-level box scores for many NCAA® tournaments
+
+list_of_dfs$`rawdata/MNCAATourneyCompactResults.csv`
+#identifies the game-by-game NCAA® tournament results 
+#for all seasons of historical data.
+
+list_of_dfs$`rawdata/MRegularSeasonDetailedResults.csv`
+# provides team-level box scores for many regular seasons 
+# of historical data, starting with the 2003 season. 
+
+list_of_dfs$`rawdata/MRegularSeasonCompactResults.csv`
+# identifies the game-by-game results for many 
+# seasons of historical data, starting with the 1985 season 
+
+# Probally not useful -----------------------------------------------------
[…]

### 2021-04-17 R/fitting.R

@@ -1,0 +1,68 @@
+logistic_fit <- glm(win ~ x3fg +
+                       opposingx3fg +
+                       # field goal pct
+                       fg_percent +
+                       opposingfg_percent +
+                       # free throws
+                       ft_percent +
+                       opposingft_percent +
+                       # rebound per game
+                       rpg +
+                       opposingrpg +
+                       # steels
+                       st +
+                       opposingst +
+                       #turnover
+                       to +
+                       opposingto +
+                       # blocks
+                       opposingbkpg +
+                       bkpg,
+                     data = merged, family = "binomial")
+
+logistic_fit_null <- glm(win ~ 1,
+                    data = merged, family = "binomial") 
+
+ fit_poisson <- glm(team_score ~ x3fg +
+              opposingx3fg +
+              # field goal pct
+              fg_percent +
+              opposingfg_percent +
+              # free throws
+              ft_percent +
+              opposingft_percent +
+              # rebound per game
+              rpg +
+              opposingrpg +
+              # steels
+              st +
+              opposingst +
+              #turnover
+              to +
+              opposingto +
+              # blocks
+              opposingbkpg +
+              bkpg,
+              data = merged, family = poisson(link = "log"))
+ 
+ fit_multinom <- VGAM::vglm(x_tile ~ x3fg +
+                  opposingx3fg +
+                  # field goal pct
+                  fg_percent +
+                  opposingfg_percent +
+                  # free throws
+                  ft_percent +
+                  opposingft_percent +
+                  # rebound per game
+                  rpg +
+                  opposingrpg +
+                  # steels
+                  st +
+                  opposingst +
+                  #turnover
+                  to +
+                  opposingto +
+                  # blocks
+                  opposingbkpg +
+                  bkpg, family = VGAM::cumulative(parallel=TRUE),
+                  data = merged)

### 2021-04-17 R/helper_functions.R

@@ -1,0 +1,55 @@
+align_teams <- function(df){
+  
+  team <- filter(df, game != "team")
+  other <- filter(df, game == "team")
+  
+  output <- other %>% 
+    rename_all(~ paste0("other",.)) %>% 
+    bind_cols(team) 
+  
+  return(output)
+}
+
+advance_round <- function(df){
+  df <- df %>% 
+    select(predicted_winner, predicted_winnerid) %>% 
+    rename(team = predicted_winner, teamid = predicted_winnerid) %>% 
+    mutate(game = rep(c("team", "otherteam"), nrow(df)/2))
+  
+  return(df)
+}
+
+nice_format <- function(df){
+  output <- df %>%
+    mutate(pred_winner_prob = if_else(team_prob > 0.5, team_prob, otherteam_prob),
+           pred_loser_prob = if_else(team_prob < 0.5, team_prob, otherteam_prob)) %>% 
+    select(game, predicted_winner, 
+           pred_winner_prob, pred_loser_prob)
+  
+  return(output)
+}
+
+add_probs <- function(df, probs){
+  
+  df <- df %>% 
+    mutate(team_prob = probs,
+           otherteam_prob = 1 - team_prob, 
+           game = paste0(team, " vs. ", otherteam),
+           predicted_winner = if_else(team_prob > 0.5, team, otherteam),
+           predicted_winnerid = if_else(team_prob > 0.5, teamid, otherteamid),
+           game = paste0(team, " vs. ", otherteam))
+  
+  return(df)
+}
+
+logistic_predictor_fn <- function (team_1_id, team_2_id) {
+  stats_team_1 <- filter(s2021, TeamID == team_1_id)
+  stats_team_2 <- filter(s2021, TeamID == team_2_id) %>% 
+                         rename_all(~ paste0("opposing",.)) 
+  bound_cols <- bind_cols(stats_team_1, stats_team_2)
+  pred <- predict(logistic_fit, newdata = bound_cols, type = "response")
+  
+  return(pred)
+}
+
+logistic_predictor <- purrr::possibly(.f = logistic_predictor_fn, otherwise = "ERROR")

### 2021-04-17 R/independent_modeling.R

@@ -1,0 +1,187 @@
+# source(here::here("R", "cleaningScript.R"))
+
+library(tidymodels)
+source(here::here("R", "helper_functions.R"))
+ids <- readr::read_csv(here::here("data", "ids.csv"))
+s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+team_stats <- readr::read_csv(here::here("data", "team_stats.csv"))
+begining_bracket <- readr::read_csv(here::here("data", "begining_bracket.csv"))
+options(tibble.print_min = 20)
+
+# TODO: recipes::step_normalize() try that see if there is a difference
+
+# logsitic fit ------------------------------------------------------------
+
+# logistic_fit <- glm(win ~ x3fg +
+#                       opposingx3fg +
+#                       # field goal pct
+#                       fg_percent +
+#                       opposingfg_percent +
+#                       # free throws
+#                       ft_percent +
+#                       opposingft_percent +
+#                       # rebound per game
+#                       rpg +
+#                       opposingrpg +
+#                       # steels
+#                       st +
+#                       opposingst +
+#                       #turnover
+#                       to +
+#                       opposingto +
+#                       # blocks
+#                       opposingbkpg +
+#                       bkpg,
+#                     data = merged, family = "binomial")
+# 
+# 
+# start <- Sys.time()
+# 
+# logistic_predictions <- ids %>%
+#   mutate(predicted_probs = map2_dbl(.x = ids$teamid,
+#                                     .y = ids$otherid,
+#                                     .f = logistic_predictor)) %>%
+#    mutate(
+#     predicted_winner = if_else(predicted_probs > .5, team, other_team),
+#     predicted_winner_probs = if_else(predicted_probs > .5, predicted_probs , 1 - predicted_probs),
+#     predicted_loser_probs = 1 - predicted_winner_probs,
+#     prediction_points = predicted_winner_probs - predicted_loser_probs,
+#     .after = game
+#   )
+# paste0("runtime: ", round(Sys.time() - start, 2))
+# 
+# saveRDS(logistic_predictions, here::here("cache", "logistic_predictions.RDS"))
+# 
+
+logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
+logistic_predictions %>% 
[…]

### 2021-04-17 R/logistic_dependent_model.R

@@ -1,0 +1,116 @@
+source(here::here("R", 'cleaningScript.R'))
+source(here::here("R", "helper_functions.R"))
+theme_set(theme_test())
+options(tibble.print_min = 38)
+
+# merged <- readr::read_csv(here::here("data", "merged.csv"))
+
+# model selection ---------------------------------------------------------
+
+fit <- glm(win ~ x3fg +
+         opposingx3fg +
+# field goal pct
+           fg_percent +
+   opposingfg_percent +
+# free throws
+           ft_percent +
+   opposingft_percent +
+# rebound per game
+                  rpg + 
+          opposingrpg +
+# steels
+                   st +
+           opposingst +
+#turnover
+                   to +
+           opposingto +
+# blocks
+         opposingbkpg +
+                 bkpg,
+           data = merged, family = "binomial")
+
+# First round -------------------------------------------------------------
+
+begining_bracket
+
+probs <- map2_dbl(.x = begining_bracket$teamid, 
+                  .y = begining_bracket$otherteamid, 
+                  .f = logistic_predictor)
+
+
+begining_bracket <- add_probs(begining_bracket, probs = probs)
+
+nice_format(begining_bracket)
+
+# second round ------------------------------------------------------------
+
+second_round  <- advance_round(begining_bracket)
+
+second_round_games <- align_teams(second_round)
+
+probs <- map2_dbl(.x = second_round_games$teamid, 
+                  .y = second_round_games$otherteamid, 
+                  .f = logistic_predictor)
+
+second_round <- add_probs(second_round_games, probs)
+
+nice_format(second_round)
+
+# sweet sixteen -------------------------------------------------------------
+
+sweet_sixteen <- advance_round(second_round)
+
+sweet_sixteen_games <- align_teams(sweet_sixteen)
+
+sweet_sixteen_probs <- map2_dbl(.x = sweet_sixteen_games$teamid, 
+                                .y = sweet_sixteen_games$otherteamid, 
+                                .f = logistic_predictor)
+
+sweet_sixteen_prediction <- add_probs(sweet_sixteen_games, probs = sweet_sixteen_probs)
+
+nice_format(sweet_sixteen_prediction)
+
+
+# elite eight -------------------------------------------------------------
+
+elite_eight <- advance_round(sweet_sixteen_prediction)
+
+elite_eight_games <- align_teams(elite_eight)
+
+elite_eight_probs <- map2_dbl(.x = elite_eight_games$teamid, 
+                                .y = elite_eight_games$otherteamid, 
[…]

### 2021-04-17 R/model_valadation.R

@@ -1,0 +1,88 @@
+library(tidymodels)
+
+options(tibble.print_max = 50, tibble.print_min = 30)
+
+results <- readr::read_csv(here::here("data", "results.csv"))
+logistic_predictions <- readRDS(here::here("cache", "logistic_predictions.RDS"))
+posson_prediction <- readRDS(here::here("cache", "posson_prediction.RDS"))
+multinom_prediction <- readRDS(here::here("cache", "multinom_prediction.RDS"))
+
+
+gameplayed <- results %>% 
+  mutate(gameid = paste0(teamid, "vs", otherteamid))
+
+
+# logistic scoring --------------------------------------------------------
+
+toscore <- function (df) {
+  df %>% 
+    mutate(gameid = paste0(teamid, "vs", otherid)) %>% 
+    filter(gameid %in% pull(gameplayed, gameid)) %>% 
+    left_join(gameplayed, by = c("team", "game", "teamid", "gameid"))
+  
+}
+
+logistic_model_score <- toscore(logistic_predictions) %>% 
+  mutate(score = if_else(predicted_winner == winner, predicted_winner_probs, -1*predicted_winner_probs)) 
+
+# saveRDS(logistic_model_score, here::here("cache", "logistic_model_score.RDS"))
+
+
+logistic_model_score %>% 
+summarise(mean = mean(score, na.rm = TRUE),
+             sd = sd(score, na.rm = TRUE),
+            t_stat = mean / sd, 
+            p = pt(t_stat, n() - 1, lower.tail = FALSE))
+
+
+
+# poisson scoring ---------------------------------------------------------
+
+posson_model_score <- toscore(posson_prediction) %>% 
+  mutate(predicted_difference = predicted_score - other_predicted_score,
+         actual_difference = teamscore - otherteamscore,
+         score = predicted_difference - actual_difference) %>% 
+  mutate(predicted_winner = if_else(predicted_difference > 0, team, other_team),
+         outcome = if_else(predicted_winner == winner, "Correct", "Incorrect")) 
+
+
+# saveRDS(posson_model_score, here::here("cache", "posson_model_score.RDS"))
+
+
+posson_model_score %>% 
+  summarise(mean = mean(score, na.rm = TRUE),
+            sd = sd(score, na.rm = TRUE),
+            t_stat = mean / sd, 
+            p = pt(t_stat, n() - 1, lower.tail = FALSE))
+
[…]

### 2021-04-17 R/scrape_finals.R

@@ -1,0 +1,55 @@
+
+# Scrape finals scores ----------------------------------------------------
+
+library(rvest)
+library(dplyr)
+
+options(tibble.print_max = 50, tibble.print_min = 50)
+
+url <- "https://www.ncaa.com/march-madness-live/scores"
+
+scraped_final_scores <- read_html(url) %>% 
+  html_nodes('.lvp') %>% 
+  html_text(trim = TRUE)
+
+games  <- tibble::tibble(scraped_final_scores[12:279]) %>% 
+    rename(game = 1) %>% 
+  mutate(id = rep(c("a", "b"), 268/2)) # this will break when game is done
+
+spliter <- function(df, tag){
+  df  %>% 
+    filter(id == tag)
+}
+
+a <- spliter(games, "a")
+b <- spliter(games, "b")
+
+MTeamSpellings <- readr::read_csv("rawdata/MTeamSpellings.csv") %>% 
+  mutate(TeamID = as.character(TeamID), 
+           team = TeamNameSpelling)
+
+scores <- bind_cols(a, b) %>% 
+  select(1, 3) %>% 
+  rename(team = 1, score = 2) %>% 
+  mutate(team = stringr::str_to_lower(team)) %>% 
+  left_join(MTeamSpellings, by = "team") %>% 
+  select(1, 2, 4) %>% 
+  mutate(id = rep(c("a", "b"), 134/2)) # this will break when game is done
+  
+a <- spliter(scores, "a")
+b <- spliter(scores, "b")
+
+bind_cols(a, b) %>% 
+  rename(team = 1, 
+            otherteam = 5,
+            teamscore = 2, 
+            otherteamscore = 6, 
+            teamid = 3, 
+         otherteamid = 7) %>% 
+  select(1:3, 5:7) %>%
+  
+  mutate(team = gsub("[^[:alnum:]]", "", team),
+         otherteam = gsub("[^[:alnum:]]", "", otherteam),
+    game = paste0(team, " vs. ", otherteam),
+         winner = if_else(teamscore > otherteamscore, team, otherteam)) %>% 
+  readr::write_csv(here::here("data", "results.csv"))

### 2021-04-17 R/testing_normalized.R

@@ -1,0 +1,221 @@
+# source(here::here("R", "cleaningScript.R"))
+
+library(tidymodels)
+source(here::here("R", "helper_functions.R"))
+ids <- readr::read_csv(here::here("data", "ids.csv"))
+s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
+merged <- readr::read_csv(here::here("data", "merged.csv"))
+team_stats <- readr::read_csv(here::here("data", "team_stats.csv"))
+begining_bracket <- readr::read_csv(here::here("data", "begining_bracket.csv"))
+options(tibble.print_min = 20)
+
+
+recipe(water + fat + protein ~ ., data = meat_tr) %>% 
+  step_normalize(all_predictors())
+
+
+normaliser <- function(x){
+ output <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
+ return(output)
+}
+
+normalized <- merged %>% 
+  select( x3fg ,
+            opposingx3fg ,
+            # field goal pct
+            fg_percent ,
+            opposingfg_percent ,
+            # free throws
+            ft_percent ,
+            opposingft_percent ,
+            # rebound per game
+            rpg ,
+            opposingrpg ,
+            # steels
+            st ,
+            opposingst ,
+            #turnover
+            to ,
+            opposingto ,
+            # blocks
+            opposingbkpg ,
+            bkpg) %>% 
+  map_df(normaliser) %>% 
+  bind_cols(win = merged$win)
+
+logistic_fit <- glm(win ~ x3fg +
+      opposingx3fg +
+      # field goal pct
+      fg_percent +
+      opposingfg_percent +
+      # free throws
+      ft_percent +
+      opposingft_percent +
+      # rebound per game
+      rpg +
+      opposingrpg +
+      # steels
+      st +
+      opposingst +
+      #turnover
+      to +
+      opposingto +
+      # blocks
+      opposingbkpg +
+      bkpg,
+    data = normalized, family = "binomial")
+  
+# TODO: recipes::step_normalize() try that see if there is a difference
+
+
+# logsitic fit ------------------------------------------------------------
+
+start <- Sys.time()
+
+logistic_predictions <- ids %>%
+  mutate(predicted_probs = map2_dbl(.x = ids$teamid,
+                                    .y = ids$otherid,
+                                    .f = logistic_predictor)) %>%
+   mutate(
+    predicted_winner = if_else(predicted_probs > .5, team, other_team),
[…]

### 2021-04-17 slides/slides.Rmd
added citation
adding final slides

@@
 ---
-title: "March Madness Basketball Turniment"
+title: "March Madness Basketball Tournament"
 author: "Sabrina Ball and Joe Despres"
@@
 ### These predictions are not wildly different than betting markets and seeds because they use similar statistics to derive their predictions.
+
+---
+class: inverse
+
+# References
+
+Faraway, Julian James. Extending the Linear Model with R:  Generalized Linear,    Mixed Effects and Nonparametric Regression Models.        Chapman &amp; Hall/CRC, 2016.

### 2021-04-18 paper/formula.rmd
worked on paper

---
title: ""
subtitle: ""
author: ""
institute: ""
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r eval=FALSE, echo=FALSE}
equatiomatic::extract_eq(model = logistic_fit, 
                         use_coefs = FALSE, 
                         show_distribution = FALSE,
                         intercept = "beta")
# results='asis' in code chunk to render as LaTex
```

$$
\begin{align}
\beta_{0} + \beta_{1}(\operatorname{x3fg}) + \beta_{2}(\operatorname{opposingx3fg}) + \beta_{3}(\operatorname{fg\_percent}) + \beta_{4}(\operatorname{opposingfg\_percent}) + \beta_{5}(\operatorname{ft\_percent}) + \beta_{6}(\operatorname{opposingft\_percent}) + \beta_{7}(\operatorname{rpg}) + \beta_{8}(\operatorname{opposingrpg}) + \beta_{9}(\operatorname{st}) + \beta_{10}(\operatorname{opposingst}) + \beta_{11}(\operatorname{to}) + \beta_{12}(\operatorname{opposingto}) + \beta_{13}(\operatorname{opposingbkpg}) + \beta_{14}(\operatorname{bkpg})
\begin{align}
$$

---
$$
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$

### 2021-04-18 paper/paper.tex
worked on paper

@@
   man,floatsintext]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
@@
 \usepackage{csquotes}
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \caption{\label{tab:unnamed-chunk-1}Descriptive Statistics}\\
-\toprule{}
+\toprule
 Variable & Mean & Median & Std & Min & Max & Range\\
-\midrule{}
+\midrule
 Three-Points goals Scored & 242.08 & 241.00 & 48.18 & 107.0 & 454.00 & 347.00\\
@@
 Blocks Per game & 3.30 & 3.20 & 0.94 & 1.3 & 6.80 & 5.50\\
-\bottomrule{}
+\bottomrule
 \end{longtable}
@@
 \caption{\label{tab:unnamed-chunk-5}Regression Output}\\
-\toprule{}
+\toprule
 Terms & Multinomial Model & Logistic Model & Poisson Model\\
-\midrule{}
+\midrule
 <-12 & -1.2516 & -- & --\\
@@
 \multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{3} All the above terms are significant at P < 0.01.}\\
-\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{4} Multinomeal Intercepts are differences in predicted score.}\\{}
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{4} Multinomeal Intercepts are differences in predicted score.}\\
 \end{longtable}
@@
 \hypertarget{refs}{}
-\begin{cslreferences}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-R-purrr}{}%
@@
 (2019).
-Welcome to the tidyverse.
+Welcome to the {tidyverse}.
 \emph{Journal of Open Source Software}, \emph{4}(43), 1686.
@@
 \leavevmode\hypertarget{ref-R-kableExtra}{}%
-Zhu, H. (2020). \emph{KableExtra: Construct complex table with 'kable' and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}
-\end{cslreferences}
+Zhu, H. (2020). \emph{kableExtra: Construct complex table with 'kable' and pipe syntax}. Retrieved from \url{https://CRAN.R-project.org/package=kableExtra}
+
+\end{CSLReferences}
 

### 2021-04-20 paper/paper.Rmd
added paper revisions

@@
 The objective is to predict the individual game outcomes as accurately as possible and determine which GLM is the most accurate.
-Our response terms is the outcome of the game, *win or loss*, without the possibility of a tie.
+Our response terms is the outcome of the game, *win or loss* without the possibility of a tie.
 Our aim is to derive a function that will accurately predict this using team-level statistics.
-In our dataset, we have many predictors however, we only selected the predictors that are not co-linear.
-First, *field goal percentage* which is the ratio of attempted scores to made scores.
-*Free-throw percentage*,  is the the rate of successfully scoring a penalty shot.
+In our dataset, we have many predictors however, we only selected ones that are not co-linear.
+First, *field goal percentage* which is the ratio of attempted scores to successful scores.
+*Free-throw percentage*, is the the rate of scoring a penalty shot.
 Cumulative *Three-point goals made*, field goals made from a sufficient distance.
@@
 *Steals* is when a team was able to remove possession of the ball from the opposing team.
-*Turnover*, the number of times the team lost possession.
-*Blocks*, the number of times the team was able to block a shot made by the opposing team.
+*Turnover*, is the number of times the team lost possession.
+*Blocks*, is the number of times the team was able to block a shot made by the opposing team.
 March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
@@
 We did make predictions using that method, however our models will be compared by treating the games as independent of the previous round.
-Meaning, we will filter our data down to include only the 64 teams that qualify, predict every single possible combination of games ${64\choose 2} = 2016$, then use the 63 games played as a sample from the population of all possible games.
-Then use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
-Then we use the results to determine which regression method performs the best.
-We will use three regression methods, logistic, poisson, and multinomial, to generate prediction models, compare the results, and make a determination as to which is best suited to this problem.
-Logistic regression most naturally suits this problem because games cannot tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
-This problem could also be suited to poison regression because the number of points scored is roughly poisson distributed.
-After fitting a poisson model, we will predict the number of points scored by Team A, predict the number of points scored by Team B, and take the team with the highest predicted score to be the winner.
-A multinomial model is less naturally suited to this problem, however it may address the shortcomings of the logistic model.
-In particular, many games are close scoring and are won by merely a few points.
-Therefore, we will use multinomeal regression to predict the difference in game score by assigning the differences into five categories.
+Meaning, we will filter our data down to include only the 64 teams that qualify, predict every single one of the ${64\choose 2} = 2016$ possible combination, then use the 63 games played as a sample from the population of all possible games.
+After that, we use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
+Then we use the results to determine which regression model performs the best.
+We will employ three regression methods, logistic, Poisson, and multinomial, to generate predictive models, compare the results, and decide which is best suited to this problem.
+Logistic regression most naturally suits this problem because games cannot tie and there is a perfectly equal number of wins and losses.
+This problem could also be suited to Poisson regression because the number of points scored is roughly Poisson distributed.
+After fitting a Poisson model, we will predict the number of points scored by Team A, and Team B.
+Then select the team with the highest predicted score to be the predicted winner.
+A multinomial model is less naturally suited to this problem, however, it may address the shortcomings of the logistic model.
+In particular, many games are close scoring and won by merely a few points.
+Therefore, we will use multinomial regression to predict the difference in-game score by assigning the differences into five categories.
 From there, assign a predicted winner based on which outcome the model considers to be the most likely.
@@
 GLMs are accurate relative to randomly guessing March Madness tournament outcomes.
-Team-level statistics, recorded by the NCAA, are highly significant and helpful predicting the tournament outcomes but do not yield fantastic results.
+Team-level statistics, recorded by the NCAA, are highly statistically significant and helpful in predicting the tournament outcomes but do not yield fantastic results.
 Also, when predicting the probability of a win or a loss using a logit link, the coefficients are symmetric because basketball games are a zero-sum-contest.
-From our results, we see that logistic and poisson models are the best at predicting wins and losses.
+From our results, we see that logistic and Poisson models perform the best at predicting wins and losses.
 All three GLM models are well over 50% accurate, therefore, we claim that using GLM models are more accurate than random chance.
@@
 
-As mentioned above, there are substantially more variables in the dataset than we used, however we omitted the ones that were co-linear.
-When making a decision on which to keep, we selected the variable using deviance and likelihood ratio tests.
-For our modeling and prediction we are going to use the statistics associated with both teams.
+As mentioned above, there are substantially more variables in the dataset than we used, however, we omitted co-linear predictors.
+When deciding on which to keep, we selected the variable using deviance and likelihood ratio tests.
+For our modeling and prediction, we are going to use the statistics associated with both teams.
 Basic descriptive statistics can be found in Table 1.
@@
 The standard deviations can be high, but not substantially high, and the ranges are reasonable for the data we have.
-We remark on the predictors being roughly normally distributed with no substantial skew, heavy tails, or slow decay as shown in Figure 1.
+As shown in Figure 1, we remark on the predictors being roughly normally distributed with no substantial skew, heavy tails, or slow decay..
 
@@
 The logistic model is straight forward, only needing a win to be coded as 1 and a loss to be coded as 0.
-We fit a poisson model by taking a score as a count.
-This is not perfectly suited to poisson because a winning team's score has a mean of `r mean(merged$w_score) %>% round(2)` and a variance of `r var(merged$w_score) %>% round(2)`.
-Regardless we will see how it performs by predicting the score for Team A and Team B, then choose the one with the highest predicted score as the winner.
+We fit a Poisson model by taking a score as a count.
+This is not perfectly suited to Poisson because a winning team's score has a mean of `r mean(merged$w_score) %>% round(2)` and a variance of `r var(merged$w_score) %>% round(2)`.
+Regardless, we will see how it performs by predicting the score for Team A and Team B, then selecting the one with the highest predicted score as the predicted winner.
 
@@
 
-The multinomeal case is not obviously suited to this problem, however, we can adapt it.
+The multinomeal case is not suited to this problem, however, we can adapt it.
 Commonly seen in betting markets, we will take a look at the difference in score between winner and loser.
 Then assign it into five categories using the quintiles as shown in Figure 2.
-First category, is the probability of the team losing by more than twelve points.
+The first category is the probability of the team losing by more than twelve points.
 Second, the probability of the team losing between twelve and four points.
@@
 To make an accurate comparison, we use the same formula for the three models with only the dependent variable being different.
-Win or loss for our logistic model, the amount Team A scores for the poisson model, and the difference in score for the multinomeal model.
+Win or loss for our logistic model, the amount Team A scores for the Poisson model, and the difference in score for the multinomeal model.
 We tried normalizing the data by subtracting the mean and dividing by the standard deviation for all predictors.
-However, that really did not have an effect on the prediction accuracy.
+However, that did not affect the prediction accuracy.
 Also, we individually tested each predictor using the likelihood ratio test before adding each term.
@@
 
-In Table 2, made manually with the kableExtra package [@R-kableExtra], you can see that we have the results for the poisson, logistic, and multinomial models.
-We have a sample of over 35,000 and all of the coefficients are highly significant.
-Which is to be expected because these are the statistics that the NCAA collects as the metrics useful in measuring a team's ability to win games.
-The aim of this study is to compare predictive models so we will not cover it exhaustively or include individual z-statistics and p-values.
-First notice, in the case of the logistic and multinomial models we see that when comparing factors that affect a team's probability of winning, the associated coefficient is nearly equal to the factor capturing the opposing teams metric.
+In Table 2, made with the assistance of the kableExtra package [@R-kableExtra], you can see that we have the results for the Poisson, logistic, and multinomial models.
+We have a sample of over 35,000 and all of the coefficients are highly significant, which is to be expected because these are the statistics that the NCAA collects as the metrics useful in measuring a team's ability to win games.
+This study aims to compare predictive models, so we will not cover it exhaustively or include individual z-statistics and p-values.
+First, the case of the logistic and multinomial models we see that when comparing factors that affect a team's probability of winning, the associated coefficient is nearly equal to the factor capturing the opposing team's metric.
 That is because basketball is a zero-sum-game, anything good for team A is proportionately bad for team B.
-Note this is not true for the poisson model because that is measuring points scored.
-Take three pointers in the poisson model, for instance, where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
+Note this is not true for the poisson model because that is measuring points scored rather than estimating probabilities.
+Take three pointers in the Poisson model, for instance, where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
 More points scored is not deterministic of winning, however, it is an indicator.
@@
 
-Now that we have fitted the model, lets take a look at how well it performed.
+Now that we have fitted the model, let us take a look at performance.
 Using the tidyr [@-tidyr] package, we make every combination of teams and associated statistics in one dataframe.
-Then wrote custom functions that would take two of the 64 teams as inputs and output the probabilities of winning, predicted amounts of points scored, or probabilities of the score resulting in one of the above mentioned multinomeal categories.
-After that we used the purrr [@R-purrr] package to iterate over all possible games that could be played.
-From there, wrote and ran a web scraping script to obtain a dataframe of the results.
-Then counted the games each model predicted correctly and incorrectly.
-In this case our population is all ${64\choose 2}$ possible games in this tournament and a sample is the 63 games that were played.
+We wrote custom functions tha take two of the 64 teams as inputs and output the probability of winning, predicted amount of points scored, or probabilities of the score resulting in one of the above-mentioned multinomial categories.
+After that, we used the purrr [@R-purrr] package to iterate over all possible games that could be played.
+From there, wrote and ran a web scraping script to obtain a dataframe of the tournament results.
+Finally, we counted the games each model predicted correctly and incorrectly.
+In this case, our population is all ${64\choose 2}$ possible games in this tournament and a sample is the 63 games that were played.
 We are not assuming that we have a random sample from the whole population as these are the best teams in the league.
 Also, there is a selection bias towards games that were played.
-Therefore, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams only are represented once.
-Therefore, we do not claim that these results will hold for games in general.
-However, for the purpose of comparing models that make predictions in this specific tournament this is appropriate.
+For example, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams are represented only once.
+Therefore, we do not claim that these results will hold for all NCAA games.
+However, for the purpose of comparing models that make predictions for this specific tournament is appropriate.
 
@@
 ```
-
-```{r}
-# plot(residuals(fit_poisson))
-# 
-# plot(residuals(logistic_fit))
-```
+m
 ## Conculsion
 
-From these results we see that logistic and poisson model are best at predicting wins and losses.
+From these results, we see that the logistic and Poisson models outperform the multinomial.
 All three models are well over 50% accurate, confirming the claim that GLMs can predict March Madness outcomes better than random chance.
@@
 The limits of this study have a lot to do with data limitations.
-Ideally, we would have more tournaments played with lower performing teams.
+Ideally, we would have more tournaments to test on with lower-skilled teams.
 Also, the results would be much more robust if the tournament was larger.
-We had a lot of data points.
-over 35,000, which is more than sufficient, however, we do not have an abundance of non co-linear covariates.
+We had a lot of data points over 35,000, which is more than sufficient, however, we do not have an abundance of non-co-linear covariates.
 An accuracy of 64.5% is not superb considering betting markets take these exact factors into account, however, being higher than random chance indicates that GLMs are useful in predicting March Madness outcomes.
@@
 While exploring, we found that teams from different parts of the country tended to have a different coefficient on statistics and outcomes.
-A team's conferences is likely a random effect that should be taken into account.
+A team's conferences are likely a random effect that should be taken into account.
 Therefore, additional research into the random effect would certainly yield more robust predictions.

### 2021-04-20 paper/paper.tex
added paper revisions

@@
 The objective is to predict the individual game outcomes as accurately as possible and determine which GLM is the most accurate.\\
-Our response terms is the outcome of the game, \emph{win or loss}, without the possibility of a tie.
+Our response terms is the outcome of the game, \emph{win or loss} without the possibility of a tie.
 Our aim is to derive a function that will accurately predict this using team-level statistics.
-In our dataset, we have many predictors however, we only selected the predictors that are not co-linear.
-First, \emph{field goal percentage} which is the ratio of attempted scores to made scores.
-\emph{Free-throw percentage}, is the the rate of successfully scoring a penalty shot.
+In our dataset, we have many predictors however, we only selected ones that are not co-linear.
+First, \emph{field goal percentage} which is the ratio of attempted scores to successful scores.
+\emph{Free-throw percentage}, is the the rate of scoring a penalty shot.
 Cumulative \emph{Three-point goals made}, field goals made from a sufficient distance.
@@
 \emph{Steals} is when a team was able to remove possession of the ball from the opposing team.
-\emph{Turnover}, the number of times the team lost possession.
-\emph{Blocks}, the number of times the team was able to block a shot made by the opposing team.\\
+\emph{Turnover}, is the number of times the team lost possession.
+\emph{Blocks}, is the number of times the team was able to block a shot made by the opposing team.\\
 March Madness is a winner-take-all tournament and teams do not have a second chance to play a game.
@@
 We did make predictions using that method, however our models will be compared by treating the games as independent of the previous round.
-Meaning, we will filter our data down to include only the 64 teams that qualify, predict every single possible combination of games \({64\choose 2} = 2016\), then use the 63 games played as a sample from the population of all possible games.
-Then use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
-Then we use the results to determine which regression method performs the best.\\
-We will use three regression methods, logistic, poisson, and multinomial, to generate prediction models, compare the results, and make a determination as to which is best suited to this problem.
-Logistic regression most naturally suits this problem because games cannot tie and there is a perfect 50/50 split of wins and losses in our training and testing data.
-This problem could also be suited to poison regression because the number of points scored is roughly poisson distributed.
-After fitting a poisson model, we will predict the number of points scored by Team A, predict the number of points scored by Team B, and take the team with the highest predicted score to be the winner.
-A multinomial model is less naturally suited to this problem, however it may address the shortcomings of the logistic model.
-In particular, many games are close scoring and are won by merely a few points.
-Therefore, we will use multinomeal regression to predict the difference in game score by assigning the differences into five categories.
+Meaning, we will filter our data down to include only the 64 teams that qualify, predict every single one of the \({64\choose 2} = 2016\) possible combination, then use the 63 games played as a sample from the population of all possible games.
+After that, we use basic statistical methods to determine if the predictions were better than random chance, better than betting markets, and better than seeds.
+Then we use the results to determine which regression model performs the best.\\
+We will employ three regression methods, logistic, Poisson, and multinomial, to generate predictive models, compare the results, and decide which is best suited to this problem.
+Logistic regression most naturally suits this problem because games cannot tie and there is a perfectly equal number of wins and losses.
+This problem could also be suited to Poisson regression because the number of points scored is roughly Poisson distributed.
+After fitting a Poisson model, we will predict the number of points scored by Team A, and Team B.
+Then select the team with the highest predicted score to be the predicted winner.
+A multinomial model is less naturally suited to this problem, however, it may address the shortcomings of the logistic model.
+In particular, many games are close scoring and won by merely a few points.
+Therefore, we will use multinomial regression to predict the difference in-game score by assigning the differences into five categories.
 From there, assign a predicted winner based on which outcome the model considers to be the most likely.
@@
 GLMs are accurate relative to randomly guessing March Madness tournament outcomes.
-Team-level statistics, recorded by the NCAA, are highly significant and helpful predicting the tournament outcomes but do not yield fantastic results.
+Team-level statistics, recorded by the NCAA, are highly statistically significant and helpful in predicting the tournament outcomes but do not yield fantastic results.
 Also, when predicting the probability of a win or a loss using a logit link, the coefficients are symmetric because basketball games are a zero-sum-contest.
-From our results, we see that logistic and poisson models are the best at predicting wins and losses.
+From our results, we see that logistic and Poisson models perform the best at predicting wins and losses.
 All three GLM models are well over 50\% accurate, therefore, we claim that using GLM models are more accurate than random chance.
@@
 The logistic model is straight forward, only needing a win to be coded as 1 and a loss to be coded as 0.
-We fit a poisson model by taking a score as a count.
-This is not perfectly suited to poisson because a winning team's score has a mean of 77.13 and a variance of 113.22.
-Regardless we will see how it performs by predicting the score for Team A and Team B, then choose the one with the highest predicted score as the winner.
+We fit a Poisson model by taking a score as a count.
+This is not perfectly suited to Poisson because a winning team's score has a mean of 77.13 and a variance of 113.22.
+Regardless, we will see how it performs by predicting the score for Team A and Team B, then selecting the one with the highest predicted score as the predicted winner.
 
@@
 
-In Table 2, made manually with the kableExtra package (Zhu, 2020), you can see that we have the results for the poisson, logistic, and multinomial models.
-We have a sample of over 35,000 and all of the coefficients are highly significant.
-Which is to be expected because these are the statistics that the NCAA collects as the metrics useful in measuring a team's ability to win games.
-The aim of this study is to compare predictive models so we will not cover it exhaustively or include individual z-statistics and p-values.
-First notice, in the case of the logistic and multinomial models we see that when comparing factors that affect a team's probability of winning, the associated coefficient is nearly equal to the factor capturing the opposing teams metric.
+In Table 2, made with the assistance of the kableExtra package (Zhu, 2020), you can see that we have the results for the Poisson, logistic, and multinomial models.
+We have a sample of over 35,000 and all of the coefficients are highly significant, which is to be expected because these are the statistics that the NCAA collects as the metrics useful in measuring a team's ability to win games.
+This study aims to compare predictive models, so we will not cover it exhaustively or include individual z-statistics and p-values.
+First, the case of the logistic and multinomial models we see that when comparing factors that affect a team's probability of winning, the associated coefficient is nearly equal to the factor capturing the opposing team's metric.
 That is because basketball is a zero-sum-game, anything good for team A is proportionately bad for team B.
-Note this is not true for the poisson model because that is measuring points scored.
-Take three pointers in the poisson model, for instance, where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
+Note this is not true for the poisson model because that is measuring points scored rather than estimating probabilities.
+Take three pointers in the Poisson model, for instance, where an opposing team scores a lot of three pointers has a significant coefficient for the amount of points scored.
 More points scored is not deterministic of winning, however, it is an indicator.
@@
 
-Now that we have fitted the model, lets take a look at how well it performed.
+Now that we have fitted the model, let us take a look at performance.
 Using the tidyr {[}@-tidyr{]} package, we make every combination of teams and associated statistics in one dataframe.
-Then wrote custom functions that would take two of the 64 teams as inputs and output the probabilities of winning, predicted amounts of points scored, or probabilities of the score resulting in one of the above mentioned multinomeal categories.
-After that we used the purrr (Henry \& Wickham, 2020) package to iterate over all possible games that could be played.
-From there, wrote and ran a web scraping script to obtain a dataframe of the results.
-Then counted the games each model predicted correctly and incorrectly.
-In this case our population is all \({64\choose 2}\) possible games in this tournament and a sample is the 63 games that were played.
+We wrote custom functions tha take two of the 64 teams as inputs and output the probability of winning, predicted amount of points scored, or probabilities of the score resulting in one of the above-mentioned multinomial categories.
+After that, we used the purrr (Henry \& Wickham, 2020) package to iterate over all possible games that could be played.
+From there, wrote and ran a web scraping script to obtain a dataframe of the tournament results.
+Finally, we counted the games each model predicted correctly and incorrectly.
+In this case, our population is all \({64\choose 2}\) possible games in this tournament and a sample is the 63 games that were played.
 We are not assuming that we have a random sample from the whole population as these are the best teams in the league.
 Also, there is a selection bias towards games that were played.
-Therefore, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams only are represented once.
-Therefore, we do not claim that these results will hold for games in general.
-However, for the purpose of comparing models that make predictions in this specific tournament this is appropriate.
+For example, the teams that played in the finals are represented 6 times in our sample and 32 of the 64 teams are represented only once.
+Therefore, we do not claim that these results will hold for all NCAA games.
+However, for the purpose of comparing models that make predictions for this specific tournament is appropriate.
 
 \includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
+m
+\#\# Conculsion
 
-\hypertarget{conculsion}{%
-\subsection{Conculsion}\label{conculsion}}
-
-From these results we see that logistic and poisson model are best at predicting wins and losses.
+From these results, we see that the logistic and Poisson models outperform the multinomial.
 All three models are well over 50\% accurate, confirming the claim that GLMs can predict March Madness outcomes better than random chance.
@@
 The limits of this study have a lot to do with data limitations.
-Ideally, we would have more tournaments played with lower performing teams.
+Ideally, we would have more tournaments to test on with lower-skilled teams.
 Also, the results would be much more robust if the tournament was larger.
-We had a lot of data points.
-over 35,000, which is more than sufficient, however, we do not have an abundance of non co-linear covariates.
+We had a lot of data points over 35,000, which is more than sufficient, however, we do not have an abundance of non-co-linear covariates.
 An accuracy of 64.5\% is not superb considering betting markets take these exact factors into account, however, being higher than random chance indicates that GLMs are useful in predicting March Madness outcomes.

### 2021-04-20 paper/paper.Rmd
finished paper

@@
   chunk_output_type: inline
-language:
-  label:
-    fig: "Abbildung "
 ---
+
+<!--
+
+All datasets and scripts associated with this project can be found at https://github.com/despresj/March-Madness
+
+clone this reposityory with the terminal command :
+```zsh
+git clone https://github.com/despresj/March-Madness
+```
+
+-->
 
@@
 
-<!-- All the files for this project is in a github repo found at https://github.com/despresj/March-Madness -->
 ## Introduction
@@
 A perfect bracket is when one correctly predicts the outcome of all 63 games.
-This study uses logistic, poisson, and multinomial regression models fitted with R [@R-base] to predict the outcomes of the March Madness tournament games.
+This study uses logistic, poisson, and multinomial regression models fitted with R [@R-base] to predict the outcomes of the March Madness tournament  games^[Rcripts and datasets assocaited with this project can be found in this [Github Repository](https://github.com/despresj/March-Madness)].
 Our primary objective is to determine which of these GLMs make the most accurate predictions.
@@
 Now that we have fitted the model, let us take a look at performance.
-Using the tidyr [@-tidyr] package, we make every combination of teams and associated statistics in one dataframe.
+Using the tidyr [@R-tidyr] package, we make every combination of teams and associated statistics in one dataframe.
 We wrote custom functions tha take two of the 64 teams as inputs and output the probability of winning, predicted amount of points scored, or probabilities of the score resulting in one of the above-mentioned multinomial categories.
@@
 ```
-m
+
 ## Conculsion

### 2021-04-20 paper/paper.tex
finished paper

@@
 \documentclass[
+  english,
   man,floatsintext]{apa6}
@@
   pdfauthor={Joe Despres \& Sabrina Ball},
+  pdflang={en-EN},
   pdfkeywords={keyword},
@@
 \keywords{keyword\newline\indent Word count: X}
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[main=english]{babel}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
+\fi
 \ifluatex
@@
 A perfect bracket is when one correctly predicts the outcome of all 63 games.
-This study uses logistic, poisson, and multinomial regression models fitted with R (R Core Team, 2020) to predict the outcomes of the March Madness tournament games.
+This study uses logistic, poisson, and multinomial regression models fitted with R (R Core Team, 2020) to predict the outcomes of the March Madness tournament games\footnote{Rcripts and datasets assocaited with this project can be found in this \href{https://github.com/despresj/March-Madness}{Github Repository}}.
 Our primary objective is to determine which of these GLMs make the most accurate predictions.
@@
 Now that we have fitted the model, let us take a look at performance.
-Using the tidyr {[}@-tidyr{]} package, we make every combination of teams and associated statistics in one dataframe.
+Using the tidyr (Wickham, 2021) package, we make every combination of teams and associated statistics in one dataframe.
 We wrote custom functions tha take two of the 64 teams as inputs and output the probability of winning, predicted amount of points scored, or probabilities of the score resulting in one of the above-mentioned multinomial categories.
@@
 \includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
-m
-\#\# Conculsion
+
+\hypertarget{conculsion}{%
+\subsection{Conculsion}\label{conculsion}}
 
@@
 
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+Wickham, H. (2021). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}
+
 \leavevmode\hypertarget{ref-R-tidyverse}{}%

### 2021-04-23 paper/paper.Rmd
final paper

@@
 A perfect bracket is when one correctly predicts the outcome of all 63 games.
-This study uses logistic, poisson, and multinomial regression models fitted with R [@R-base] to predict the outcomes of the March Madness tournament  games^[Rcripts and datasets assocaited with this project can be found in this [Github Repository](https://github.com/despresj/March-Madness)].
+This study uses logistic, poisson, and multinomial regression models fitted with R [@R-base] to predict the outcomes of the March Madness tournament  games^[Scripts and datasets assocaited with this project are published in this [Github Repository](https://github.com/despresj/March-Madness)].
 Our primary objective is to determine which of these GLMs make the most accurate predictions.

### 2021-04-23 paper/paper.tex
final paper

@@
 A perfect bracket is when one correctly predicts the outcome of all 63 games.
-This study uses logistic, poisson, and multinomial regression models fitted with R (R Core Team, 2020) to predict the outcomes of the March Madness tournament games\footnote{Rcripts and datasets assocaited with this project can be found in this \href{https://github.com/despresj/March-Madness}{Github Repository}}.
+This study uses logistic, poisson, and multinomial regression models fitted with R (R Core Team, 2020) to predict the outcomes of the March Madness tournament games\footnote{Scripts and datasets assocaited with this project are published in this \href{https://github.com/despresj/March-Madness}{Github Repository}}.
 Our primary objective is to determine which of these GLMs make the most accurate predictions.
